# Don't Forget: Module 00 {-}

```{r include = FALSE}
source("common.R")
htmltools::tagList(rmarkdown::html_dependency_font_awesome())
library(tidyverse)
library(devtools)
library(DT)
library(vembedr)
library(knitr)
```


This overview is designed to orient you to the class. Please watch the videos from this [playlist][ds4p-pl-00] and work your way through the notes. Although the module-level playlists are embedded in the course, you can find the full-course video playlist [here][ds4p-pl-all]. 



```{r, echo=FALSE}
library(vembedr)
embed_url("https://www.youtube.com/watch?v=HU39nOxcJYg&list=PLKrrdtYgOUYaEAnJX20Ryy4OSie375rVY") %>%
  use_align("center")
```

Data Science for Psychologists (DS4P) introduces on the principles of data science, including:

* data wrangling, 
* modeling, 
* visualization, and 
* communication. 

In this class, we link those principles to psychological methods and open science practices by emphasizing exploratory analyses and description, rather than confirmatory analyses and prediction. Through the semester we will work our way thru [Wickham and Grolemund's R for Data Science text](http://r4ds.had.co.nz/) and develop proficiency with [tidyverse](https://www.tidyverse.org/). This class emphasizes replication and reproducibility. DS4P is a practical skilled-based class and should be useful to students aiming for academia as well as those interested in industry. Applications of these methods can be applied to a full range of psychological areas, including perception (e.g, eye-tracking data), neuroscience (e.g., visualizing neural networks), and individual differences (e.g., valence analysis).

## Big Ideas
This class covers the following broad five areas:

* Reproducibility;
* Replication;
* Robust Methods;
* Resplendent Visualizations; and
* R Programming.


## Materials

### Hardware

This class is requires that you have a laptop that can run R.

### Required Texts

The text is intended to supplement the videos, lecture notes, and in-class tutorials. You probably need to consume all four in order to be successful in this class.

* [R for Data Science text][r4ds]

### Software

#### R and RStudio

```{r, echo=FALSE}
library(vembedr)
embed_url("https://www.youtube.com/watch?v=kVIZGCT5p9U&list=PLKrrdtYgOUYaEAnJX20Ryy4OSie375rVY") %>%
  use_align("center")
```


R is a free software environment for statistical computing and graphics. It compiles and runs on a wide variety of UNIX platforms, Windows, and MacOS.

RStudio is a free _integrated development environment_ (IDE), a powerful user interface for R.

### Git and Github

Git is a version control system. Its original purpose was to help groups of developers work collaboratively on big software projects. Git manages the evolution of a set of files – called a repository – in a structured way. Think of it like the "Track Changes" features from Microsoft Word.

Github is a free IDE and hosting service for Git. As a Wake Forest student, you should be able to access the [GitHub Student Developer Pack for free](https://education.github.com/benefits?type=student). It includes a free PRO upgrade for your github account


## Knowledge is Power

```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=ZbGMbdDLnyY") %>%
  use_align("center")
```

This brief video is covers the icebreaker I do in all of my classes. I encourage you to watch it. In it, I discuss stereotype threats and statistics anxiety.

## Course Modality

This class is a blended class. The online portions are asynchronous. I've created a video highlighting how to be a successful asynchronous learner. 


```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=dc3D2tbl8G0") %>%
  use_align("center")
```

Much of this information comes from Northeastern University's [Tips for Taking Online Classes](https://www.northeastern.edu/graduate/blog/tips-for-taking-online-classes/)




```{r links, child="links.md"}
```
