# (PART) Module 00 {.unnumbered}

# Don't Miss Module 00 {.unnumbered}

```{r opener-setup, include = FALSE}
source("common.R")
library(devtools)
library(DT)
```

This overview is designed to orient you to the class. Please watch the videos from this [playlist][pl_00] and work your way through the notes. Although the module-level playlists are embedded in the course, you can find the full-course video playlist [here][pl_all]. In addition, you can find the slides for this module [here](https://github.com/DataScience4Psych/DataScience4Psych/blob/main/reference/001_Introduction.pdf). Currently, there are seven videos in this playlist. The average video length is 12 minutes, 27 seconds. The total length of the playlist is 1 hour, 27 minutes, 10 seconds. <!-- https://ytplaylist-len.herokuapp.com/ -->

```{r opener-video-html, echo=FALSE,, eval = knitr::is_html_output(),error=FALSE}
"https://www.youtube.com/watch?v=HU39nOxcJYg&list=PLKrrdtYgOUYaEAnJX20Ryy4OSie375rVY" %>%
  embed_url() %>%
  use_align("center")
```
```{r opener-video-latex,echo=FALSE, eval = knitr::is_latex_output()}
embed_youtube_alt("fwJHNw9jU_U")
```

Data Science for Psychologists (DS4P) introduces on the principles of data science, including:

- data wrangling,
- modeling,
- visualization, and
- communication.

This class links those principles to psychological methods and open science practices by emphasizing exploratory analyses and description rather than confirmatory analyses and prediction. Through the semester, we will work our way through [Wickham and Grolemund's R for Data Science text](http://r4ds.had.co.nz/) and develop proficiency with [tidyverse](https://www.tidyverse.org/). This class emphasizes replication and reproducibility. DS4P is a practical skilled-based class and should be helpful to students aiming for academia and those interested in industry. Applications of these methods can be applied to a full range of psychological areas, including perception (e.g., eye-tracking data), neuroscience (e.g., visualizing neural networks), and individual differences (e.g., valence analysis).

## Big Ideas

This class covers the following broad five areas:

- Reproducibility;
- Replication;
- Robust Methods;
- Resplendent Visualizations; and
- R Programming.

## Structure and Organization

Everything in this course is written in R. Yes! Even the slides and the course notes. The course notes are written in R Markdown and rendered with [bookdown](https://bookdown.org/). The slides are written in R Markdown and rendered with [xaringan](https://slides.yihui.org/xaringan/). The labs and homework assignments are written in R Markdown and rendered with [rmarkdown](https://rmarkdown.rstudio.com/). 

Now that's a pretty cool feature ... but it has actual tangible benefits for you! For one, it means that if you want to learn how to do something, you can just look at the source code for the slides or the course notes and see how I did it! But how? Well, all of the course materials are hosted on GitHub. You can access the source code for the slides, course notes, labs, and homework assignments directly from GitHub. This means that you can see how I made the slides and course notes, and you can even make your own edits to them if you want to! This is a great way to learn how to do things in R and to see how I approached different problems. You can also see how the code evolved over time by looking at the commit history on GitHub. For example, you can see that I added in a link to the history of [this file](https://github.com/DataScience4Psych/DataScience4Psych/commits/main/0001_opener.Rmd). This is a great way to learn how to use Git and to see how I worked through different problems or revisions of the course materials. 


### GitHub Organization

You can think of the GitHub organization as the "home" for all of the course materials. 

Everything in this course lives on **GitHub**. What it **looks** like:
- "The syllabus website"   `r emo::ji("right_arrow")` [smasongarrison.github.io/syllabi](https://www.smasongarrison.github.io/syllabi)
- "The course notes site" `r emo::ji("right_arrow")` [DataScience4Psych.github.io/DataScience4Psych](https://DataScience4Psych.github.io/DataScience4Psych/)  (you're looking at them right now)
- "The slides" `r emo::ji("right_arrow")` [DataScience4Psych.github.io/slides/](https://DataScience4Psych.github.io/slides/)
- "GitHub Organization" `r emo::ji("right_arrow")` [github.com/DataScience4Psych](https://github.com/DataScience4Psych/) 


```{r ds4p-structure-diagram, echo=FALSE, message=FALSE, warning=FALSE, fig.align='center', fig.width=10, fig.height=6, fig.alt="Diagram of the DataScience4Psych GitHub organization. Inside the organization are three repositories: Course Notes Repo, Labs and HW Repos, and Slides Repo. The Course Notes Repo points to the Course Notes Website, labeled your home base. The Slides Repo points to Slide Decks, labeled in-class content."}
library(ggplot2)
library(dplyr)
library(grid)

# Box definitions
boxes <- tibble::tribble(
  ~id,             ~xmin, ~xmax, ~ymin, ~ymax, ~label,
  "outer",          0.5,   9.5,   4.8,   9.2,  "github.com/DataScience4Psych\n(GitHub Organization)",
  "course_notes",   1.2,   3.2,   6.0,   7.6,  "Course Notes\nRepo",
  "labs_hw",        3.9,   6.1,   6.0,   7.6,  "Labs & HW\nRepos",
  "slides",         6.8,   8.8,   6.0,   7.6,  "Slides\nRepo",
  "website",        1.1,   3.5,   1.0,   2.6,  "Course Notes Website\n(your home base)",
  "slide_decks",    6.4,   8.9,   1.0,   2.6,  "Slide Decks\n(in-class content)"
)

# Connector lines
segments <- tibble::tribble(
  ~x,   ~y,   ~xend, ~yend,
  2.2,  6.0,  2.2,   3.4,   # course notes down
  7.8,  6.0,  7.8,   3.4    # slides down
)

# Arrow heads
arrows <- tibble::tribble(
  ~x,   ~y,   ~xend, ~yend,
  2.2,  3.4,  2.2,   2.7,
  7.8,  3.4,  7.8,   2.7
)

ggplot() +
  # Outer container
  geom_rect(
    data = filter(boxes, id == "outer"),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "white", color = "black", linewidth = 0.6
  ) +
  # Inner repo boxes
  geom_rect(
    data = filter(boxes, id %in% c("course_notes", "labs_hw", "slides")),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "white", color = "black", linewidth = 0.6
  ) +
  # Lower output boxes
  geom_rect(
    data = filter(boxes, id %in% c("website", "slide_decks")),
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax),
    fill = "white", color = "black", linewidth = 0.6
  ) +
  # Labels
  geom_text(
    data = filter(boxes, !id %in% c("outer")),
    aes(
      x = (xmin + xmax) / 2,
      y = (ymin + ymax) / 2,
      label = label
    ),
    size = 4,
    lineheight = 1.0
  ) +
    geom_text(
    data = filter(boxes, id %in% c("outer")),
    aes(
      x = (xmin + xmax) / 2,
      y = ymax - 0.3,
      label = label
    ),
    size = 4,
    lineheight = 1.0
  ) +
  # Connector lines
  geom_segment(
    data = segments,
    aes(x = x, y = y, xend = xend, yend = yend),
    linewidth = 0.5
  ) +
  # Arrow segments
  geom_segment(
    data = arrows,
    aes(x = x, y = y, xend = xend, yend = yend),
    linewidth = 0.5,
    arrow = arrow(length = unit(0.18, "inches"), type = "closed")
  ) +
  coord_cartesian(xlim = c(0, 10), ylim = c(0, 10), clip = "off") +
  theme_void()
```

What this means is that you can access all of the course materials from the course notes website. The course notes website is your home base. From there, you can click links to access the slides and labs. The slides are where the in-class content lives, and the labs are where the out-of-class content lives. You can also access all of these materials directly from GitHub if you prefer, but I find it easier to navigate through the course notes website. 

There are often embedded slides that you can follow along with. You can open them in their own window if they are too small or do not appear below. This link to them is [here](https://datascience4psych.github.io/slides/d00_onboarding/d00_1_onboarding.html).


```{r embed-welcome-slides, echo=FALSE,error=TRUE}
"https://datascience4psych.github.io/slides/d00_onboarding/d00_1_onboarding.html" %>%
  include_url(height = "400px")
```

To navigation within the slides, you can use the left and right arrows on your keyboard. It's a little trickery on an ipad or phone as sometimes you'll end up swiping to the next section of the book. If you have trouble with it or your screenreader doesn't respond as you'd hope, please reach out! The source code for the slides is available [here](https://github.com/DataScience4Psych/slides/), and you are welcome to make edits to it if you think of ways to make it more accessible or easier to navigate.


## Course Modality

I have taught this class in practically every modality available. <!-- Technically, this course is designated as a blended course. This designation is because this course's modality changes based on the level of COVID-19 transmission in the community. When COVID-19 community transmission is low or medium, this class will be in-person and masking will be required. However, if COVID-19 transmission is high in the community, this course may be moved entirely online in order to protect the health and safety of students and instructor. In the case that we move online, the in-person sessions will become synchronous online sessions. Any changes will be clearly and promptly communicated via email. -->

Pragmatically, the face-to-face portions of the class are -- well -- face-to-face. Or that was the idea anyway... however, during the first semester I taught this course, a few members of the class were on the other side of the planet. Accordingly, I pivoted all the planned in-class activities and labs so that the entire class could complete their degrees on-time. So obviously this last-minute pivot was a little messy, but I think it turned out ok... So again, technically, this class was blended, but effectively, it can be completed from anywhere at any time. It had to be.

### Successful Asynchronous Learning

I've created a video highlighting how to be a successful asynchronous learner.

```{r async-learning-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("dc3D2tbl8G0")
```
```{r async-learning-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=dc3D2tbl8G0" %>%
  embed_url() %>%
  use_align("center")
```

Much of this information comes from Northeastern University's [Tips for Taking Online Classes](https://www.northeastern.edu/graduate/blog/tips-for-taking-online-classes/)

## Learning to Code in the Age of AI

I have thought a lot about how to approach teaching coding in the age of AI. I have a few thoughts on this, which I share in the slide deck below. I've also written a book chapter with two of my former students on this topic, which you can soon find [online](https://teachpsych.org/ebooks/index).

> S. Mason Garrison, Holland K. Tyson, and Xuanyu Lyu (2026). Pragmatic AI Integration in R Teaching: From Frustration to Facilitation. In I. Katzarska-Miller, M. Jackson, & M. Fortner. (Eds.), Integrating generative AI in psychology courses (pp. x–xx ). Society for the Teaching of Psychology.

## Knowledge is Power

```{r knowledge-is-power-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("ZbGMbdDLnyY")
```
```{r knowledge-is-power-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=ZbGMbdDLnyY" %>%
  embed_url() %>%
  use_align("center")
```

This brief video covers the icebreaker I do in all of my classes. I encourage you to watch it. In it, I discuss stereotype threats and statistics anxiety.

## Meet Prof. Mason

```{r meet-prof-mason-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("1IOCKxgowaQ")
```
```{r meet-prof-mason-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=1IOCKxgowaQ" %>%
  embed_url() %>%
  use_align("center")
```

## Website Tour

```{r website-tour-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("Fq_hAQd78zo")
```
```{r website-tour-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=Fq_hAQd78zo" %>%
  embed_url() %>%
  use_align("center")
```

```{r courselinks, child="includes/courselinks.md"}
```
