# (PART\*) Module 00 {-}

# Don't Miss Module 00 {-}

```{r include = FALSE}
source("common.R")
#htmltools::tagList(rmarkdown::html_dependency_font_awesome())
library(tidyverse)
library(devtools)
library(DT)

```

```{r links, child="links.md"}
```


This overview is designed to orient you to the class. Please watch the videos from this [playlist][ds4p-pl-00] and work your way through the notes. Although the module-level playlists are embedded in the course, you can find the full-course video playlist [here][ds4p-pl-all]. In addition, you can find the slides for this module [here](https://github.com/DataScience4Psych/DataScience4Psych/blob/main/admin/pdfs/001_Introduction.pdf). Currently, there are 7 videos in this playlist. The average video length is  12 minutes, 27 seconds. The total length of the playlist is 1 hour, 27 minutes, 10 seconds.

<!--  https://ytplaylist-len.herokuapp.com/ -->


```{r, echo=FALSE}
library(vembedr)
embed_url("https://www.youtube.com/watch?v=HU39nOxcJYg&list=PLKrrdtYgOUYaEAnJX20Ryy4OSie375rVY") %>%
  use_align("center")
```

Data Science for Psychologists (DS4P) introduces on the principles of data science, including:

* data wrangling, 
* modeling, 
* visualization, and 
* communication. 

In this class, we link those principles to psychological methods and open science practices by emphasizing exploratory analyses and description, rather than confirmatory analyses and prediction. Through the semester we will work our way thru [Wickham and Grolemund's R for Data Science text](http://r4ds.had.co.nz/) and develop proficiency with [tidyverse](https://www.tidyverse.org/). This class emphasizes replication and reproducibility. DS4P is a practical skilled-based class and should be useful to students aiming for academia as well as those interested in industry. Applications of these methods can be applied to a full range of psychological areas, including perception (e.g, eye-tracking data), neuroscience (e.g., visualizing neural networks), and individual differences (e.g., valence analysis).

## Big Ideas
This class covers the following broad five areas:

* Reproducibility;
* Replication;
* Robust Methods;
* Resplendent Visualizations; and
* R Programming.


## Course Modality

This class is a blended class. The online portions are asynchronous, and often contain pre-recorded videos. I've created a video highlighting how to be a successful asynchronous learner. 


```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=dc3D2tbl8G0") %>%
  use_align("center")
```

Much of this information comes from Northeastern University's [Tips for Taking Online Classes](https://www.northeastern.edu/graduate/blog/tips-for-taking-online-classes/)

### Productivity During Lockdown


```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=snAhsXyO3Ck") %>%
  use_align("center")
```


## Knowledge is Power

```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=ZbGMbdDLnyY") %>%
  use_align("center")
```

This brief video is covers the icebreaker I do in all of my classes. I encourage you to watch it. In it, I discuss stereotype threats and statistics anxiety.


## Meet Mason

```{r,echo=FALSE}
embed_url("https://www.youtube.com/watch?v=1IOCKxgowaQ") %>%
  use_align("center")
```

## Website Tour

```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=Fq_hAQd78zo") %>%
 use_align("center")
```

