# (PART) Week 1 {-}
```{r include = FALSE}
source("common.R")
```
# Week 1: Overview

This module is designed to <>. Please watch the videos and work your way through the notes. Although the videos are embedded into the course course, you can find the video playlist [here](). 

## Video Roadmap
- Big Ideas
	- What is Exploratory Data Analysis
	- Exploratory vs. Confirmatory Analyses
	- How to succeed
- Getting Set Up with R


# R Examples
## Big Ideas

A first example: Heights of the highest points by state

```{r}
# data-read
## load required packages and data
library(tidyverse)
options(tibble.print_min = 15)
heights = read_csv("data/highest-points-by-state.csv")
## switch from feet to meters
heights$elevation = heights$elevation * .3048

## data-print
heights

## data-arrange-ascending
arrange(heights, elevation)

## data-arrange-descending
arrange(heights, desc(elevation))

## stem-and-leaf
stem(heights$elevation)

## density-estimate
ggplot(heights, aes(x = elevation)) + geom_density()

## density-estimate-plus-rug
ggplot(heights, aes(x = elevation)) + geom_density() + geom_rug()

```

