# (PART) Week 1 {-}

```{r include = FALSE}
source("common.R")
library(vembedr)
```
# Week 1: Overview  {-}

This module is designed to <>. Please watch the videos and work your way through the notes. Although the videos are embedded into the course notes, you can find the video playlist [here](). 

## Week Roadmap  {-}
- What kind of cool stuff are we learning to do?
- What is Exploratory Data Analysis?
	- Exploratory vs. Confirmatory Analyses
- Get your R act together



# What kind of cool stuff are we learning?

```{r, echo=FALSE}
# install.packages("devtools")
# devtools::install_github("gadenbuie/tweetrmd")

library(tweetrmd)
include_tweet("https://twitter.com/Lesdoggg/status/1346584128368508930")
```

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">YASSSSSSSSSS MY LOVE STEVE IS BACK!!! <a href="https://twitter.com/hashtag/KornackiThirstcontinues?src=hash&amp;ref_src=twsrc%5Etfw">#KornackiThirstcontinues</a> <a href="https://t.co/ynK4D87Bhr">pic.twitter.com/ynK4D87Bhr</a></p>&mdash; Leslie Jones 🦋 (@Lesdoggg) <a href="https://twitter.com/Lesdoggg/status/1346584128368508930?ref_src=twsrc%5Etfw">January 5, 2021</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

# What is Exploratory Data Analysis?

```{r, echo=FALSE}
library(vembedr)
embed_url("https://www.youtube.com/watch?v=kVIZGCT5p9U") %>%
  use_align("center")
```

## A first example

<!--Original content: https://jfukuyama.github.io/teaching/stat670/notes/lecture-1.R -->

A first example: Heights of the highest points by state

```{r}
# data-read
## load required packages and data
library(tidyverse)
options(tibble.print_min = 15)
heights = read_csv("data/highest-points-by-state.csv")
## switch from feet to meters
heights$elevation = heights$elevation * .3048

## data-print
heights

## data-arrange-ascending
arrange(heights, elevation)

## data-arrange-descending
arrange(heights, desc(elevation))

## stem-and-leaf
stem(heights$elevation)

## density-estimate
ggplot(heights, aes(x = elevation)) + geom_density()

## density-estimate-plus-rug
ggplot(heights, aes(x = elevation)) + geom_density() + geom_rug()

```

