# (PART) Module 01 {-}


```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
library(knitr)
```

# Welcome to Data Science

This module is designed to introduce you to data science. Please watch the videos and work your way through the notes. You can find the module playlist [here](https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup). Most of the slides used to make the videos in this module can be found in the [slides repo](https://github.com/DataScience4Psych/slides).


## Module Materials

* Slides
  * [Welcome Slides](https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html)
  * [Meet the toolkit](https://datascience4psych.github.io/slides/u1_d02-meet-the-toolkit/u1_d02-meet-the-toolkit.html)

* Activities
  * [UN Voting](https://github.com/DataScience4Psych/ae-01a-un-votes)
  * [Covid Data](https://github.com/DataScience4Psych/ae-01b-covid)
  * [Bechdal Test](https://github.com/DataScience4Psych/ae-02-bechdel-rmarkdown)

* Lab
	* [Hello R](#lab01)

## What is Data Science?

```{r, echo=FALSE}
embed_url("https://www.youtube.com/watch?v=BpKXkkU-NiY&list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup") %>%
  use_align("center")

```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html) if they do not appear below.


```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html"

include_url(var_url, height = "400px")
```

### Fun examples
I've embedded a few examples below.

#### Shiny App

```{r, echo=FALSE}
knitr::include_app("https://minecr.shinyapps.io/unvotes/")
```


#### Hans Rosling

The below video is the shorter version. Hans Rosling's 200 Countries, 200 Years, 4 Minutes - The Joy of Stats

```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=jbkSRLYSojo&list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup") %>%
  use_align("center")
```


You can find a longer talk-length version below.

```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=hVimVzgtD6w") %>%
  use_align("center")
```

#### Social Media

Social media contains a ton of great (and terrible examples of data science in action. These examples range from entire subreddits, such as
[/r/DataisBeautiful](https://www.reddit.com/r/dataisbeautiful/) ([be sure to check out the highest voted posts](https://www.reddit.com/r/dataisbeautiful/top/?sort=top&t=all)) to celebrity tweets about data scientists.

```{r, echo=FALSE}

include_tweet("https://twitter.com/Lesdoggg/status/1346584128368508930")
```


## Course structure and some other useful things

```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=3mz3HNyew-s&list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup") %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html#28) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html#28"

include_url(var_url, height = "400px")
```


### Activity 01

You can do either activity. The choice is yours.

#### UN Votes
```{r, echo=FALSE}
include_url("https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html#25", height = "400px")

```


You can find the materials for the UN activity [here](https://github.com/DataScience4Psych/ae-01a-un-votes). The compiled version should look something like the following...

```{r, echo=FALSE}

var_url="https://datascience4psych.github.io/ae-01a-un-votes/unvotes.html"

include_url(var_url, height = "400px")
```


#### Covid Data

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d01-welcome/u1_d01-welcome.html#26"

include_url(var_url, height = "400px")
```


You can find the materials for the Covid version of this activity [here](https://github.com/DataScience4Psych/ae-01b-covid). The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-01b-covid/covid.html"

include_url(var_url, height = "400px")
```



## Meet our toolbox!

```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=SJaQtRLFb-Y&list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup") %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d02-meet-the-toolkit/u1_d02-meet-the-toolkit) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d02-meet-the-toolkit/u1_d02-meet-the-toolkit.html"

include_url(var_url, height = "400px")
```


### Activity 02: Bechdel

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d02-meet-the-toolkit/u1_d02-meet-the-toolkit#24"

include_url(var_url, height = "400px")
```


You can find the materials for the Bechdel activity [here][Bechdel]. The compiled version should look something like the following...



```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-02-bechdel-rmarkdown/bechdel.html"

include_url(var_url, height = "400px")
```

### Install R and RStudio {#install}


```{r}
library(vembedr)
embed_url("https://www.youtube.com/watch?v=kVIZGCT5p9U") %>%
  use_align("center")
```

* Install [R, a free software environment for statistical computing and graphics][r-proj] from [CRAN][cran], the Comprehensive R Archive Network. I __highly recommend__ you install a precompiled binary distribution for your operating system -- use the links up at the top of the CRAN page linked above!

* Install RStudio's IDE (stands for _integrated development environment_), a powerful user interface for R. Get the Open Source Edition of RStudio Desktop.

  - You can run either the [Preview version][rstudio-preview] or the official releases available [here][rstudio-official].
  - RStudio comes with a __text editor__, so there is no immediate need to install a separate stand-alone editor.
  - RStudio can __interface with Git(Hub)__. However, you must do all the Git(Hub) set up [described elsewhere][happy-git] before you can take advantage of this.
    
If you have a pre-existing installation of R and/or RStudio, I __highly recommend__ that you reinstall both and get as current as possible. It can be considerably harder to run old software than new.

* If you upgrade R, you will need to update any packages you have installed. The command below should get you started, though you may need to specify more arguments if, e.g., you have been using a non-default library for your packages.

``` r
update.packages(ask = FALSE, checkBuilt = TRUE)
```

__Note:__ this  code will only look for updates on CRAN. So if you use a package that lives *only* on GitHub or if you want a development version from GitHub, you will need to  update manually, e.g. via `devtools::install_github()`.


#### Testing testing

* Do whatever is appropriate for your OS to launch RStudio. You should get a window similar to the screenshot you see [here][rstudio-workbench], but yours will be more boring because you haven't written any code or made any figures yet!

* Put your cursor in the pane labeled Console, which is where you interact with the live R process. Create a simple object with code like `x <- 3 * 4` (followed by enter or return). Then inspect the `x` object by typing `x` followed by enter or return. You should see the value 12 print to screen. If yes, you've succeeded in installing R and RStudio.

#### Add-on packages

R is an extensible system and many people share useful code they have developed as a _package_ via CRAN and GitHub. To install a package from CRAN, for example the [dplyr][dplyr-cran] package for data manipulation, here is one way to do it in the R console (there are others).

```r
install.packages("dplyr", dependencies = TRUE)
```

By including `dependencies = TRUE`, we are being explicit and extra-careful to install any additional packages the target package, dplyr in the example above, needs to have around.

You could use the above method to install the following packages, all of which we will use:

* tidyr, [package webpage][tidyr-web]
* ggplot2, [package webpage][ggplot2-web]


#### Further resources

The above will get your basic setup ready but here are some links if you are interested in reading a bit further.

* [How to Use RStudio][rstudio-support]
* [RStudio's leads for learning R][rstudio-R-help]
* [R FAQ][cran-faq]
* [R Installation and Administration][cran-R-admin]
* [More about add-on packages in the R Installation and Administration Manual][cran-add-ons]
    
### Git and Github

#### What is Github?
```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=w3jLJU7DT5E") %>%
  use_align("center")
```

```{r links, child="links.md"}
```
