---
title: "0101.2_activities"
author: "Cynthia Jiao"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


install.packages("devtools")
## install.packages("tweetrmd") 
## library(tweetrmd) # ... embedding tweets - package not available???
install.packages("vembedr")
library(devtools)
if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(vembedr)

##open the file
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```


## Module 1 In-Class Activity

You can do either activity. The choice is yours.

* Activities
  * [UN Voting][ae01a_unvotes]
  * [Covid Data][ae01b_covid]
  * [Bechdal Test][ae02_bechdel]
  * [Oh My Git](https://ohmygit.org/)

## UN Voting

```{r, echo=FALSE}
slide_url(ds4p_urls, "d01_welcome", "#25") %>%
  include_url(height = "400px")
```

You can find the materials for the UN activity [here][ae01a_unvotes]. The compiled version should look something like the following...

```{r, echo=FALSE}
"https://datascience4psych.github.io/ae01a_unvotes/unvotes.html" %>%
  include_url(height = "400px")
```

## COVID Data

```{r, echo=FALSE}
slide_url(ds4p_urls, "d01_welcome", "#26") %>%
  include_url(height = "400px")
```

You can find the materials for the Covid version of this activity [here][ae01b_covid]. The compiled version should look something like the following...

```{r, echo=FALSE}
"https://datascience4psych.github.io/ae01b_covid/covid.html" %>%
  include_url(height = "400px")

```

## Bechdel Activity

```{r, echo=FALSE}
slide_url(ds4p_urls, "d02_toolkit", "#24") %>%
  include_url(height = "400px")
```

You can find the materials for the Bechdel activity [here][ae02_bechdel]. The compiled version should look something like the following...

```{r, echo=FALSE}
  include_url("https://datascience4psych.github.io/ae-02-bechdel-rmarkdown/bechdel.html",
              height = "400px")
```

## Oh My Git

```{r, echo=FALSE}
  include_url("https://ohmygit.org/",
              height = "400px")
```

