
# Getting Started with Github

These instructions are adapted from Jenny Bryan's [happygitwithr](https://happygitwithr.com/). I encourage you to check out the unabridged version as it has so much more detail.

## Half the battle {#install-intro}



Getting all the necessary software installed, configured, and playing nicely together is honestly half the battle when first adopting Git. Brace yourself for some pain. The upside is that you can give yourself a pat on the back once you get through this. And you WILL get through this.

You will find far more resources for how to *use Git* than for installation and configuration. Why? The experts ...

  * Have been doing this for years. It's simply not hard for them anymore.
  * Probably use some flavor of Unix. They may secretly (or not so secretly) take pride in neither using nor knowing Windows.
  * Get more satisfaction and reward for thinking and writing about Git concepts and workflows than Git installation.

In their defense, it's hard to write installation instructions. Failures can be specific to an individual OS or even individual computer. If you have some new problem and, especially, the corresponding solution, [she'd love to hear from you!](https://github.com/jennybc/happy-git-with-r/issues)

## Register a GitHub account {#github-acct}

Register an account with GitHub. It's free!

  * <https://github.com>

### Username advice

You will be able to upgrade to a paid level of service, apply discounts, join organizations, etc. in the future, so don't fret about any of that now. **Except your username. You might want to give that some thought.**

A few tips, which sadly tend to contradict each other:

  * Incorporate your actual name! People like to know who they're dealing with. Also makes your username easier for people to guess or remember.
  * Reuse your username from other contexts, e.g., Twitter or Slack. But, of course, someone with no GitHub activity will probably be squatting on that.
  * Pick a username you will be comfortable revealing to your future boss.
  * Shorter is better than longer.
  * Be as unique as possible in as few characters as possible. In some settings GitHub auto-completes or suggests usernames.
  * Make it timeless. Don't highlight your current university, employer, or place of residence, e.g. JennyFromTheBlock.
  * Avoid words laden with special meaning in programming. In my first inept efforts to script around the GitHub API, I assigned lots of issues to [the guy with username `NA`](https://github.com/na) because my vector of GitHub usernames contained missing values. A variant of [Little Bobby Tables](https://xkcd.com/327/).
  * Avoid the use of upper vs. lower case to separate words. We highly recommend all lowercase. GitHub treats usernames in a case insensitive way, but using all lowercase is kinder to people doing downstream regular expression work with usernames, in various languages. A better strategy for word separation is to use a hyphen `-` or underscore `_`.

You can change your username later, but better to get this right the first time.

  * <https://help.github.com/articles/changing-your-github-username/>
  * <https://help.github.com/articles/what-happens-when-i-change-my-username/>

### Free private repos

GitHub offers free unlimited private repositories for all users. These free private repositories support up to three external collaborators, making them a perfect place for your personal projects, for job applications, and testing things out before making your project open source.

Go ahead and register your free account NOW and then pursue any special offer that applies to you:

  * Students, faculty, and educational/research staff: [GitHub Education](https://education.github.com).
    - GitHub "Organizations" can be extremely useful for courses or research/lab groups, where you need some coordination across a set of repos and users.
  * Official nonprofit organizations and charities: [GitHub for Good](https://github.com/nonprofit)
  
## Install Git {#install-git}

You need Git, so you can use it at the command line and so RStudio can call it.

If there's any chance it's installed already, verify that, rejoice, and skip this step.

Otherwise, find installation instructions below for your operating system.

```{r setup, include = FALSE}
has_bash <- Sys.which('bash') != '' && .Platform$OS.type != 'windows'
```

### Git already installed?

Go to the shell (Appendix \@ref(shell)). Enter `which git` to request the path to your Git executable:

```{bash collapse = TRUE, eval = has_bash}
which git
```

and `git --version` to see its version:
 
```{bash collapse = TRUE, eval = has_bash}
git --version
```

If you are successful, that's great! You have Git already. No need to install! Move on.

If, instead, you see something more like `git: command not found`, keep reading.

macOS users might get an immediate offer to install command line developer tools. Yes, you should accept! Click "Install" and read more below.

## Windows  {#install-git-windows}

**Option 1** (*highly recommended*): Install [Git for Windows](https://git-for-windows.github.io/), also known as `msysgit` or "Git Bash", to get Git in addition to some other useful tools, such as the Bash shell. Yes, all those names are totally confusing, but you might encounter them elsewhere and I want you to be well-informed.

We like this because Git for Windows leaves the Git executable in a conventional location, which will help you and other programs, e.g. RStudio, find it and use it. This also supports a transition to more expert use, because the "Git Bash" shell will be useful as you venture outside of R/RStudio.

  *  **NOTE:** When asked about "Adjusting your PATH environment", make sure to select "Git from the command line and also from 3rd-party software". Otherwise, we believe it is good to accept the defaults.
  * Note that RStudio for Windows prefers for Git to be installed below `C:/Program Files` and this appears to be the default. This implies, for example, that the Git executable on my Windows system is found at `C:/Program Files/Git/bin/git.exe`. Unless you have specific reasons to otherwise, follow this convention.

This also leaves you with a Git client, though not a very good one. So check out Git clients we recommend (chapter \@ref(git-client)).

FYI, this appears to be equivalent to what you would download from here: <https://git-scm.com/download/>.

**Option 2** (*recommended*): Install [Git for Windows](https://git-for-windows.github.io/) via the [Chocolatey](https://chocolatey.org) package manager. If this means anything to you, Chocolatey is like [`apt-get`](https://en.wikipedia.org/wiki/APT_(Debian)) or [Homebrew](https://brew.sh), but for Windows instead of Debian/Ubuntu Linux or macOS. As far as I can tell, using Chocolatey to install Git for Windows gives the same result as installing it yourself (option 1).

This obviously requires that you already have [Chocolatey](https://chocolatey.org) installed or that you are up for installing it. It is not hard and the [instructions are here](https://chocolatey.org/install). This may be worthwhile if it seems likely you will be installing more open source software in the future.

After you install Chocolatey, in a shell (Appendix \@ref(shell)), do:

``` bash
choco install git.install
```

This installs the most current [Git (Install) X.Y.Z](https://chocolatey.org/packages/git.install) Chocolatey package. At the time of writing, that is "Git (Install) 2.29.2.2", but that version number will increment over time.

If you [search Chocolatey packages](https://chocolatey.org/packages) yourself, you might see two packages that install Git -- "Git (Install) 2.29.2.2" and "Git 2.29.2.2", at the time of writing. I believe "Git (Install) 2.29.2.2" is technically the more correct, but I also think it doesn't really matter which one you use. A rather confusing explanation is found [here](https://chocolatey.org/docs/chocolatey-faqs#what-is-the-difference-between-packages-no-suffix-as-compared-toinstallportable). Don't worry too much about whether you do `choco install git.install` or `choco install git`.

### macOS

**Option 1** (*highly recommended*): Install the Xcode command line tools (**not all of Xcode**), which includes Git.

Go to the shell and enter one of these commands to elicit an offer to install developer command line tools:

``` bash
git --version
git config
```

Accept the offer! Click on "Install".

Here's another way to request this installation, more directly:

``` bash
xcode-select --install
```

We just happen to find this Git-based trigger apropos.

Note also that, after upgrading macOS, you might need to re-do the above and/or re-agree to the Xcode license agreement. We have seen this cause the RStudio Git pane to disappear on a system where it was previously working. Use commands like those above to tickle Xcode into prompting you for what it needs, then restart RStudio.

**Option 2** (*recommended*): Install Git from here: <http://git-scm.com/downloads>.

  * This arguably sets you up the best for the future. It will certainly get you the latest version of Git of all approaches described here.
  * The GitHub home for the macOS installer is here: <https://github.com/timcharper/git_osx_installer>.
    - At that link, you can find more info if something goes wrong or you are working on an old version of macOS.

**Option 3** (*recommended*): If you anticipate getting heavily into scientific computing, you're going to be installing and updating lots of software. You should check out [Homebrew](http://brew.sh), "the missing package manager for OS X". Among many other things, it can install Git for you. Once you have Homebrew installed, do this in the shell:

```
brew install git
```

## Introduce yourself to Git {#hello-git}

In the shell (Appendix \@ref(shell)):

``` bash
git config --global user.name 'Jane Doe'
git config --global user.email 'jane@example.com'
git config --global --list
```

substituting your name and **the email associated with your GitHub account**.

The [usethis package](https://usethis.r-lib.org) offers an alternative approach. You can set your Git user name and email from within R:

```{r, eval = FALSE}
## install if needed (do this exactly once):
## install.packages("usethis")
library(usethis)
use_git_config(user.name = "Jane Doe", user.email = "jane@example.org")
```

### More about `git config`

An easy way to get into a shell from RStudio is *Tools > Terminal* or *Tools > Shell*. More about the shell in the Appendix \@ref(shell).

Special Windows gotchas: If you are struggling on Windows, consider there are different types of shell and you might be in the wrong one. You want to be in a "Git Bash" shell, as opposed to Power Shell or the legacy `cmd.exe` command prompt. Read more in [the Appendix](#windows-shell-hell). This might also be a reason to do this configuration via the usethis package in R.

What user name should you give to Git? This does not have to be your GitHub user name, although it can be. Another good option is your actual first name and last name. If you commit from different machines, sometimes people work that info into the user name. Your commits will be labelled with this user name, so make it informative to potential collaborators and future you.

What email should you give to Git? This __must__ be the email associated with your GitHub account.

These commands return nothing. You can check that Git understood what you typed by looking at the output of `git config --global --list`.

### Configure the Git editor {#git-editor}

Another Git option that many people eventually configure is the editor. At some point, you will fail to give Git what it wants in terms of a commit message and it will kick you into an editor. This can be distressing, if it's not your editor of choice and you don't even know how to save and quit. You can enforce your will with something along these lines:

``` bash
git config --global core.editor "emacs"
```

Substitute your preferred editor for `"emacs"` here. Software Carpentry's Git lesson has a comprehensive listing of the exact `git config` command needed for [many combinations of OS and editor](http://swcarpentry.github.io/git-novice/02-setup/).
