# RDD: Quick Starting with Github {#quickhappygit}

One of the goals for this course is to familiarize yourself with Git and GitHub. This section is a concise version of a more detailed document you'll find later in the class, combining insights from GitHub Classroom's [github-starter-course](https://github.com/education/github-starter-course) and Jenny Bryan's [happygitwithr](https://happygitwithr.com/). I *strongly* encourage you to check out the unabridged version of [happygit](https://happygitwithr.com/) as it has so much more detail. Alternatively,  [this](https://www.youtube.com/watch?v=8Dd7KRpKeaE) video by [Jess Chan](https://coder-coder.com/about/) of [Coder Coder](https://coder-coder.com/) is an excellent resource for beginners.

```{r include=FALSE}
source("common.R")
```

```{r echo=FALSE}
"https://twitter.com/jmtraver1/status/1449752767510220807" %>%
  try_include_tweet()
```

## The Basics of GitHub and Git


```{r child = './supporting-docs/basicsofgithubandgit.Rmd'}
```


## Getting Started with GitHub 

### Create a GitHub Account {#github-acct}

Sign up at [github.com](https://github.com/join). Choose a thoughtful username as it represents you in the GitHub community. You will be able to upgrade to a paid level of service, apply discounts, join organizations, etc., in the future. 

#### Username advice

You will be able to upgrade to a paid level of service, apply discounts, join organizations, etc. in the future, so don't fret about any of that now. **Except your username. You might want to give that some thought.**

A few tips, which sadly tend to contradict each other:

- Incorporate your actual name! People like to know who they're dealing with. Also makes your username easier for people to guess or remember.
- Reuse your username from other contexts, e.g., Twitter or Slack. But, of course, someone with no GitHub activity will probably be squatting on that.
- Pick a username you will be comfortable revealing to your future boss.
- Shorter is better than longer.
- Be as unique as possible in as few characters as possible. In some settings GitHub auto-completes or suggests usernames.
- Make it timeless. Don't highlight your current university, employer, or place of residence, e.g. JennyFromTheBlock.
- Avoid words laden with special meaning in programming. In Jenny's first inept efforts to script around the GitHub API, she assigned lots of issues to [the guy with username `NA`](https://github.com/na) because my vector of GitHub usernames contained missing values. A variant of [Little Bobby Tables](https://xkcd.com/327/).
- Avoid the use of upper vs. lower case to separate words. We highly recommend all lowercase. GitHub treats usernames in a case insensitive way, but using all lowercase is kinder to people doing downstream regular expression work with usernames, in various languages. A better strategy for word separation is to use a hyphen `-` or underscore `_`.

You can change your username later, but better to get this right the first time.

- <https://help.github.com/articles/changing-your-github-username/>
- <https://help.github.com/articles/what-happens-when-i-change-my-username/>

### Install Git and a Git client

```{r child = './supporting-docs/installgit.Rmd'}
```

```{r child = './supporting-docs/installgitclient.Rmd'}
```


## Half the battle {#install-intro}

```{r child = './supporting-docs/halfthebattle.Rmd'}
```


### What is a Git client? Why would you want one?

```{r child = './supporting-docs/whygitclient.Rmd'}
```

```{r child = './supporting-docs/installgitclient.Rmd'}
```

## 📚 Resources

- [A short video explaining what GitHub is](https://www.youtube.com/watch?v=w3jLJU7DT5E&feature=youtu.be)
- [Git and GitHub learning resources](https://docs.github.com/en/github/getting-started-with-github/git-and-github-learning-resources)
- [Understanding the GitHub flow](https://guides.github.com/introduction/flow/)
- [How to use GitHub branches](https://www.youtube.com/watch?v=H5GJfcp3p4Q&feature=youtu.be)
- [Interactive Git training materials](https://githubtraining.github.io/training-manual/#/01_getting_ready_for_class)
- [GitHub's Learning Lab](https://lab.github.com/)
- [Education community forum](https://education.github.community/)
- [GitHub community forum](https://github.community/)


```{r, echo=FALSE,eval = knitr::is_latex_output()}
 embed_youtube_alt("8Dd7KRpKeaE")
```
```{r, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=8Dd7KRpKeaE" %>%
  embed_url() %>%
  use_align("center")
# jess chan's video from coder coder
```

```{r links, child="admin/md/courselinks.md"}
```

### Oh My Git

Oh My Git! is a game that "visualizes the internal structures of Git repositories in realtime". It's a great way to learn how Git works and to debug problems with your repositories.

```{r, echo=FALSE}
  include_url("https://ohmygit.org/",
              height = "400px")
```


```{r links, child="admin/md/courselinks.md"}