# (PART) Module 02 {-}


```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
```

# Data and Visualization

This module is designed to introduce you to graphical data analysis. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-02]. The slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Data and visualization][d03_dataviz]
  * [Building plots for various data types][d04_ggplot2]
* Suggested Readings
  * [Data visualisation](https://r4ds.had.co.nz/data-visualisation.html)
  * [Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html)
* Activities
  * [Star Wars!][ae-03-starwars-dataviz]
  
* Lab
	* [Plastic waste](#lab02)

# Exploratory Data Analysis
 
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=IUP3qCIjk4Q"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d03_dataviz] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d03_dataviz/d03_dataviz.html"
include_url(var_url, height = "400px")%>%
  use_align("center")
```

## What is in a dataset?
```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d03_dataviz] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d03_dataviz/d03_dataviz.html#23"
include_url(var_url, height = "400px")
```

## Why do we visualize?

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d03_dataviz] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d03_dataviz/d03_dataviz.html#35"
include_url(var_url, height = "400px")
```
# Visualising data with ggplot2

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d04_ggplot2] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d04_ggplot2/d04_ggplot2.html"
include_url(var_url, height = "400px")
```


# Activity 03: Star Wars!

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d04_ggplot2/d04_ggplot2.html#19"
include_url(var_url, height = "400px")
```

You can find the materials for the Star Wars activity [here][ae-03-starwars-dataviz]. The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html"
include_url(var_url, height = "400px")
```

# Visualization Examples

Here are some fun examples of data visualization. Eventually, this section will include examples from previous classes. :) For now, it's going to just include random stuff.

```{r, echo=FALSE}
# Tukey and the Prim9!
embed_url("https://www.youtube.com/watch?v=sN2gCCd2Rr8") %>%
  use_align("center")
```

## Census Reporter Data from North Carolina

<iframe id="cr-embed-04000US37-demographics-age-distribution_by_category" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=demographics-age-distribution_by_category&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Population+by+age+category&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px; float: right;"></iframe>

<iframe id="cr-embed-04000US37-social-language-children" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-language-children&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Language+at+home%2C+children+5-17&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px;float: left;"></iframe>

<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>
<iframe id="cr-embed-04000US37-social-educational_attainment_distribution" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-educational_attainment_distribution&dataYear=2019&releaseID=ACS_2019_1-year&chartType=histogram&chartHeight=200&chartQualifier=Universe%3A+Population+25+years+and+over&chartTitle=Population+by+minimum+level+of+education&initialSort=&statType=scaled-percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 720px; "></iframe>


<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>


