
```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
```

# Data and Visualization: Module 02

This module is designed to <>. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here](https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup). Most of the slides used to make the videos in this module can be found in the [slides repo](https://github.com/DataScience4Psych/slides).


## Module Materials

* Slides
  * [Data and visualization](https://datascience4psych.github.io/slides/u1_d03-data-viz-1/u1_d03-data-viz-1.html)
  * [Building plots for various data types](https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html)

* Activities
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
  
* Lab
	* [Plastic waste]


## Data and visualization!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8&") %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d03-data-viz-1/u1_d03-data-viz-1.html) if they do not appear below.

<iframe seamless width="100%" frameborde="0" src="https://datascience4psych.github.io/slides/u1_d03-data-viz-1/u1_d03-data-viz-1.html"></iframe>


## Building plots for various data types!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8&") %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html) if they do not appear below.

<iframe seamless width="100%" frameborde="0" src="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html"></iframe>


## Activity 03: Star Wars!

<iframe seamless width="100%" frameborde="0" src="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html#19"></iframe>

You can find the materials for the Star Wars activity [here](https://github.com/DataScience4Psych/ae-03-starwars-dataviz). The compiled version should look something like the following...

<iframe seamless width="100%" frameborde="0" src="https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html"></iframe>

