# (PART) Module 02 {-}


```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```



# Welcome to Data and Visualization

This module is designed to introduce you to exploratory and graphical data analysis. 
Please watch the videos and work your way through the notes. **The videos start on the next page.**  
You can find the video playlist for this module [here][pl_02]. The slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Data and visualization][d03_dataviz]
  * [Visualizing data with ggplot2][d04_ggplot2]
  * [Visualizing numerical data][d05_viznum]
  * [Visualizing categorical data][d06_vizcat]
* Suggested Readings
  * All subchapters of this module, including
      *  [Basic care and feeding of data in R](#basic_data_care)
  * R4DS
      * [Data Exploratation Section](https://r4ds.had.co.nz/explore-intro.html), including
      * [Data visualization](https://r4ds.had.co.nz/data-visualization.html)
      * [Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html)
* Activities
  * [Star Wars!][ae03_starwars]
* Lab
  * [Plastic waste](#lab02)
  
### Estimated Video Length

<!-- https://www.youtube.com/playlist?list=PLKrrdtYgOUYZmr_T3PnuxjVIlj0C0kUNI & https://ytplaylist-len.herokuapp.com/ -->

No of videos : 9

Average length of video : 13 minutes, 34 seconds

Total length of playlist : 2 hours, 2 minutes, 11 seconds

# Exploratory Data Analysis
 
You can follow along with the slides [here][d03_dataviz] if they do not appear below.
 
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=IUP3qCIjk4Q"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}

var_url=slide_url(ds4p_urls,"d03_dataviz")
include_url(var_url[1], height = "400px")
```

## What is in a dataset?

```{r, echo=FALSE}
# 
"https://www.youtube.com/watch?v=J6arziOCQpk" %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
# is unhappy for some reason
var_url=slide_url(ds4p_urls,"d03_dataviz","#23")
include_url(var_url[1], height = "400px")
```

### Why do we visualize?

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=E5IY1FQv0OM" %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
# is unhappy for some reason
var_url=slide_url(ds4p_urls,"d03_dataviz","#35")
include_url(var_url[1], height = "400px")
```

# Visualizing data with ggplot2

```{r, echo=FALSE}
"https://twitter.com/valeriepolad/status/1388733898926870533" %>%
include_tweet()

"https://twitter.com/_stevenshoe/status/1469332905503731713" %>%
include_tweet()
```



You can follow along with the slides [here][d04_ggplot2] if they do not appear below.

```{r, echo=FALSE}

"https://www.youtube.com/watch?v=-tdEQXdx-SA" %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}

var_url=slide_url(ds4p_urls,"d04_ggplot2")
include_url(var_url, height = "400px")
```


## ggplot2 and aesthetics

```{r, echo=FALSE}
# Aesthetics 
"https://www.youtube.com/watch?v=K2u2_lcgdks" %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}

var_url=slide_url(ds4p_urls,"d04_ggplot2","#26")
include_url(var_url, height = "400px")
```


# Visualizing numerical data

You can follow along with the slides [here][d05_viznum] if they do not appear below.

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=hGylk9UaM2k"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}

var_url=slide_url(ds4p_urls,"d05_viznum")
include_url(var_url, height = "400px")
```


## Looking at Data

```{r, echo=FALSE}
# Tukey and the Prim9!
embed_url("https://www.youtube.com/watch?v=sN2gCCd2Rr8") %>%
  use_align("center")
```

## More on visualizing numerical data

Fun fact, when you screen capture with [f.lux](https://justgetflux.com/) running in the background, f.lux is captured too.

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=7OWyhBWMoM8" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}

slide_url(ds4p_urls,"d05_viznum","#12") %>%
include_url(height = "400px")
```

# Visualizing categorical data

You can follow along with the slides [here][d06_vizcat] if they do not appear below.

```{r, echo=FALSE}

"https://www.youtube.com/watch?v=xY819zNKdnY" %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}

slide_url(ds4p_urls,"d06_vizcat") %>%
include_url( height = "400px")
```



# Star Wars Activity!

```{r, echo=FALSE}

slide_url(ds4p_urls,"d06_vizcat","#15") %>%
include_url(height = "400px")
```


You can find the materials for the Star Wars activity [here][ae03_starwars]. The compiled version should look something like the following...

```{r, echo=FALSE}
"https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html" %>%
include_url(height = "400px")
```


# Visualization Examples

Here are some fun examples of data visualization. Eventually, this section will include examples from previous classes. :) For now, it's going to just include random stuff.

## Census Reporter Data from North Carolina

<iframe id="cr-embed-04000US37-demographics-age-distribution_by_category" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=demographics-age-distribution_by_category&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Population+by+age+category&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px; float: right;"></iframe>

<iframe id="cr-embed-04000US37-social-language-children" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-language-children&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Language+at+home%2C+children+5-17&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px;float: left;"></iframe>

<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>
<iframe id="cr-embed-04000US37-social-educational_attainment_distribution" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-educational_attainment_distribution&dataYear=2019&releaseID=ACS_2019_1-year&chartType=histogram&chartHeight=200&chartQualifier=Universe%3A+Population+25+years+and+over&chartTitle=Population+by+minimum+level+of+education&initialSort=&statType=scaled-percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 720px; "></iframe>


<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>


```{r links, child="admin/md/courselinks.md"}
```