# (PART) Module 02 {-}


```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
library(knitr)
```

# Data and Visualization

This module is designed to <>. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-02]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Data and visualization][u1_d03-data-viz-1]
  * [Building plots for various data types][u1_d04-data-viz-2]

* Activities
  * [Star Wars!][ae-03-starwars-dataviz]
  
* Lab
	* [Plastic waste](#lab02)



## Data and visualization!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8") %>%
  use_align("center")
```

You can follow along with the slides [here][u1_d03-data-viz-1] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d03-data-viz-1/u1_d03-data-viz-1.html"
include_url(var_url, height = "400px")
```


## Building plots for various data types!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8") %>%
  use_align("center")
```

You can follow along with the slides [here][u1_d04-data-viz-2] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html"
include_url(var_url, height = "400px")
```


## Activity 03: Star Wars!

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html#19"
include_url(var_url, height = "400px")
```

You can find the materials for the Star Wars activity [here][ae-03-starwars-dataviz]. The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html"
include_url(var_url, height = "400px")
```

## Visualization Examples

Here are some fun examples of data visualization.

### Pedigrees

```{r, echo=FALSE}
# Who Would Be King of America if George Washington had been made a monarch?
embed_url("https://www.youtube.com/watch?v=ZxnBveop5no") %>%
  use_align("center")
```



### Census Reporter Data from North Carolina

<iframe id="cr-embed-04000US37-demographics-age-distribution_by_category" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=demographics-age-distribution_by_category&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Population+by+age+category&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px; float: right;"></iframe>

<iframe id="cr-embed-04000US37-social-language-children" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-language-children&dataYear=2019&releaseID=ACS_2019_1-year&chartType=pie&chartHeight=200&chartQualifier=&chartTitle=Language+at+home%2C+children+5-17&initialSort=-value&statType=percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 300px;float: left;"></iframe>

<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>
<iframe id="cr-embed-04000US37-social-educational_attainment_distribution" class="census-reporter-embed" src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/iframe.html?geoID=04000US37&chartDataID=social-educational_attainment_distribution&dataYear=2019&releaseID=ACS_2019_1-year&chartType=histogram&chartHeight=200&chartQualifier=Universe%3A+Population+25+years+and+over&chartTitle=Population+by+minimum+level+of+education&initialSort=&statType=scaled-percentage" frameborder="0" width="100%" height="300" style="margin: 1em; max-width: 720px; "></iframe>


<script src="https://s3.amazonaws.com/embed.censusreporter.org/1.0/js/embed.chart.make.js"></script>


