# (PART) Module 02 {-}

```{r visualization-setup, include = FALSE}
source("common.R")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")
```

# Welcome to Data and Visualization

This module is designed to introduce you to exploratory and graphical data analysis.
Please watch the videos and work your way through the notes. **The videos start on the next page.**  
You can find the video playlist for this module [here][pl_02]. The slides used to make the videos in this module can be found in the [slides repo][course_slides].

## Module Materials

* Slides from Lectures
  * [Data and visualization][d03_dataviz]
  * [Visualizing data with ggplot2][d04_ggplot2]
  * [Visualizing numerical data][d05_viznum]
  * [Visualizing categorical data][d06_vizcat]
* Suggested Readings
  * All subchapters of this module, including
    * [Basic care and feeding of data in R](#basiccare)
  * R4DS
    * [Data Exploratation Section](https://r4ds.had.co.nz/explore-intro.html), including
    * [Data visualization](https://r4ds.had.co.nz/data-visualization.html)
    * [Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html)
* Activities
  * [Star Wars!][ae03_starwars]
* Lab
  * [Plastic waste](#lab02)
  
### Estimated Video Length

<!-- https://www.youtube.com/playlist?list=PLKrrdtYgOUYZmr_T3PnuxjVIlj0C0kUNI & https://ytplaylist-len.herokuapp.com/ -->

No of videos : 9

Average length of video : 13 minutes, 34 seconds

Total length of playlist : 2 hours, 2 minutes, 11 seconds

# Exploratory Data Analysis

You can follow along with the slides [here][d03_dataviz] if you would like to open them full-screen.

```{r eda-intro-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("IUP3qCIjk4Q")
```
```{r eda-intro-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=IUP3qCIjk4Q" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-dataviz-slides, echo=FALSE}
var_url <- slide_url(ds4p_urls, "d03_dataviz")
include_url(var_url[1], height = "400px")
```

## What is in a dataset?

```{r dataset-intro-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("J6arziOCQpk")
```
```{r dataset-intro-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=J6arziOCQpk" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-dataset-slides, echo=FALSE}
# is unhappy for some reason
var_url <- slide_url(ds4p_urls, "d03_dataviz", "#25")
include_url(var_url[1], height = "400px")
```

### Why do we visualize?

```{r why-visualize-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("E5IY1FQv0OM")
```
```{r why-visualize-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=E5IY1FQv0OM" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-why-visualize-slides, echo=FALSE}
# is unhappy for some reason
var_url <- slide_url(ds4p_urls, "d03_dataviz", "#37")
include_url(var_url[1], height = "400px")
```

```{r embed-visualization-tweet, echo=FALSE}
"https://twitter.com/LinChujun/status/1478075789379653633" %>%
  try_include_tweet()
```

Data visualization is a powerful tool for understanding and communicating insights from data. Research has shown that effective visualizations can reveal patterns, relationships, and anomalies that might be missed in raw numerical data [@lin2022power].


## What we learned

```{r embed-why-visualize-learning, echo=FALSE}
# is unhappy for some reason
var_url <- slide_url(ds4p_urls, "d03_dataviz", "#45")
include_url(var_url[1], height = "400px")
```


# LEC: Visualizing data with ggplot2

```{r embed-ggplot2-tweets, echo=FALSE}
"https://twitter.com/valeriepolad/status/1388733898926870533" %>%
  try_include_tweet()

"https://twitter.com/_stevenshoe/status/1469332905503731713" %>%
  try_include_tweet()
```

You can follow along with the slides [here][d04_ggplot2] if you would like to open them full-screen. This lecture will break down the components of a ggplot2 graph and how to use it to visualize data in R. 

```{r ggplot2-intro-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("-tdEQXdx-SA")
```
```{r ggplot2-intro-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=-tdEQXdx-SA" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-ggplot2-slides, echo=FALSE}
slide_url(ds4p_urls, "d04_ggplot2") %>%
  include_url(height = "400px")
```


## Learning goals for ggplot2

```{r embed-ggplot2-slides-goals, echo=FALSE}
slide_url(ds4p_urls, "d04_ggplot2","#4") %>%
  include_url(height = "400px")
```


## ggplot2 and aesthetics

```{r aesthetics-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("K2u2_lcgdks")
```
```{r aesthetics-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=K2u2_lcgdks" %>%
  embed_url() %>%
  use_align("center")
# Aesthetics
```

```{r embed-aesthetics-slides, echo=FALSE}
slide_url(ds4p_urls, "d04_ggplot2", "#28") %>%
  include_url(height = "400px")
```


# LEC: Visualizing numerical data

You can follow along with the slides [here][d05_viznum] if you would like to open them full-screen.

```{r viznum-intro-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("hGylk9UaM2k")
```
```{r viznum-intro-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=hGylk9UaM2k" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-viznum-slides, echo=FALSE}
slide_url(ds4p_urls, "d05_viznum") %>%
  include_url(height = "400px")
```

## Looking at Data

```{r tukey-prim9-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("sN2gCCd2Rr8")
```
```{r tukey-prim9-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=sN2gCCd2Rr8" %>%
  embed_url() %>%
  use_align("center")
# Tukey and the Prim9!
```

## More on visualizing numerical data

Fun fact, when you screen capture with [f.lux](https://justgetflux.com/) running in the background, f.lux is captured too.

```{r viznum-details-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("7OWyhBWMoM8")
```
```{r viznum-details-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=7OWyhBWMoM8" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-viznum-details-slides, echo=FALSE}
slide_url(ds4p_urls, "d05_viznum", "#14") %>%
  include_url(height = "400px")
```


### What we learned about visualizing numerical data

```{r embed-viznum-learned-slides, echo=FALSE}
slide_url(ds4p_urls, "d05_viznum", "#41") %>%
  include_url(height = "400px")
```

# Lec: Visualizing categorical data

You can follow along with the slides [here][d06_vizcat] if you would like to open them full-screen.

```{r vizcat-intro-video-latex, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("xY819zNKdnY")
```
```{r vizcat-intro-video-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=xY819zNKdnY" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-vizcat-slides, echo=FALSE}
slide_url(ds4p_urls, "d06_vizcat") %>%
  include_url(height = "400px")
```

# Act: Star Wars Activity

```{r embed-starwars-activity-slides, echo=FALSE}
slide_url(ds4p_urls, "d06_vizcat", "#17") %>%
  include_url(height = "400px")
```

You can find the materials for the Star Wars activity [here][ae03_starwars]. The compiled version should look something like the following...

```{r embed-starwars-activity, echo=FALSE}
include_url("https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html", height = "400px")
```

### What we learned about visualizing categorical data

```{r embed-starwars-activity-learning, echo=FALSE}
slide_url(ds4p_urls, "d06_vizcat", "#24") %>%
  include_url(height = "400px")
```

```{r courselinks, child="includes/courselinks.md"}
```

