# (PART) Module 03 {-}


```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```


# Welcome to the tidyverse!

This module is designed to introduce you to the key ideas of data wrangling and the grammar of tidyverse. In essence, data wrangling is the process of transforming and mapping data from one "raw" data form into a better-suited format. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_03]. The slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Tidy data][d07_tidy]
  * [Grammar of data wrangling][d08_grammar]
  * [Hands on Data Wrangling][d09_wrangle]
  * [Working with Multiple Data Frames][d10_dfs]
* Suggested Readings
  * All subchapters of this module, including
      * [Introduction to dplyr](#dplyr-intro)
      * [Merges on Github](#merges)
  * R4DS     
      * [Data Wrangling](https://r4ds.had.co.nz/wrangle-intro.html), including
        * [Tidy Data](https://r4ds.had.co.nz/tidy-data.html)
        * [Pipes](https://r4ds.had.co.nz/pipes.html)
* Activities
  * [Hotels!](https://github.com/DataScience4Psych/ae-04-hotels-datawrangling)
* Lab
  * [Nobel Laureates](#lab03)

## Estimated Video Length

<!--https://ytplaylist-len.herokuapp.com/ -->


No of videos : 8

Average length of video : 14 minutes, 11 seconds

Total length of playlist : 1 hour, 53 minutes, 34 seconds



# Tidy data

You can follow along with the slides [here][d07_tidy] if they do not appear below.

```{r, echo=FALSE}
# 03.01
video_url="https://www.youtube.com/watch?v=wdDKr92ZT3s"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d07_tidy")
include_url(var_url, height = "400px")
```

## Data structures in R

```{r, echo=FALSE}
# Data structures in R
video_url="https://www.youtube.com/watch?v=fE2zILoY53Y"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d07_tidy","#11")
include_url(var_url, height = "400px")
```


# Grammar of data wrangling

You can follow along with the slides [here][d08_grammar] if they do not appear below.

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=XLj1sJWIv84"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d08_grammar")
include_url(var_url, height = "400px")
```

## Piping

```{r, echo=FALSE}
# PIPES!
video_url="https://www.youtube.com/watch?v=nOrblReuiiA"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d08_grammar","#19")
include_url(var_url, height = "400px")
```

```{r links, child="admin/md/courselinks.md"}
```