# (PART) Module 03 {-}


```{r include = FALSE}
source("common.R")
# install.packages("devtools")

#if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets

ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```

```{r links, child="links.md"}
```
# Data wrangling

This module is designed to introduce you to the key ideas of data wrangling and the grammar of tidyverse. In essence, data wrangling is the process of transforming and mapping data from one "raw" data form into a better-suited format. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-03].The slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Tidy data][d07_tidy]
  * [Grammar of data wrangling][d08_grammar]
  * [Hands on Data Wrangling][d09_wrangle]
* Suggested Readings
  * [Data Wrangling](https://r4ds.had.co.nz/wrangle-intro.html)
  * [Tidy Data](https://r4ds.had.co.nz/tidy-data.html)  
* Activities
  * [Hotels!](UPDATE)
* Lab
  * [302-nobel-laureates](#lab03)



# Tidy data

```{r, echo=FALSE}
# 03.01
video_url="https://www.youtube.com/watch?v=sT5ELdDBUuU"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d07_tidy]) if they do not appear below.

```{r, echo=FALSE}

var_url=slide_url(ds4p_urls,"d07_tidy")
include_url(var_url, height = "400px")

```

# Grammar of data wrangling

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=dInUWpJZG-c"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d08_grammar]) if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d08_grammar")
include_url(var_url, height = "400px")
```

## Piping

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d08_grammar]) if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d08_grammar","#19")
include_url(var_url, height = "400px")
```


# Hands on Data Wrangling

## Working with a single data frame

```{r, echo=FALSE}

# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d09_wrangle]) if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d09_wrangle")
include_url(var_url, height = "400px")
```



## Activity 04: Hotels!

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d09_wrangle","#18")
include_url(var_url, height = "400px")
```

You can find the materials for the Hotels activity [here](https://github.com/DataScience4Psych/ae-04-hotels-datawrangling). The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-04-hotels-datawrangling/hotels-datawrangling.html"
include_url(var_url, height = "400px")
```

