# (PART) Module 03 {-}


```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
library(knitr)
```

# Grammar of data wrangling

This module is designed to <>. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-03]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Tidy data and data wrangling][u1_d05-data-wrangle]
  * [Joining data from multiple sources][u1_d06-data-join]
  * [Data tidying and reshaping][u1_d07-tidy-data]
<!-- 
* Activities
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
--->  
* Lab
	* [302-nobel-laureates](#lab03)



## Tidy data and data wrangling!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8") %>%
  use_align("center")
```

You can follow along with the slides [here][u1_d05-data-wrangle]) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d05-data-wrangle/u1_d05-data-wrangle.html"
include_url(var_url, height = "400px")
```


## Building plots for various data types!

```{r, echo=FALSE}
# Filler Video
embed_url("https://www.youtube.com/watch?v=nYYJmmcgzr8") %>%
  use_align("center")
```

You can follow along with the slides [here][u1_d06-data-join]) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d06-data-join/u1_d06-data-join.html"
include_url(var_url, height = "400px")
```


<!-- 

## Activity 03: Star Wars!

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html#19"
include_url(var_url, height = "400px")
```

You can find the materials for the Star Wars activity [here](https://github.com/DataScience4Psych/ae-03-starwars-dataviz). The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html"
include_url(var_url, height = "400px")
```
--->
