# (PART) Module 04 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Data Diving with Types

Welcome to the exciting world of data types and data importing! This module is your guide to understanding the basics of these important topics. Get ready to dive deep into the world of data and learn everything you need to know about data types and data importing.

To get started, simply watch the videos and follow along with the notes. The video playlist for this module can be found [here][pl_04] and the slides used in the videos are available in the [slides repo][course_slides]. So, grab a notebook and pen and turn to the next page to start learning! With each video, you'll gain a deeper understanding of data types and data importing, and by the end of this module, you'll be a pro. Let's get started!

## Module Materials

* Slides from Lectures
  * [Data types and recoding][d11_types]
  * [Importing data][d12_import]
* Suggested Readings
  * All subchapters of this module
  * R4DS
    * [Wrangle](https://r4ds.had.co.nz/wrangle-intro.html)
    * [Data Transformation](https://r4ds.had.co.nz/transform.html)
    * [Data Importing](https://r4ds.had.co.nz/data-import.html)
* Activities
  * [Hotels, again!](https://github.com/DataScience4Psych/ae-05-hotels-datatypes)
  * [Data Import](https://github.com/DataScience4Psych/ae-06-nobels-sales-dataimport)
* Lab
  * [Visualizing spatial data](#lab04)
  
## Estimated Video Length

No of videos : 9

Average length of video : 8 minutes, 28 seconds

Total length of playlist : 1 hour, 16 minutes, 15 seconds

# Data types and recoding

You can follow along with the slides [here][d11_types] if they do not appear below.

```{r, echo=FALSE}
try_include_tweet("https://twitter.com/IsabellaGhement/status/1473649548778106889")

try_include_tweet("https://twitter.com/patilindrajeets/status/1476171334400954371")
```

## Why should you care about data types?

```{r, echo=FALSE}
# 04.01
"https://www.youtube.com/watch?v=1pErauWQTQY" %>%
  embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
slide_url(ds4p_urls, "d11_types") %>%
  include_url(height = "400px")
```

## Data types

```{r, echo=FALSE}
#
"https://www.youtube.com/watch?v=d8pBipXpDVU" %>%
  embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
slide_url(ds4p_urls, "d11_types", "#14") %>%
  include_url(height = "400px")
```

```{r links, child="admin/md/courselinks.md"}
```
