# (PART) Module 04 {-}


```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```



# Welcome to Data Diving with Types

This module is designed to dive into data types and data importing. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_04]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Data types and recoding][d11_types]
  * [Importing data][d12_import]
* Suggested Readings
  * All subchapters of this module
  * R4DS     
    * [Wrangle](https://r4ds.had.co.nz/wrangle-intro.html)
    * [Data Transformation](https://r4ds.had.co.nz/transform.html)
    * [Data Importing](https://r4ds.had.co.nz/data-import.html)
* Activities
  * [Hotels, again!](https://github.com/DataScience4Psych/ae-05-hotels-datatypes)
  * [Data Import](https://github.com/DataScience4Psych/ae-06-nobels-sales-dataimport)
* Lab
  * [Visualizing spatial data](#lab04)
  
## Estimated Video Length

No of videos : 9

Average length of video : 8 minutes, 28 seconds

Total length of playlist : 1 hour, 16 minutes, 15 seconds



# Data types and recoding

You can follow along with the slides [here][d11_types] if they do not appear below.

```{r, echo=FALSE}
include_tweet("https://twitter.com/IsabellaGhement/status/1473649548778106889")

include_tweet("https://twitter.com/patilindrajeets/status/1476171334400954371")
```

## Why should you care about data types?

```{r, echo=FALSE}
# 04.01
"https://www.youtube.com/watch?v=1pErauWQTQY"  %>%
embed_url() %>%
  use_align("center")
```



```{r, echo=FALSE}
slide_url(ds4p_urls,"d11_types") %>%
include_url(height = "400px")
```




## Data types!

```{r, echo=FALSE}
#
"https://www.youtube.com/watch?v=d8pBipXpDVU" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}
slide_url(ds4p_urls,"d11_types","#14") %>%
include_url( height = "400px")
```


### Another Hotels Activity


```{r, echo=FALSE}
slide_url(ds4p_urls,"d11_types","#24") %>%
include_url(height = "400px")
```

You can find the materials for the Hotels activity [here](https://github.com/DataScience4Psych/ae-05-hotels-datatypes). The compiled version should look something like the following...

```{r, echo=FALSE}
"https://datascience4psych.github.io/ae-05-hotels-datatypes/type-coercion.html" %>%
include_url(height = "400px")
```


## Special Values


```{r, echo=FALSE}
# Special values
"https://www.youtube.com/watch?v=3YoFhEKvieA"  %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
slide_url(ds4p_urls,"d11_types","#26")  %>%
include_url(height = "400px")
```

## Data classes


```{r, echo=FALSE}
# Data classes
"https://www.youtube.com/watch?v=9yNpsBKyXRE"  %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d11_types","#35")
include_url(var_url, height = "400px")
```


## Working with factors


```{r, echo=FALSE}
# Data factors
video_url="https://www.youtube.com/watch?v=yeEnw9rMJe4"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d11_types","#45")
include_url(var_url, height = "400px")
```


### (An) Another Hotels Activity


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d11_types","#50")
include_url(var_url, height = "400px")
```

You can find the materials for the Hotels activity [here](https://github.com/DataScience4Psych/ae-05-hotels-datatypes). The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-05-hotels-datatypes/hotels-forcats.html"
include_url(var_url, height = "400px")
```

## Working with Dates


```{r, echo=FALSE}
# Working with Dates
video_url="https://www.youtube.com/watch?v=R6cEDlzCBfQ"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d11_types","#52")
include_url(var_url, height = "400px")
```

```{r links, child="admin/md/courselinks.md"}
```
