# (PART) Module 04 {-}


```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")

```

# Data types and Data Transformations

This module is designed to dive into data types, data transformations, and data importing. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-04]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Data types and recoding][d11_types]
  * [Importing data][d12_import]
* Suggested Readings
  * [Wrangle](https://r4ds.had.co.nz/wrangle-intro.html)
  * [Data Transformation](https://r4ds.had.co.nz/transform.html)
  * [Data Importing](https://r4ds.had.co.nz/data-import.html)
  * []
* Activities
<!-- 
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
-->  
* Lab
	* [Visualizing spatial data](#lab04)



## Data types and recoding

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][u1_d08-data-types] if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d08-data-types/u1_d08-data-types.html"
include_url(var_url, height = "400px")
```


