```{r types-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```


## Special Values

You can follow along with the slides [here][d11_types] if you would like to open them full-screen.

```{r embed-special-values-video, echo=FALSE}
# Special values
"https://www.youtube.com/watch?v=3YoFhEKvieA" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-special-values-slides, echo=FALSE}
slide_url(ds4p_urls, "d11_types", "#26") %>%
  include_url(height = "400px")
```

R has a few special values you'll bump into when your calculations don't work as expected.


Try dividing by zero in R and you'll get `Inf` (infinity) - unlike other languages that might crash or throw an error. R is pretty chill about it (to an annoying degree):

```{r divide-by-zero, }
pi / 0 # Returns Inf
```

But if you try something truly undefined like dividing zero by zero, you'll get `NaN` (Not a Number):

```{r divide-zero-by-zero, }
0 / 0 # Returns NaN
```

These special values appear when dealing with unusual or undefined operations in R:

- When dividing by zero: `pi / 0` gives `Inf`
- When performing undefined math: `0 / 0` gives `NaN`
- With contradictory operations: `1/0 - 1/0` gives `NaN`
- With consistent operations: `1/0` + `1/0` gives `Inf`


The most common special value you'll encounter is `NA` - missing data. `NA`s are sneaky because they're "contagious" - almost any calculation involving an `NA` will give you `NA` as the result:

```{r demonstrate-na-contagious, }
mean(c(1, 2, NA, 4)) # Returns NA
```





The cool thing about `NA`s is that they're logically consistent. When you work with them in logical operations:

- `TRUE | NA` is `TRUE` (because "true or anything" is always true)
- `FALSE | NA` is `NA` (because we need to know what `NA` is to determine the result)

It's like `NA` is saying "I don't know what I am, but I'll follow the rules of logic!"



## Data classes

You can follow along with the slides [here][d11_types] if you would like to open them full-screen.

```{r embed-data-classes-video, echo=FALSE}
# Data classes
"https://www.youtube.com/watch?v=9yNpsBKyXRE" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-data-classes-slides, echo=FALSE}
slide_url(ds4p_urls, "d11_types", "#35") %>%
  include_url(height = "400px")
```

Think of R's data classes as Lego sets built from basic building blocks. The basic types (`logical`, `character`, `numeric`) are the individual Lego pieces, but classes are the cool structures you build with them.


Take factors - they look like character strings when you print them, but under the hood they're actually integers with labels:


```{r demonstrate-factor-integers, }
x <- factor(c("BS", "MS", "PhD", "MS"))
x # Looks like text
typeof(x) # But it's stored as integers!
as.integer(x) # See the numbers behind the scenes
```

Or dates - they look like calendar dates when you print them:
```{r create-date-object, }
Y2kday <- as.Date("2000-01-01")
Y2kday # Shows as "2000-01-01"
```

But they're actually just counting days since January 1, 1970:

```{r convert-date-to-integer, }
as.integer(Y2kday) # Days since 1970-01-01
```

Because you're just counting time since a fixed date. This explains why you can do math with dates, like finding out what date is 30 days from now:
```{r add-days-to-date, }
Y2kday + 30 # Adds 30 days
```

Even data frames are secretly just lists where all the elements have the same length:
```{r inspect-dataframe-type, }
df <- data.frame(x = 1:2, y = 3:4)
typeof(df) # "list"
```

Understanding these "secret identities" helps you avoid common pitfalls and work more effectively with your data.


## Working with factors

You can follow along with the slides [here][d11_types] if you would like to open them full-screen.

```{r embed-factors-video, echo=FALSE}
# Data factors
"https://www.youtube.com/watch?v=yeEnw9rMJe4" %>%
  embed_url() %>%
  use_align("center")
```

```{r embed-factors-slides, echo=FALSE}
slide_url(ds4p_urls, "d11_types", "#45") %>%
  include_url(height = "400px")
```

Ever made a bar chart where the categories were in a weird order? That's where factors come to the rescue!

When you make a simple plot like this:

```{r load-cat-lovers-data, include=FALSE}
cat_lovers <- read.csv("./data/cat-lovers.csv")
library(ggplot2)
library(viridis)
```
```{r plot-handedness-alphabetical, }
ggplot(cat_lovers, mapping = aes(x = handedness, fill = handedness)) +
  geom_bar() +
  labs(title = "Cat lovers by handedness") +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma")
```
R automatically converts your text variable to a factor, but it uses alphabetical order by default. That's rarely what you want!
The `forcats` package (part of the tidyverse) gives you superpowers for controlling factor order. Want categories ordered by frequency? Just use `fct_infreq`():

```{r plot-handedness-by-frequency, }
cat_lovers %>%
  mutate(handedness = fct_infreq(handedness)) %>%
  ggplot(mapping = aes(x = handedness, fill = handedness)) +
  geom_bar() +
  labs(title = "Cat lovers by handedness") +
  theme_minimal() +
  scale_fill_viridis_d(option = "plasma")
```


Now your most common category appears first - much more informative!
The slides show an example with months, which is a classic problem. If you don't use factors, your months end up in alphabetical order (April, August, December...) instead of calendar order. Using `fct_relevel`() with `month.name` fixes this:

```r
hotels %>%
  mutate(FOO = fct_relevel(BAR, month.name))
```

So next time your plot looks oddly ordered, remember: there's probably a forcats function that can fix it in one line!

```{r courselinks, child="admin/md/courselinks.md"}
```
