# LEC: Working with Dates

You can follow along with the slides [here][d11_types] if you would like to open them full-screen.

```{r dates-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("metadata/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")
```

```{r dates-embed-video, echo=FALSE}
# Working with Dates
"https://www.youtube.com/watch?v=R6cEDlzCBfQ" %>%
  embed_url() %>%
  use_align("center")
```

```{r dates-show-slides, echo=FALSE}
slide_url(ds4p_urls, "d11_types", "#52") %>%
  include_url(height = "400px")
```


Dates are a special type of data that have their own set of functions and methods for working with them. In R, the `lubridate` package is a popular tool for working with dates and times. It provides a set of functions for parsing, manipulating, and formatting dates and times. For example, you can use the `ymd()` function to parse a date in the format "year-month-day": 
```{r lubridate-example, echo=TRUE}
library(lubridate)
date <- ymd("2024-06-01")
date
```

```{r courselinks, child="includes/courselinks.md"}
```

