# LEC: Importing data

```{r import-setup, include = FALSE}
source("common.R")
```

## Importing data!

```{r import-embed-intro-video, echo=FALSE}
"https://www.youtube.com/watch?v=hnUGwlyqx1M" %>%
  embed_url() %>%
  use_align("center") # 07
```

You can follow along with the slides [here][d12_import] if you would like to open them full-screen.

```{r import-show-import-slides, echo=FALSE}
slide_url(ds4p_urls, "d12_import") %>%
  include_url(height = "400px")
```

## Importing and Variable Types

```{r import-embed-vartypes-video, echo=FALSE}
"https://www.youtube.com/watch?v=g7G8ZTnDSv8" %>%
  embed_url() %>%
  use_align("center") # 08
```

You can follow along with the slides [here][d12_import] if you would like to open them full-screen.

```{r import-show-vartypes-slides, echo=FALSE}
slide_url(ds4p_urls, "d12_import", "#19") %>%
  include_url(height = "400px")
```

### More Activity

```{r import-show-activity-slides, echo=FALSE}
slide_url(ds4p_urls, "d12_import", "#24") %>%
  include_url(height = "400px")
```

You can find the materials for the Nobels and sales activities [here](https://github.com/DataScience4Psych/ae-06-nobels-sales-dataimport).

## Vroom

```{r import-embed-vroom-video, echo=FALSE}
"https://www.youtube.com/watch?v=EmB1yyzRX0c" %>%
  embed_url() %>%
  use_align("center")
# :D
```

You can follow along with the slides [here][#d12_import] if you would like to open them full-screen.

```{r import-show-vroom-slides, echo=FALSE}
slide_url(ds4p_urls, "d12_import", "#32") %>%
  include_url(height = "400px")
```

```{r import-embed-tweet, echo=FALSE}
library(tweetrmd)
try_include_tweet("https://twitter.com/statsepi/status/1323619993263755264")
```


```{r courselinks, child="includes/courselinks.md"}
```
