
# (PART) Module 05 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```
```{r links, child="links.md"}
```

# Tips for effective data visualization

This module is designed to introduce ideas related to effective data visualization. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-05]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Tips for effective data visualization][d13_goodviz]
* Activities
  * [Brexit](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)

* Lab
	* [More Dennys](#lab05) 



```{r, echo=FALSE}
library(tweetrmd)
include_tweet("https://twitter.com/ChelseaParlett/status/1361516386925744129")
```
	


# Good Visualization!

## Designing effective visualizations

```{r, echo=FALSE}
# Designing effective visualizations
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```



You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz")
include_url(var_url, height = "400px")
```



## Principles for effective visualizations!

```{r, echo=FALSE}
# Principles for effective visualizations
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz","#7")
include_url(var_url, height = "400px")
```


```{r, echo=FALSE}
# appropriately choosing a y-axis window
include_tweet("https://twitter.com/bmwiernik/status/1362437077820968960")
```

# Design choices in data visualization

I have currated this collection of external video sources on design choices in data visualization.

## How to spot a misleading graph 

When they’re used well, graphs can help us intuitively grasp complex data. But as visual software has enabled more usage of graphs throughout all media, it has also made them easier to use in a careless or dishonest way — and as it turns out, there are plenty of ways graphs can mislead and outright manipulate. Lea Gaslowitz shares some things to look out for.

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=E91bGT9BjYk"
embed_url(video_url) %>%
  use_align("center")
```

## Data Visualization and Misrepresentation

This animatation was produced by some of my colleagues at Brown.

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=x-rDVXVwW9s"
embed_url(video_url) %>%
  use_align("center")
```
## Vox on How coronavirus charts can mislead us


```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=O-3Mlj3MQ_Q"
embed_url(video_url) %>%
  use_align("center")
```

## Vox on Shut up about the y-axis. It shouldn’t always start at zero.

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=14VYnFhBKcY"
embed_url(video_url) %>%
  use_align("center")
```

