
# (PART) Module 05 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```


# Welcome to Tips for effective data visualization

This module is designed to introduce ideas related to effective data visualization. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_05]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides
  * [Tips for effective data visualization][d13_goodviz]
  * [Deep Diving into ggplot][d13b_moreggplot]
* Activities
  * [Brexit](https://github.com/DataScience4Psych/ae-07-brexit-story-dataviz)
* Lab
  * [More Dennys](#lab05) 
* Suggested Readings
  * [A layered grammar of graphics](http://vita.had.co.nz/papers/layered-grammar.html)
  * All subchapters of this module
  * [r4ds](https://r4ds.had.co.nz/)
    * [Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html)
    * [Data visualisation](https://r4ds.had.co.nz/data-visualisation.html)

## Estimated Video Length

No of videos : 7

Average length of video : 12 minutes, 23 seconds

Total length of playlist : 1 hour, 26 minutes, 46 seconds



```{r, echo=FALSE}
library(tweetrmd)
include_tweet("https://twitter.com/ChelseaParlett/status/1361516386925744129")


include_tweet("https://twitter.com/joshuadarr/status/1466150736283242498")
```
	


# Designing effective visualizations

```{r, echo=FALSE}
# Designing effective visualizations
video_url="https://www.youtube.com/watch?v=H9NHHs5VKZc"
embed_url(video_url) %>%
  use_align("center")
```



You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz")
include_url(var_url, height = "400px")
```



## Principles for effective visualizations!

```{r, echo=FALSE}
# Principles for effective visualizations
video_url="https://www.youtube.com/watch?v=oxLB5rg3prA"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz","#7")
include_url(var_url, height = "400px")
```

```{r, echo=FALSE}
include_tweet("https://twitter.com/Notawful/status/1457441237787676672?s=20")
```



# Deeper Diving into ggplot2!

```{r, echo=FALSE}
# 503
video_url="https://www.youtube.com/watch?v=xLBMJMkwc-U"
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here][d13b_moreggplot] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot")
include_url(var_url, height = "400px")
```

## What are the components of a plot?


```{r, echo=FALSE}
# 504
video_url="https://www.youtube.com/watch?v=Eg6hH-1qX5I"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot","#15")
include_url(var_url, height = "400px")
```


## Stats, Geoms, and Positions


```{r, echo=FALSE}
# Stats, Geoms, and Positions
video_url="https://www.youtube.com/watch?v=SklAl8uyuv4"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot","#27")
include_url(var_url, height = "400px")
```

### Jitter to the rescue!


```{r, echo=FALSE}
include_tweet("https://twitter.com/mijkenijk/status/1365913365755822089")
```

## Scales and Coordinates


```{r, echo=FALSE}
# Scales and Coordinates
video_url="https://www.youtube.com/watch?v=vn8RToV1cc4"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot","#39")
include_url(var_url, height = "400px")
```



## How this all works with Minard


```{r, echo=FALSE}
# Minnard
video_url="https://www.youtube.com/watch?v=9nWaGrEMA54"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot","#57")
include_url(var_url, height = "400px")
```



```{r links, child="admin/md/courselinks.md"}
```