# (PART) Module 05 {-}

```{r goodviz-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Tips for Effective Data Visualization

This module is designed to introduce ideas related to effective data visualization. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_05]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

## Module Materials

* Slides
  * [Tips for effective data visualization][d13_goodviz]
  * [Deep Diving into ggplot][d13b_moreggplot]
* Activities
  * [Brexit](https://github.com/DataScience4Psych/ae-07-brexit-story-dataviz)
* Lab
  * [More Dennys](#lab05)
* Suggested Readings
  * [A layered grammar of graphics](http://vita.had.co.nz/papers/layered-grammar.html)
  * All subchapters of this module
  * [r4ds](https://r4ds.had.co.nz/)
    * [Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html)
    * [Data visualisation](https://r4ds.had.co.nz/data-visualisation.html)

## Estimated Video Length

No of videos: 7

Average length of video: 12 minutes, 23 seconds

Total length of playlist: 1 hour, 26 minutes, 46 seconds

```{r goodviz-include-video, echo=FALSE}
library(tweetrmd)
try_include_tweet("https://twitter.com/ChelseaParlett/status/1361516386925744129")
```

# Designing effective visualizations

```{r goodviz-include-video-designing, echo=FALSE}
# Designing effective visualizations
"https://www.youtube.com/watch?v=H9NHHs5VKZc" %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here][d13_goodviz] if you would like to open them full-screen.

```{r goodviz-include-slides-designing, echo=FALSE}
slide_url(ds4p_urls, "d13_goodviz") %>%
  include_url(height = "400px")
```

## Principles for effective visualizations

```{r goodviz-include-video-principles, echo=FALSE}
# Principles for effective visualizations
"https://www.youtube.com/watch?v=oxLB5rg3prA" %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here][d13_goodviz] if you would like to open them full-screen.

```{r goodviz-include-slides, echo=FALSE}
slide_url(ds4p_urls, "d13_goodviz", "#7") %>%
  include_url(height = "400px")
```

```{r goodviz-include-tweet, echo=FALSE}
try_include_tweet("https://twitter.com/joshuadarr/status/1466150736283242498")

# try(try_include_tweet("https://twitter.com/Notawful/status/1457441237787676672?s=20"))
### I think this one is broken, it doesn't show any visual and it seems like it references one
```


```{r links, child="admin/md/courselinks.md"}
```
