
# (PART) Module 05 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```
```{r links, child="links.md"}
```

# Tips for effective data visualization

This module is designed to introduce ideas related to effective data visualization. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-05]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Tips for effective data visualization][d13_goodviz]
  * [Deep Diving into ggplot][d13b_moreggplot]
* Activities
  * [Brexit](https://github.com/DataScience4Psych/ae-07-brexit-story-dataviz)
* Lab
  * [More Dennys](#lab05) 
* Suggested Readings
  * [A layered grammar of graphics](http://vita.had.co.nz/papers/layered-grammar.html)
  * [r4ds](https://r4ds.had.co.nz/)
    * [Graphics for communication](https://r4ds.had.co.nz/graphics-for-communication.html)
    * [Data visualisation](https://r4ds.had.co.nz/data-visualisation.html)



```{r, echo=FALSE}
library(tweetrmd)
include_tweet("https://twitter.com/ChelseaParlett/status/1361516386925744129")
```
	


# Good Visualization!

## Designing effective visualizations

```{r, echo=FALSE}
# Designing effective visualizations
video_url="https://www.youtube.com/watch?v=H9NHHs5VKZc"
embed_url(video_url) %>%
  use_align("center")
```



You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz")
include_url(var_url, height = "400px")
```



## Principles for effective visualizations!

```{r, echo=FALSE}
# Principles for effective visualizations
video_url="https://www.youtube.com/watch?v=oxLB5rg3prA"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d13_goodviz] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13_goodviz","#7")
include_url(var_url, height = "400px")
```



# Deeper Diving into ggplot2!



```{r, echo=FALSE}
# 503
video_url="https://www.youtube.com/watch?v=xLBMJMkwc-U"
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here][d13b_moreggplot] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d13b_moreggplot")
include_url(var_url, height = "400px")
```
