# (PART) Module 06 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```

# Confounding and Spring Break!

This module introduces ideas related to study design (and has additional optional material on science communication). Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-06]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides]. Please note that you only have to do one of the labs. Pick either.


## Module Materials

* Slides
  * [Scientific studies and confounding][d14_confound]
* Suggested Readings
  * Judea Pearl on [Understanding Simpson's Paradox](https://ftp.cs.ucla.edu/pub/stat_ser/r414-reprint.pdf)
  * R4DS section on [communication](https://r4ds.had.co.nz/communicate-intro.html), including
    * [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)

* Lab (Choose one)
  * [Ugly Charts](#lab06a)
  * [Smokers](#lab06b)

## Video Length

No of videos : 4

Average length of video : 10 minutes, 19 seconds

Total length of playlist : 41 minutes, 17 seconds


# Scientific studies and confounding!

## Scientific studies
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=OHfk4dbzNBc"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound")
include_url(var_url, height = "400px")
```

##  Climate Change: A Conditional Probability Case Study

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=Fv-yucms7lQ"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#17")
include_url(var_url, height = "400px")
```

##  Introducing Simpson's Paradox with a case study
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=My68gFE_usU"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#29")
include_url(var_url, height = "400px")
```

## Revisiting Simpson's Paradox

```{r, echo=FALSE}
# 
video_url="https://www.youtube.com/watch?v=vVq_G-a6K30"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#37")
include_url(var_url, height = "400px")
```
