# (PART) Module 06 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```

# Confounding and Communication

This module introduces related to science communication and study design. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-06]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Scientific studies and confounding][d14_confound]
  * [Communicating data science results effectively][d15_goodtalk]
* Suggested Readings
  * R4DS, including
    * [TBD](https://r4ds.had.co.nz/explore-intro.html)
* Lab (Choose one)
  * [Ugly Charts](#lab06a)
  * [Smokers](#lab06b)

# Scientific studies and confounding!

## Scientific studies
```{r, echo=FALSE}
# Filler Video
video_url="https://www.youtube.com/watch?v=OHfk4dbzNBc"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound")
include_url(var_url, height = "400px")
```

##  Climate Change: A Conditional Probability Case Study

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=Fv-yucms7lQ"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#17")
include_url(var_url, height = "400px")
```

##  Introducing Simpson's Paradox with a case study
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=My68gFE_usU"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#29")
include_url(var_url, height = "400px")
```

## Revisiting Simpson's Paradox

```{r, echo=FALSE}
# 
video_url="https://www.youtube.com/watch?v=vVq_G-a6K30"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d14_confound] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#37")
include_url(var_url, height = "400px")
```


# Communicating data science results effectively!


```{r, echo=FALSE}
include_tweet("https://twitter.com/DaphnaHarel/status/1366792094279495685")
```


```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d15_goodtalk] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d15_goodtalk")
include_url(var_url, height = "400px")
```


