# (PART) Module 06 {-}

```{r confounds-setup, include = FALSE}
source("common.R")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Confounding and Communication! {#mod06}

This module introduces ideas related to study design and science communication. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_06]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides]. Note that you only have to do one of the labs. Pick either (or both if you want...)

## Module Materials

* Slides from Lectures
  * [Scientific studies and confounding][d14_confound]
  * [Communicating data science results effectively][d15_goodtalk]
* Suggested Readings
  * Judea Pearl on [Understanding Simpson's Paradox](https://ftp.cs.ucla.edu/pub/stat_ser/r414-reprint.pdf)
  * R4DS on [communication](https://r4ds.had.co.nz/communicate-intro.html), including
    * [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)
  * All subchapters of this module
* Lab
  * [Ugly Charts](#lab06)

## Video Length

No of videos : 5

Average length of video : 11 minutes, 16 seconds

Total length of playlist : 56 minutes, 20 seconds

# Scientific studies and confounding

```{r confounds-include-tweet-intro, echo=FALSE}
try_include_tweet("https://twitter.com/JkayFlake/status/1475533542302093315")
```

You can follow along with the slides [here][d14_confound] if you would like to open them full-screen.

## Scientific studies

```{r confounds-include-video-studies, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("OHfk4dbzNBc")
```
```{r confounds-include-video-studies-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=OHfk4dbzNBc" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-studies, echo=FALSE}
slide_url(ds4p_urls, "d14_confound") %>%
  include_url(height = "400px")
```

## Climate Change: A Conditional Probability Case Study

```{r confounds-include-video-climate, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("Fv-yucms7lQ")
```
```{r confounds-include-video-climate-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=Fv-yucms7lQ" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-climate, echo=FALSE}
slide_url(ds4p_urls, "d14_confound", "#20") %>%
  include_url(height = "400px")
```

## Introducing Simpson's Paradox with a case study

```{r confounds-include-video-simpson-intro, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("My68gFE_usU")
```
```{r confounds-include-video-simpson-intro-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=My68gFE_usU" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-simpson-intro, echo=FALSE}
slide_url(ds4p_urls, "d14_confound", "#40") %>%
  include_url(height = "400px")
```

## Revisiting Simpson's Paradox

```{r confounds-include-video-simpson-revisit, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("vVq_G-a6K30")
```
```{r confounds-include-video-simpson-revisit-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=vVq_G-a6K30" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-simpson, echo=FALSE}
slide_url(ds4p_urls, "d14_confound", "#60") %>%
  include_url(height = "400px")
```

# Communicating data science results effectively

You can follow along with the slides [here][d15_goodtalk] if you would like to open them full-screen.

```{r confounds-include-tweet-comm, echo=FALSE}
try_include_tweet("https://twitter.com/DaphnaHarel/status/1366792094279495685")
```

```{r confounds-include-video-goodtalk, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("xGP08paoAKI")
```
```{r confounds-include-video-goodtalk-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=xGP08paoAKI" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-goodtalk, echo=FALSE}
slide_url(ds4p_urls, "d15_goodtalk") %>%
  include_url(height = "400px")
```

```{r courselinks, child="admin/md/courselinks.md"}
```
