# (PART) Module 06 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Confounding and Communication! {#mod06}

This module introduces ideas related to study design and science communication. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_06]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides]. Note that you only have to do one of the labs. Pick either (or both if you want...)


## Module Materials

* Slides from Lectures
  * [Scientific studies and confounding][d14_confound]
  * [Communicating data science results effectively][d15_goodtalk]
* Suggested Readings
  * Judea Pearl on [Understanding Simpson's Paradox](https://ftp.cs.ucla.edu/pub/stat_ser/r414-reprint.pdf)
  * R4DS on [communication](https://r4ds.had.co.nz/communicate-intro.html), including
      * [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)
  * All subchapters of this module
* Lab (Choose one or both)
  * [Ugly Charts](#lab06a)
  * [Smokers](#lab06b)

## Video Length

No of videos : 5

Average length of video : 11 minutes, 16 seconds

Total length of playlist : 56 minutes, 20 seconds


# Scientific studies and confounding!
```{r, echo=FALSE}
include_tweet("https://twitter.com/JkayFlake/status/1475533542302093315")
```


You can follow along with the slides [here][d14_confound] if they do not appear below.



## Scientific studies
```{r, echo=FALSE}

"https://www.youtube.com/watch?v=OHfk4dbzNBc" %>%
embed_url() %>%
  use_align("center")
```



```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound")
include_url(var_url, height = "400px")
```

##  Climate Change: A Conditional Probability Case Study

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=Fv-yucms7lQ"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#20")
include_url(var_url, height = "400px")
```

##  Introducing Simpson's Paradox with a case study
```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=My68gFE_usU"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d14_confound","#29")
include_url(var_url, height = "400px")
```

## Revisiting Simpson's Paradox

```{r, echo=FALSE}

"https://www.youtube.com/watch?v=vVq_G-a6K30"  %>%
embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
slide_url(ds4p_urls,"d14_confound","#37")  %>%
include_url(height = "400px")
```


# Communicating data science results effectively!

You can follow along with the slides [here][d15_goodtalk] if they do not appear below.

```{r, echo=FALSE}
include_tweet("https://twitter.com/DaphnaHarel/status/1366792094279495685")
```

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=xGP08paoAKI"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d15_goodtalk")
include_url(var_url, height = "400px")
```

```{r links, child="admin/md/courselinks.md"}
```
