# (PART) Module 06 {-}

```{r confounds-setup, include = FALSE}
source("common.R")
```

# Welcome to Confounding and Communication! {#mod06}

This module introduces ideas related to study design and science communication. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_06]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

## Module Materials

* Slides from Lectures
  * [Scientific studies and confounding][d14_confound]
  * [Communicating data science results effectively][d15_goodtalk]
* Suggested Readings
  * Judea Pearl on [Understanding Simpson's Paradox](https://ftp.cs.ucla.edu/pub/stat_ser/r414-reprint.pdf)
  * R4DS on [communication](https://r4ds.had.co.nz/communicate-intro.html), including
    * [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)
  * All subchapters of this module
* Lab
  * [Ugly Charts](#lab06)

## Video Length

No of videos : 5

Average length of video : 11 minutes, 16 seconds

Total length of playlist : 56 minutes, 20 seconds



```{r courselinks, child="includes/courselinks.md"}
```
