# LEC: Simpson's Paradox {#simpsons_paradox}

```{r simpsons-setup, include = FALSE}
source("common.R")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")
```

```{r simpsons-packages, echo = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(knitr)
library(DT)
```

## Introducing Simpson's Paradox with a case study

You can follow along with the slides [here][d14_confound] if you would like to open them full-screen.

```{r confounds-include-video-simpson-intro, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("My68gFE_usU")
```
```{r confounds-include-video-simpson-intro-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=My68gFE_usU" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-simpson-intro, echo=FALSE}
slide_url(ds4p_urls, "d14_confound", "#42") %>%
  include_url(height = "400px")
```


## Revisiting Simpson's Paradox

```{r confounds-include-video-simpson-revisit, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("vVq_G-a6K30")
```
```{r confounds-include-video-simpson-revisit-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=vVq_G-a6K30" %>%
  embed_url() %>%
  use_align("center")
```

```{r confounds-include-slides-simpson, echo=FALSE}
slide_url(ds4p_urls, "d14_confound", "#62") %>%
  include_url(height = "400px")
```


# ODD: Deeper into Simpson's Paradox

This deep dive explores Simpson's Paradox: a phenomenon where a trend that appears in several groups of data reverses when the groups are combined. It is one of the most striking examples of confounding in action, and understanding it is essential for anyone who works with data. If you take away one thing from the confounding module, let it be this: **always ask whether there is a lurking variable that could reverse your conclusions.**

## What Is Simpson's Paradox?

Imagine you are comparing two hospitals. Hospital A has a higher overall survival rate than Hospital B. Sounds like Hospital A is better, right? But when you break the data down by patient severity --- mild vs. critical cases --- Hospital B actually has a higher survival rate in *both* categories. How is that possible?

The answer is that Hospital B treats a much higher proportion of critical patients. Their overall rate gets dragged down by a harder caseload, even though they outperform Hospital A within each severity group. The **lurking variable** --- patient severity --- reverses the apparent conclusion.

This reversal is called **Simpson's Paradox**, named after Edward Simpson who described it in 1951, though the idea was known earlier. It shows up everywhere: in medical studies, in education data, in legal cases, and yes, in psychology research.


## The UC Berkeley Admissions Case

The most famous real-world example comes from UC Berkeley's graduate admissions in 1973. At first glance, the university appeared to discriminate against women: men had a higher overall admission rate. But when researchers examined each department separately, women were admitted at equal or higher rates in most departments. The paradox arose because women disproportionately applied to more competitive departments.

R has this dataset built in as `UCBAdmissions`. Let's explore it.

```{r simpsons-ucb-data}
# UCBAdmissions is a 3-dimensional array; let's convert to a tidy tibble
ucb <- as_tibble(UCBAdmissions)
ucb
```

### Overall admission rates by gender

```{r simpsons-ucb-overall}
ucb_overall <- ucb %>%
  group_by(Gender, Admit) %>%
  summarise(total = sum(n), .groups = "drop") %>%
  group_by(Gender) %>%
  mutate(prop = total / sum(total)) %>%
  filter(Admit == "Admitted")

ucb_overall
```

```{r simpsons-ucb-overall-plot, fig.width=5, fig.height=3.5, fig.cap="Overall admission rates by gender at UC Berkeley, 1973. Men appear to be admitted at a higher rate."}
ggplot(ucb_overall, aes(x = Gender, y = prop, fill = Gender)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 0.55)) +
  scale_fill_manual(values = c("#E48957", "#56B4E9")) +
  labs(
    title = "Overall Admission Rates by Gender",
    subtitle = "UC Berkeley, 1973",
    x = NULL, y = "Admission Rate"
  ) +
  theme_minimal()
```

It looks like men are admitted at a noticeably higher rate. But let's not stop here.

### Admission rates by department

```{r simpsons-ucb-by-dept}
ucb_dept <- ucb %>%
  group_by(Dept, Gender, Admit) %>%
  summarise(total = sum(n), .groups = "drop") %>%
  group_by(Dept, Gender) %>%
  mutate(prop = total / sum(total)) %>%
  filter(Admit == "Admitted")

ucb_dept
```

```{r simpsons-ucb-dept-plot, fig.width=8, fig.height=5, fig.cap="Admission rates by department and gender. In most departments, the gender gap disappears or reverses."}
ggplot(ucb_dept, aes(x = Gender, y = prop, fill = Gender)) +
  geom_col(width = 0.6, show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_fill_manual(values = c("#E48957", "#56B4E9")) +
  facet_wrap(~Dept, nrow = 1) +
  labs(
    title = "Admission Rates by Department and Gender",
    subtitle = "UC Berkeley, 1973",
    x = NULL, y = "Admission Rate"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 8))
```

Now we can see what's really going on. In departments A and B, women are actually admitted at *higher* rates than men. In most other departments, the rates are roughly similar. The overall difference was driven by **where** people applied, not by **who** they were.

### What explains the reversal?

Women applied in larger numbers to departments with low admission rates (like departments C, D, E, and F), while men applied more heavily to departments with high admission rates (A and B). When you combine the data, the composition effect overwhelms the within-department pattern.

```{r simpsons-ucb-applicants, fig.width=8, fig.height=4, fig.cap="Number of applicants by department and gender. Women applied disproportionately to more competitive departments."}
ucb_applicants <- ucb %>%
  group_by(Dept, Gender) %>%
  summarise(applicants = sum(n), .groups = "drop")

ggplot(ucb_applicants, aes(x = Dept, y = applicants, fill = Gender)) +
  geom_col(position = "dodge") +
  scale_fill_manual(values = c("#E48957", "#56B4E9")) +
  labs(
    title = "Number of Applicants by Department",
    subtitle = "UC Berkeley, 1973",
    x = "Department", y = "Number of Applicants"
  ) +
  theme_minimal()
```

This is the confound in action: the lurking variable (department choice) is associated with *both* gender and admission rate.


## A Simulated Example: Treatment Effectiveness

The Berkeley data is compelling, but it can help to see the paradox constructed from scratch. Let's simulate a medical study where a treatment appears to be less effective overall, but is actually more effective within each patient group.

```{r simpsons-sim-data}
set.seed(42)

# Mild cases: treatment works well, and most mild patients get the treatment
mild <- tibble(
  severity = "Mild",
  group = c(rep("Treatment", 200), rep("Control", 50)),
  recovered = c(rbinom(200, 1, 0.90), rbinom(50, 1, 0.85))
)

# Severe cases: treatment still works better, but fewer severe patients get it
severe <- tibble(
  severity = "Severe",
  group = c(rep("Treatment", 50), rep("Control", 200)),
  recovered = c(rbinom(50, 1, 0.50), rbinom(200, 1, 0.40))
)

patients <- bind_rows(mild, severe)
```

### Overall recovery rates

```{r simpsons-sim-overall}
patients %>%
  group_by(group) %>%
  summarise(
    n = n(),
    recovered = sum(recovered),
    rate = mean(recovered)
  ) %>%
  kable(digits = 3, col.names = c("Group", "N", "Recovered", "Recovery Rate"))
```

The control group has a higher overall recovery rate! Should we conclude that the treatment is harmful?

### Recovery rates by severity

```{r simpsons-sim-by-severity}
patients %>%
  group_by(severity, group) %>%
  summarise(
    n = n(),
    recovered = sum(recovered),
    rate = mean(recovered),
    .groups = "drop"
  ) %>%
  kable(digits = 3, col.names = c("Severity", "Group", "N", "Recovered", "Recovery Rate"))
```

Within *both* severity groups, the treatment group has a higher recovery rate. The paradox arises because the treatment group contains a disproportionate number of mild patients (who recover at high rates regardless), while the control group is loaded with severe patients.

```{r simpsons-sim-plot, fig.width=7, fig.height=4.5, fig.cap="Recovery rates by treatment group, overall and stratified by severity. The trend reverses when the lurking variable is accounted for."}
# Prepare summary data
sim_overall <- patients %>%
  group_by(group) %>%
  summarise(rate = mean(recovered), .groups = "drop") %>%
  mutate(severity = "Overall")

sim_by_severity <- patients %>%
  group_by(severity, group) %>%
  summarise(rate = mean(recovered), .groups = "drop")

sim_combined <- bind_rows(sim_by_severity, sim_overall) %>%
  mutate(severity = factor(severity, levels = c("Mild", "Severe", "Overall")))

ggplot(sim_combined, aes(x = group, y = rate, fill = group)) +
  geom_col(width = 0.5, show.legend = FALSE) +
  scale_y_continuous(labels = scales::percent_format(), limits = c(0, 1)) +
  scale_fill_manual(values = c("#CA235F", "#56B4E9")) +
  facet_wrap(~severity) +
  labs(
    title = "Recovery Rates: Overall vs. Stratified",
    x = NULL, y = "Recovery Rate"
  ) +
  theme_minimal()
```


## Visualizing the Paradox with Scatter Plots

Simpson's Paradox isn't limited to proportions. It can also appear in continuous data, where the direction of a correlation reverses when you account for a grouping variable.

```{r simpsons-continuous-data}
set.seed(123)

# Three groups with positive within-group slopes but negative overall slope
continuous_example <- bind_rows(
  tibble(
    group = "A",
    x = rnorm(50, mean = 2, sd = 0.8),
    y = 0.8 * x + rnorm(50, mean = 8, sd = 0.5)
  ),
  tibble(
    group = "B",
    x = rnorm(50, mean = 5, sd = 0.8),
    y = 0.8 * x + rnorm(50, mean = 4, sd = 0.5)
  ),
  tibble(
    group = "C",
    x = rnorm(50, mean = 8, sd = 0.8),
    y = 0.8 * x + rnorm(50, mean = 0, sd = 0.5)
  )
)
```

```{r simpsons-continuous-plot, fig.width=8, fig.height=4, fig.cap="Simpson's Paradox with continuous data. The overall trend (left) is negative, but the within-group trends (right) are all positive."}
p_overall <- ggplot(continuous_example, aes(x = x, y = y)) +
  geom_point(color = "gray40", alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 1) +
  labs(title = "Ignoring Groups", x = "X", y = "Y") +
  theme_minimal()

p_grouped <- ggplot(continuous_example, aes(x = x, y = y, color = group)) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "lm", se = FALSE, linewidth = 1) +
  scale_color_manual(values = c("#E48957", "#56B4E9", "#CA235F")) +
  labs(title = "Accounting for Groups", x = "X", y = "Y", color = "Group") +
  theme_minimal()

library(gridExtra)
grid.arrange(p_overall, p_grouped, ncol = 2)
```

In the left panel, the overall trend line slopes downward. In the right panel, every group's trend line slopes upward. The negative overall slope is entirely an artifact of the group structure.


## Why This Matters for Psychology

Simpson's Paradox is not just a statistical curiosity --- it has real consequences for research conclusions. Here are a few scenarios where psychologists should be especially vigilant:

1. **Treatment effectiveness studies**: A therapy that works well for both mild and severe cases can appear ineffective overall if it is disproportionately used for severe cases (as in our simulation above).

2. **Educational interventions**: A teaching method might improve outcomes for both high- and low-performing students, but if it's primarily adopted by schools with low-performing students, the aggregate data could suggest the method is harmful.

3. **Cross-cultural comparisons**: A relationship between two variables might hold within every culture studied, but reverse when cultures are pooled together, because cultures differ on a third variable.

4. **Longitudinal data**: Trends over time within individuals can look very different from trends across individuals at a single time point (an instance of the ecological fallacy, which is closely related).

The general lesson: **aggregated data can mask or reverse relationships that exist at a finer level of analysis.** Whenever you see a surprising pattern in grouped data, ask yourself what happens when you break the data down further.


## How to Protect Yourself

Here are some practical strategies for avoiding Simpson's Paradox:

- **Think about confounds before analyzing.** What variables might be associated with both your predictor and your outcome? This is the same logic behind experimental design and statistical control.

- **Stratify or condition on potential confounders.** Look at your data broken down by relevant subgroups before drawing conclusions from aggregated numbers.

- **Visualize at multiple levels.** Make plots that show both the overall pattern and the within-group patterns. Side-by-side comparisons (like the faceted plots above) make reversals immediately obvious.

- **Use regression with appropriate controls.** Including confounding variables in a regression model is a formal way to "adjust" for them, though it comes with its own assumptions.

- **Be skeptical of aggregate statistics.** When someone reports an overall rate or trend, always ask: "Overall for whom? Compared to what? Broken down by what?"


## Common Pitfalls

1. **Assuming the aggregated result is always wrong.** Simpson's Paradox doesn't tell you which level of analysis is "correct" --- that depends on the research question. Sometimes the overall rate is what matters; sometimes the within-group rate is more informative. You need subject-matter knowledge to decide.

2. **Over-stratifying.** If you break data into too many subgroups, you can end up with tiny samples and unreliable estimates. There's a balance between controlling for confounds and maintaining statistical power.

3. **Ignoring the paradox entirely.** Many published studies report only aggregate results. If the data could plausibly be confounded by a grouping variable, the aggregate result could be misleading.


## Further Reading

- Judea Pearl, [Understanding Simpson's Paradox](https://ftp.cs.ucla.edu/pub/stat_ser/r414-reprint.pdf) --- the reading from the module materials, a classic treatment of the paradox from a causal inference perspective
- [Wikipedia: Simpson's Paradox](https://en.wikipedia.org/wiki/Simpson%27s_paradox) --- thorough overview with many real-world examples
- [Bickel, P. J., Hammel, E. A., & O'Connell, J. W. (1975). Sex bias in graduate admissions: Data from Berkeley. *Science*, *187*(4175), 398--404](https://www.science.org/doi/10.1126/science.187.4175.398) ([PDF](https://homepage.stat.uiowa.edu/~mbognar/1030/Bickel-Berkeley.pdf)) --- the original Berkeley admissions paper
- [Kievit, R. A., Frankenhuis, W. E., Waldorp, L. J., & Borsboom, D. (2013). Simpson's paradox in psychological science: A practical guide. *Frontiers in Psychology*, *4*, 513](https://www.frontiersin.org/journals/psychology/articles/10.3389/fpsyg.2013.00513/full) --- directly relevant to psychology research


```{r courselinks, child="includes/courselinks.md"}
```
