# (PART) Module 07 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```

# Science Communication and Web scraping

This module is designed to introduce you to the basic ideas behind web scraping. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-07]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Webscraping][d16_webscraping]
  * [Communicating data science results effectively][d15_goodtalk]
* Suggested Readings
  * All subchapters of this module
  * R4DS,
    * Section on [communication](https://r4ds.had.co.nz/communicate-intro.html), including
      * [Graphics for Communication](https://r4ds.had.co.nz/graphics-for-communication.html)
  * [Wrap Scraping from An Introduction to Statistical Programming Methods with R](https://smac-group.github.io/ds/section-web-scraping.html)
* Activities
  * [Scraping IMDB](https://github.com/DataScience4Psych/ae-08-imdb-webscraping)
* Lab
  * [0702_uoe_art]


# Scraping the web

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=iy11CqHnfcM"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d16_webscraping] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d16_webscraping")
include_url(var_url, height = "400px")
```


## Using the SelectorGadget!

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=B7tGbfDgx3c"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d16_webscraping] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d16_webscraping","#12")
include_url(var_url, height = "400px")
```


## Top 250 movies on IMDB!

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=c0s1Im0Q7rA"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d16_webscraping] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d16_webscraping","#25")
include_url(var_url, height = "400px")
```



## Activity 08: IMDB!

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d16_webscraping","#28")
include_url(var_url, height = "400px")
```

You can find the materials for the IMDB activity [here](https://github.com/DataScience4Psych/ae-08-imdb-webscraping).



# Communicating data science results effectively!


```{r, echo=FALSE}
include_tweet("https://twitter.com/DaphnaHarel/status/1366792094279495685")

```

```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=xGP08paoAKI"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d15_goodtalk] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d15_goodtalk")
include_url(var_url, height = "400px")
```

