# (PART) Module 08 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv",sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to web scraping

This module is designed to introduce you to the basic ideas behind web scraping. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_07]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides]. In the original filming, this unit was module 07.


## Module Materials

* Slides from Lectures
  * [Webscraping][d16_webscraping]
* Suggested Readings
  * All subchapters of this module
  * R4DS,
    * Section on [functions](https://r4ds.had.co.nz/functions.html)
  * [Wrap Scraping from An Introduction to Statistical Programming Methods with R](https://smac-group.github.io/ds/section-web-scraping.html)
* Activities
  * [Scraping IMDB](https://github.com/DataScience4Psych/ae-08-imdb-webscraping)
* Lab
  * [lab07]

## Estimated Video Length

No of videos : 3

Average length of video : 14 minutes, 40 seconds

Total length of playlist : 44 minutes

# Scraping the web

You can follow along with the slides [here][d16_webscraping] if they do not appear below.

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=iy11CqHnfcM" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}
slide_url(ds4p_urls,"d16_webscraping") %>%
include_url(height = "400px")
```


## Using the SelectorGadget!

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=B7tGbfDgx3c" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}
slide_url(ds4p_urls,"d16_webscraping","#12") %>%
include_url(height = "400px")
```


## Top 250 movies on IMDB!

```{r, echo=FALSE}
"https://www.youtube.com/watch?v=c0s1Im0Q7rA" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}
slide_url(ds4p_urls,"d16_webscraping","#25") %>%
include_url( height = "400px")
```



## Activity 08: IMDB!

```{r, echo=FALSE}
slide_url(ds4p_urls,"d16_webscraping","#28") %>%
include_url(height = "400px")
```

```{r, echo=FALSE}
slide_url(ds4p_urls,"d16_webscraping","#68") %>%
include_url(height = "400px")
```


You can find the materials for the IMDB activity [here](https://github.com/DataScience4Psych/ae-08-imdb-webscraping).


## Useful RegEx things...
```{r, echo=FALSE}
library(tweetrmd)

include_tweet("https://twitter.com/patilindrajeets/status/1463508045103869963")
```

```{r links, child="admin/md/courselinks.md"}
```