# (PART) Module 08 {-}

```{r welcomefunctions-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Functions and Automation

This module is designed to introduce functions. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_08]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

## Module Materials

* Slides from Lectures
  * [Functions and Automation](https://datascience4psych.github.io/slides/d17_functions/d17_functions.html)
* Activities
  * [Functions and Art!](https://github.com/DataScience4Psych/ae-09-uoeart-functions)
* Suggested Readings
  * All subchapters of this module, including 'Notes on Functions'
  * [r4ds](https://r4ds.had.co.nz/)
    * Sections on [functions](https://r4ds.had.co.nz/functions.html), and
    * [Iterations](https://r4ds.had.co.nz/iteration.html)
* Lab
 	* [Conveying the right message through visualization][lab08]
 	* [Optional Lab Money in Politics][lab08b]



```{r courselinks, child="admin/md/courselinks.md"}
```
