# (PART) Module 08 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```
```{r links, child="links.md"}
```


# Functions Overview

This module is designed to introduce functions. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-08]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides for lectures
  * [Functions and Automation](https://datascience4psych.github.io/slides/d17_functions/d17_functions.html)

* Activities
  * [Functions and Art!](https://github.com/DataScience4Psych/ae-09-uoeart-functions)
* Suggested Readings
  * All subchapters of this module
  * [r4ds](https://r4ds.had.co.nz/)
    * Sections on [functions](https://r4ds.had.co.nz/functions.html), and
    * [Iterations](https://r4ds.had.co.nz/iteration.html)
* Lab
	* [Conveying the right message through visualization][lab08]
	* [Optional Lab Money in Politics][lab08b]

# Functions!

```{r, echo=FALSE}
# Extra Scaffolding
video_url="https://www.youtube.com/watch?v=gfmVLvgztSw"
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here][d17_functions] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions")
include_url(var_url, height = "400px")
```

## Code Along pt 1

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#7")
include_url(var_url, height = "400px")
```

You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-09-uoeart-functions). 

## Functions for real!

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=w0YF_KZpBQM"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d17_functions] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#14")
include_url(var_url, height = "400px")
```

### Code Along pt 2

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#26")
include_url(var_url, height = "400px")
```

You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-09-uoeart-functions). 

## Writing Functions

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=dlmkgbB6Pjo"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d17_functions] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#32")
include_url(var_url, height = "400px")
```


# Automation


```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=BZrm_tlvCq4"
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here][d17_functions] if they do not appear below.

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#42")
include_url(var_url, height = "400px")
```


## Code Along pt 3

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions","#54")
include_url(var_url, height = "400px")
```

You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-09-uoeart-functions). 

