# (PART) Module 08 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
```

# Functions

This module is designed to <>. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-02]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
<!-- 
  * [Tidy data and data wrangling](https://datascience4psych.github.io/slides/u1_d05-data-wrangle/u1_d05-data-wrangle.html)
  * [Joining data from multiple sources](https://datascience4psych.github.io/slides/u1_d06-data-join/u1_d06-data-join.html)
-->  
* Activities
<!-- 
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
-->  
* Lab
<!-- 
	* Pulling yourself up by your bootstraps  
-->  

## Topic 1!

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d05-data-wrangle/u1_d05-data-wrangle.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d05-data-wrangle/u1_d05-data-wrangle.html"
include_url(var_url, height = "400px")
```


## Topic 2!

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/u1_d06-data-join/u1_d06-data-join.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d06-data-join/u1_d06-data-join.html"
include_url(var_url, height = "400px")
```


<!-- 

## Activity 03: Star Wars!

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/u1_d04-data-viz-2/u1_d04-data-viz-2.html#19"
include_url(var_url, height = "400px")
```

You can find the materials for the Star Wars activity [here](https://github.com/DataScience4Psych/ae-03-starwars-dataviz). The compiled version should look something like the following...

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/ae-03-starwars-dataviz/starwars.html"
include_url(var_url, height = "400px")
```
-->