# LEC: Functions

```{r functions-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("metadata/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("metadata/ds4p_urls.csv")
```

```{r functions-include-tweet, echo=FALSE}
try_include_tweet("https://twitter.com/MissyElliott/status/1386707191617527812")
```

You can follow along with the slides [here][d17_functions] if you would like to open them full-screen.

```{r functions-include-video-intro, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("gfmVLvgztSw")
```

```{r functions-include-video-intro-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=gfmVLvgztSw" %>%
  embed_url() %>%
  use_align("center")

# Extra Scaffolding
```

```{r functions-include-slides-intro, echo=FALSE}
slide_url(ds4p_urls, "d17_functions") %>%
  include_url(height = "400px")
```

## Code Along pt 1

```{r functions-include-slides-codealong1, echo=FALSE}
slide_url(ds4p_urls, "d17_functions", "#7") %>%
  include_url(height = "400px")
```


You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-09-uoeart-functions). This time we will be working with .R files, which offer a straightforward method for executing R scripts. Unlike `.Rmd` (R Markdown) files, which integrate R code with Markdown for a blend of code, output, and narrative, `.R` files are focused on script execution, making them simpler for direct script-based activities. Both methods have their uses, depending on your own goals.

## Functions for real


```{r functions-include-video-forreal, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("w0YF_KZpBQM")
```
```{r functions-include-video-forreal-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=w0YF_KZpBQM" %>%
  embed_url() %>%
  use_align("center")
```

```{r functions-include-slides-forreal, echo=FALSE}
slide_url(ds4p_urls, "d17_functions", "#14") %>%
  include_url(height = "400px")
```

## Code Along pt 2

```{r functions-include-slides-codealong2, echo=FALSE}
slide_url(ds4p_urls, "d17_functions", "#26") %>%
  include_url(height = "400px")
```

You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-09-uoeart-functions).

## Writing Functions

```{r functions-include-video-writing, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("dlmkgbB6Pjo")
```
```{r functions-include-video-writing-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=dlmkgbB6Pjo" %>%
  embed_url() %>%
  use_align("center")
```

```{r functions-include-slides-writing, echo=FALSE}
slide_url(ds4p_urls, "d17_functions", "#32") %>%
  include_url(height = "400px")
```


```{r courselinks, child="includes/courselinks.md"}
```
