# (PART) Module 09 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Data and Ethics

This module is a bit different than the typical module. We'll be introducing ideas related to privacy and ethics in the context of data science. There is only one video this week. This week is dedicated to your mid-semester check-in with Mason. I encourage you to watch the video and work your way through the notes. You can find the video playlist for this module [here][pl_09]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Ethics](https://datascience4psych.github.io/slides/d18_ethics/d18_ethics.html)
  * [Algorithmic bias](https://datascience4psych.github.io/slides/d19_bias/d19_bias.html) 
* Readings
	* [How to make a racist AI without really trying](https://blog.conceptnet.io/posts/2017/how-to-make-a-racist-ai-without-really-trying/)
	* [How to write a racist AI in R without really trying](https://notstatschat.rbind.io/2018/09/27/how-to-write-a-racist-ai-in-r-without-really-trying/)
	* Check out the annotated bibliography.
* Activities
  * Annotated Bibliography
* Lab
  * [Lab](#lab09)



```{r, echo=FALSE}
try(include_tweet("https://twitter.com/ChelseaParlett/status/1377361474952200194"))
```

# Data Science and Ethics

## Module Commentary

```{r, echo=FALSE}
# 
"https://www.youtube.com/watch?v=ViXI50uaPOM" %>%
embed_url() %>%
  use_align("center")
```


## Misrepresenting Data


```{r, echo=FALSE}
# 
"https://www.youtube.com/watch?v=31puixbm9n0" %>%
embed_url() %>%
  use_align("center")
```


```{r, echo=FALSE}
slide_url(ds4p_urls,"d18_ethics") %>%
include_url(height = "400px")
```

### Graph Detective 

Source: Lucy D'Agostino McGowan. [Graph detective](https://livefreeordichotomize.com/2020/05/17/graph-detective/). Live Free or Dichotomize. 17 May 2020.
<center>
<iframe width="900" height="1000" src="https://livefreeordichotomize.com/2020/05/17/graph-detective/" frameborder="0"></iframe></center>

## Maps

```{r, echo=FALSE}
slide_url(ds4p_urls,"d18_ethics","#14") %>%
include_url(height = "400px")
```


