# (PART) Module 09 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
```

# Welcome to Data and Ethics

This module is a bit different than the typical module. We'll be introducing ideas related to privacy and ethics in the context of data science. There is only one video this week. This week is dedicated to your mid-semester check-in with Mason. I encourage you to watch the video and work your way through the notes. You can find the video playlist for this module [here][pl_09]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Ethics](https://datascience4psych.github.io/slides/d18_ethics/d18_ethics.html)
  * [Algorithmic bias](https://datascience4psych.github.io/slides/d19_bias/d19_bias.html) 
* Readings
	* [How to make a racist AI without really trying](https://blog.conceptnet.io/posts/2017/how-to-make-a-racist-ai-without-really-trying/)
	* [How to write a racist AI in R without really trying](https://notstatschat.rbind.io/2018/09/27/how-to-write-a-racist-ai-in-r-without-really-trying/)
	* Check out the annotated bibliography.
* Activities
  * Annotated Bibliography
* Lab
  * [Lab](#lab09)




```{r, echo=FALSE}
include_tweet("https://twitter.com/ChelseaParlett/status/1377361474952200194")
```

# Data Science and Ethics

## Module Commentary

```{r, echo=FALSE}
# 
video_url="https://www.youtube.com/watch?v=ViXI50uaPOM"
embed_url(video_url) %>%
  use_align("center")
```

Bonus slides  [here](https://datascience4psych.github.io/slides/d18_ethics/d18_ethics.html) and [here](https://datascience4psych.github.io/slides/d19_bias/d19_bias.html)  if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d18_ethics/d18_ethics.html"
include_url(var_url, height = "400px")
```

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d19_bias/d19_bias.html"
include_url(var_url, height = "400px")
```



## Curated Videography

### Data Science Ethics in 6 Minutes

```{r, echo=FALSE}
# Register, Yim, 2020, ‘Data Science Ethics in 6 Minutes’, YouTube, 29 December, https://youtu.be/mA4gypAiRYU.
embed_url("https://www.youtube.com/watch?v=mA4gypAiRYU") %>%
  use_align("center")
```

### AI for Good in the R and Python ecosystems
```{r, echo=FALSE}
# Keynote "AI for Good in the R and Python ecosystems" by Julien Cornebise
embed_url("https://www.youtube.com/watch?v=fgf2VjnhpCs?start=1162") %>%
  use_align("center")
```

### Are We Automating Racism?
```{r, echo=FALSE}
# Are We Automating Racism?
embed_url("https://www.youtube.com/watch?v=Ok5sKLXqynQ") %>%
  use_align("center")
```

### Big Tech's B.S. about AI ethics

```{r, echo=FALSE}
include_tweet("https://twitter.com/_KarenHao/status/1381981018295578629")

include_tweet("https://twitter.com/JAldrichPL/status/1451524487275552775")
```

## Annotated Bibliography Instructions

An annotated bibliography is a list of citations but with commentary! Ok, more like just an enhanced list where you summarize the source and explain why it is important to include.

Your mission is to add either a citation or an annotation (or both) to this list of Data Science and Ethics Readings. 

1. Carole Cadwalladr and Emma Graham-Harrison. [How Cambridge Analytica turned Facebook ‘likes’ into a lucrative political tool](https://www.theguardian.com/technology/2018/mar/17/facebook-cambridge-analytica-kogan-data-algorithm). The Guardian. 17 March 2018.
2. Chen Wenhong and Anabel Quan-Haase. [Big Data Ethics and Politics: Toward New Understandings](https://journals.sagepub.com/doi/full/10.1177/0894439318810734). Social Science Computer Review. 14 November 2018.
3. Nitasha Tiku. [Google hired Timnit Gebru to be an outspoken critic of unethical AI. Then she was fired for it.]. (https://www.washingtonpost.com/technology/2020/12/23/google-timnit-gebru-ai-ethics/). The Washington Post. 23 December 2020. 
4. Dan Swinhoe. [The biggest data breach fines, penalties, and settlements so far]. (https://www.csoonline.com/article/3410278/the-biggest-data-breach-fines-penalties-and-settlements-so-far.html). CSO.  5 March 2021.
5. Sara Morrison.[Why you should care about data privacy even if you have “nothing to hide”](https://www.vox.com/recode/22250897/facebook-data-privacy-collection-algorithms-extremism). Vox. Jan 28 2021.
6. Richard Van Noorden. [The ethical questions that haunt facial-recognition research](https://www.nature.com/articles/d41586-020-03187-3). Nature. 18 November 2020.
7. Karen Hao. [Big Tech's guide to talking about AI ethics](https://www.technologyreview.com/2021/04/13/1022568/big-tech-ai-ethics-guide/). MIT Technology Review. April 13 2021
8. Catherine D'Ignazio and Lauren F. Klein. [Data feminism](https://mitpressonpubpub.mitpress.mit.edu/data-feminism). Mit Press, 2020.


```{r links, child="admin/md/courselinks.md"}
```
