# (PART) Module 09 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
```

# Welcome to Data and Ethics

This module is a bit different than the typical module. We'll be introducing ideas related to privacy and ethics in the context of data science. There is only one video this week. This week is dedicated to your midsemester check-in with Mason. I encourage you to watch the video and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-09]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
<!-- 
  * [Tidy data and data wrangling](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
  * [Joining data from multiple sources](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
-->  
* Activities
<!-- 
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
-->  
* Lab
	* [Lab](#lab09)



## Module Commentary

```{r, echo=FALSE}
# 
video_url="https://www.youtube.com/watch?v=ViXI50uaPOM"
embed_url(video_url) %>%
  use_align("center")
```



## Data Science Ethics in 6 Minutes



```{r, echo=FALSE}
# Register, Yim, 2020, ‘Data Science Ethics in 6 Minutes’, YouTube, 29 December, https://youtu.be/mA4gypAiRYU.
embed_url("https://www.youtube.com/watch?v=mA4gypAiRYU") %>%
  use_align("center")
```

