# (PART) Module 09 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to Data and Ethics

This module is a bit different than the typical module. We'll be introducing ideas related to privacy and ethics in the context of data science. There are fewer Mason-created videos this week. Instead, this week is dedicated to your mid-semester check-in with Mason. The video playlist for this module can be accessed [here][pl_09]. The slides used in the videos are available in the [slides repo][course_slides].

## Module Materials

* Slides from Lectures
  * [Ethics](https://datascience4psych.github.io/slides/d18_ethics/d18_ethics.html)
  * [Algorithmic bias](https://datascience4psych.github.io/slides/d19_bias/d19_bias.html)
* Readings
  * [How to make a racist AI without really trying](https://blog.conceptnet.io/posts/2017/how-to-make-a-racist-ai-without-really-trying/)
  * [How to write a racist AI in R without really trying](https://notstatschat.rbind.io/2018/09/27/how-to-write-a-racist-ai-in-r-without-really-trying/)
  * Check out the annotated bibliography.
* Activities
  * Annotated Bibliography
* Lab
  * [Lab](#lab09)

## Estimated Reading Time

```{r echo=FALSE}
bytes <- file.size("0901_ethics.Rmd")
words <- bytes / 10
minutes <- words / 200
```

Reading time: `r round(minutes)` minute(s) @ 200 WPM

```{r, echo=FALSE}
try_include_tweet("https://twitter.com/ChelseaParlett/status/1377361474952200194")
```

# Data Science and Ethics

## Module Commentary

```{r, echo=FALSE}
#
"https://www.youtube.com/watch?v=ViXI50uaPOM" %>%
  embed_url() %>%
  use_align("center")
```

## Misrepresenting Data

```{r, echo=FALSE}
#
"https://www.youtube.com/watch?v=31puixbm9n0" %>%
  embed_url() %>%
  use_align("center")
```

```{r, echo=FALSE}
slide_url(ds4p_urls, "d18_ethics") %>%
  include_url(height = "400px")
```
<!--
### Graph Detective 

Source: Lucy D'Agostino McGowan. [Graph detective](https://livefreeordichotomize.com/2020/05/17/graph-detective/). Live Free or Dichotomize. 17 May 2020.
<center>
<iframe width="900" height="1000" src="https://livefreeordichotomize.com/2020/05/17/graph-detective/" frameborder="0"></iframe></center>

-->
## Maps

```{r, echo=FALSE}
slide_url(ds4p_urls, "d18_ethics", "#14") %>%
  include_url(height = "400px")
```
