# Society and AI

```{r include = FALSE}
source("common.R")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

Artificial intelligence (AI) is one of the most transformative technologies of our time, with the potential to revolutionize the way we live, work, and interact with each other. I'll *eventually* have a lecture where I'll explore the ethical and social implications that arise as a result. I will examine how AI is enhancing productivity, improving healthcare, and providing personalized education, while also exploring its potential to displace jobs, perpetuate bias and discrimination, and threaten privacy and security. Furthermore, I will delve into the ethical concerns that arise with the development and deployment of AI, such as transparency, accountability, fairness, and human values. Finally, I will consider the future of AI and its potential outcomes, as well as the role of society in shaping its development and deployment. This lecture aims to provide an in-depth understanding of the complex relationship between AI and society and to encourage responsible and ethical practices in its development and deployment.

In the meantime, this curated videography contains a broad overview of the current landscape...

## Curated Videography

### Last Week Tonight with John Oliver


#### AI Images
```{r, echo=FALSE}
# AI Images: Last Week Tonight with John Oliver
"https://www.youtube.com/watch?v=3YNku5FKWjw" %>%
  embed_url() %>%
  use_align("center")
```


#### Artificial Intelligence
```{r, echo=FALSE}
# Artificial Intelligence: Last Week Tonight with John Oliver
"https://www.youtube.com/watch?v=Sqa8Zo2XWc4" %>%
  embed_url() %>%
  use_align("center")
```
