# (PART) Module 10 {.unnumbered}

```{r language-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
```

# Welcome to modeling the tidy way!

This module introduces you to the ideas behind fitting and interpreting models within the framework of tidymodels. Please watch the videos and work your way through the notes. **The videos start on the next page.** You can find the video playlist for this module [here][pl_10]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides]. In the spirit of transparency, I have strong [non-positive][tidymodelthoughts] thoughts on tidymodels, but I also think that it is useful for you to learn them. I have similar feelings about the necessary evil that is ANOVA or SPSS.

## Module Materials

- Videos
  - Located in the subchapters of this module
- Slidedecks
  - [Language of Models](https://datascience4psych.github.io/slides/d20_language/d20_language.html)
  - [Fitting and interpreting models](https://datascience4psych.github.io/slides/d21_fitting/d21_fitting.html)
  - [Modeling non-linear relationships](https://datascience4psych.github.io/slides/d22_nonlinear/d22_nonlinear.html)
  - [Models with multiple predictors](https://datascience4psych.github.io/slides/d23_multiple/d23_multiple.html)

- Suggested Readings
  - All subchapters of this module
  - Articles
    - [Rodgers, J. L. (2010). The epistemology of mathematical and statistical modeling: A quiet methodological revolution. American Psychologist, 65, 1-12.](https://core.ac.uk/reader/191832935)
  - R4DS
    - [Model Section](https://r4ds.had.co.nz/model-intro.html), including
      - [Model Basics](https://r4ds.had.co.nz/model-basics.html)
      - [Model Building](https://r4ds.had.co.nz/model-building.html)

- Lab
  - Lab 10

# Language of Models

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d20_language/d20_language.html) if you would like to open them full-screen.

## What is a model?

```{r language-include-video, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("XbyXxGDFmFQ")
```
```{r language-include-video, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=XbyXxGDFmFQ" %>%
  embed_url() %>%
  use_align("center")
```

```{r language-include-video, echo=FALSE}
"https://datascience4psych.github.io/slides/d20_language/d20_language.html#1" %>%
  include_url(height = "400px")
```

## Modeling the relationship between variables

```{r language-include-video, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("Y0ehX5VI_Wo")
```
```{r language-include-video, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=Y0ehX5VI_Wo" %>%
  embed_url() %>%
  use_align("center")
```

```{r language-include-slides, echo=FALSE}
"https://datascience4psych.github.io/slides/d20_language/d20_language.html#16" %>%
  include_url(height = "400px")
```



```{r links, child="admin/md/courselinks.md"}
```
