# (PART) Module 10 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
```

```{r links, child="links.md"}
```

# Welcome to fitting and interpreting models

This module is designed to introduce you to the ideas behind fitting and interpreting models. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-02]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Language of Models](https://datascience4psych.github.io/slides/d20_language/d20_language.html)

* Suggested Readings
  * All subchapters of this module
  * R4DS
      * [Model Section](https://r4ds.had.co.nz/model-intro.html), including
         * [Model Basics](https://r4ds.had.co.nz/model-basics.html)
         * [Model Building](https://r4ds.had.co.nz/model-building.html)

* Lab
	* [lab-10]

# Language of Models!

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d20_language/d20_language.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d20_language/d20_language.html#1"
include_url(var_url, height = "400px")
```


## What is a model?
```{r, echo=FALSE}
# 
video_url="https://www.youtube.com/watch?v=XbyXxGDFmFQ"
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d20_language/d20_language.html#2"
include_url(var_url, height = "400px")
```


## Modeling the relationship between variables

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=Y0ehX5VI_Wo"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d20_language/d20_language.html#16"
include_url(var_url, height = "400px")
```


