# (PART) Module 11 {-}

```{r cv-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("metadata/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")
```

# Welcome to Overfitting and Cross-Validation

This module is designed to introduce you to cross-validation and overfitting. Please watch the videos and work your way through the notes. **The videos start on the next page.** You can find the video playlist for this module [here][pl_11]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

```{r cv-include-tweet, echo=FALSE, eval=FALSE}
try_include_tweet("https://twitter.com/Chops310/status/1382481152489979913")
```

## Module Materials

* Videos and Slides from Lectures
  * [Overfitting](https://datascience4psych.github.io/slides/d24_overfitting/d24_overfitting.html)
  * [Cross-validation](https://datascience4psych.github.io/slides/d25_crossvalidation/d25_crossvalidation.html)
* Suggested Readings
  * All subchapters of this module
  * Articles
    * [de Rooij, M., & Weeda, W. (2020). Cross-validation: A method every psychologist should know. Advances in Methods and Practices in Psychological Science, 3(2), 248-263.](https://journals.sagepub.com/doi/full/10.1177/2515245919898466)
    * [MacCallum, R. C., Zhang, S., Preacher, K. J., & Rucker, D. D. (2002). On the practice of dichotomization of quantitative variables. Psychological Methods, 7, 19-40.](http://www.quantpsy.org/pubs/maccallum_zhang_preacher_rucker_2002.pdf)
  * R4DS
    * [Many Models](https://r4ds.had.co.nz/many-models.html)

# LEC: Overfitting

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d24_overfitting/d24_overfitting.html) if you would like to open them full-screen. The embedded code for feature enginering can be found [here](#featurenotes)

```{r cv-include-slides-intro, echo=FALSE}
"https://datascience4psych.github.io/slides/d24_overfitting/d24_overfitting.html#1" %>%
  include_url(height = "400px")
```

## Prediction

```{r cv-include-video-prediction, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("U70OmbO-DP4")
```
```{r cv-include-video-prediction-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=U70OmbO-DP4" %>%
  embed_url() %>%
  use_align("center")
```

```{r cv-include-slides-prediction, echo=FALSE}
"https://datascience4psych.github.io/slides/d24_overfitting/d24_overfitting.html#2" %>%
  include_url(height = "400px")
```

## Workflow

```{r cv-include-video-workflow, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("R4h9u-sQHwI")
```
```{r cv-include-video-workflow-html, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=R4h9u-sQHwI" %>%
  embed_url() %>%
  use_align("center")
```

```{r cv-include-slides-workflow, echo=FALSE}
"https://datascience4psych.github.io/slides/d24_overfitting/d24_overfitting.html#17" %>%
  include_url(height = "400px")
```

# LEC: Cross-Validation

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d25_crossvalidation/d25_crossvalidation.html) if you would like to open them full-screen.

```{r cv-include-video-crossval, echo=FALSE}
# CV
"https://www.youtube.com/watch?v=KQ9f8s7RB5g" %>%
  embed_url() %>%
  use_align("center")
```

```{r cv-include-slides-crossval, echo=FALSE}
slide_url(ds4p_urls, "d25_crossvalidation") %>%
  include_url(height = "400px")
```

## V-Fold

```{r cv-include-video-vfold, echo=FALSE}
# v fold
"https://www.youtube.com/watch?v=quEVKV-Tk0Y" %>%
  embed_url() %>%
  use_align("center")
```

```{r cv-include-slides-vfold, echo=FALSE}
slide_url(ds4p_urls, "d25_crossvalidation", "#35") %>%
  include_url(height = "400px")
```


```{r courselinks, child="includes/courselinks.md"}
```
