# (PART) Module 11 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
library(knitr)
```

# Welcome to prediction and overfitting

This module is designed to expand up ideas introduced in the previous module. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-11]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides
  * [Modeling non-linear relationships](https://datascience4psych.github.io/slides/d22_nonlinear/d22_nonlinear.html)
  * [Models with multiple predictors](https://datascience4psych.github.io/slides/d23_multiple/d23_multiple.html)

* Suggested Readings
  * All subchapters of this module
  * R4DS
  
* Lab 10

# Modeling non-linear relationships

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d22_nonlinear/d22_nonlinear.html) if they do not appear below.

```{r, echo=FALSE}
video_url="https://www.youtube.com/watch?v=tSzwRyO7cNw"
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d22_nonlinear/d22_nonlinear.html#1"
include_url(var_url, height = "400px")
```




# Modeling with multiple predictors

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d23_multiple/d23_multiple.html) if they do not appear below.


## The linear model with multiple predictors

```{r, echo=FALSE}

embed_url("https://www.youtube.com/watch?v=xTv1wqNKbHI") %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d23_multiple/d23_multiple.html#2"
include_url(var_url, height = "400px")
```


## Two numerical predictors


```{r, echo=FALSE}

video_url="https://www.youtube.com/watch?v=SVoazCcAZ"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d23_multiple/d23_multiple.html#24"
include_url(var_url, height = "400px")
```

