# Lab: Cross Validation in Action {#lab11}

## Snappy Name {.unnumbered}

```{r include=FALSE}
library(tufte)
library(knitr)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE, # for regression output
  digits = 2
)
knitr::opts_chunk$set(eval = FALSE)
```



## Getting started {.unnumbered}

### Packages {.unnumbered}

In this lab we will work with the **tidyverse**, **openintro**, and **broom** packages.

```{r eval = FALSE}
library(tidyverse)
library(broom)
library(openintro)
```

<!-- 
### Housekeeping

#### Git configuration / password caching

Configure your Git user name and email. If you cannot remember the instructions, refer to an earlier lab. Also remember that you can cache your password for a limited amount of time.

#### Project name

Update the name of your project to match the lab's title.

-->

## Warm up {.unnumbered}

Before we introduce the data, let's warm up with some simple exercises.

### YAML {.unnumbered}

Open the R Markdown (Rmd) file in your project, change the author name to your name, and knit the document.

### Commiting and pushing changes {.unnumbered}

- Go to the **Git** pane in your RStudio.
- View the **Diff** and confirm that you are happy with the changes.
- Add a commit message like "Update team name" in the **Commit message** box and hit **Commit**.
- Click on **Push**. This will prompt a dialogue box where you first need to enter your user name, and then your password.

## The data {.unnumbered}

We will use the `ames` dataset from the **openintro** package. This dataset contains information about house sales in Ames, Iowa. We will use cross-validation to estimate the test error of a linear regression model predicting sale price from square footage.

## Exercises {.unnumbered}

1. Fit a linear regression model predicting `sale_price` from `sqft_living`. Use the `lm()` function.

2. Use 5-fold cross-validation to estimate the test error of this model. You can use the `sample()` function to create the folds.

3. Report the estimated test error (mean squared error) from the cross-validation.

```{r links, child="admin/md/courselinks.md"}
```
