# (PART) Module 12 {-}

```{r uncertainty-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
```

# Welcome to Quantifying Uncertainty

This module is designed to introduce ideas related to representing uncertainty. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_13]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

## Module Materials

* Slides from Lectures
  * [Quantifying Uncertainty](https://datascience4psych.github.io/slides/d26_quantify/d26_quantify.html)
  * [Bootstrapping](https://datascience4psych.github.io/slides/d27_bootstrap/d27_bootstrap.html)
* Suggested Readings
  * All subchapters of this module
  * About the Needle
     * [The NYT's election forecast needle is stressing people out with fake jitter](https://www.theverge.com/2016/11/8/13571216/new-york-times-election-forecast-jitter-needle)
   	* [Kopf, D. (2020, November 3). Why it's okay to look at the New York Times "election needle". Quartz](https://qz.com/1924611/its-okay-to-look-at-the-new-york-times-election-needle/)
* Lab
  * [Lab](#lab12)

# Quantifying Uncertainty

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d26_quantify/d26_quantify.html) if you would like to open them full-screen.  

```{r uncertainty-include-1, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("rQp3knI6Pdk")
```
```{r uncertainty-include-2, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=rQp3knI6Pdk" %>%
  embed_url() %>%
  use_align("center")
```

```{r uncertainty-include-3, echo=FALSE}
"https://datascience4psych.github.io/slides/d26_quantify/d26_quantify.html" %>%
  include_url(height = "400px")
```

```{r uncertainty-include-4, echo=FALSE}
try_include_tweet("https://twitter.com/jmtfeliciano/status/1597355324008108034")

# https://retractionwatch.com/2022/12/22/that-paper-with-the-t-error-bars-was-just-retracted/
```

```{r uncertainty-include-5, echo=FALSE}
try_include_tweet("https://twitter.com/bmwiernik/status/1451894527187431429")
```

# Bootstrapping

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d27_bootstrap/d27_bootstrap.html) if you would like to open them full-screen.

```{r uncertainty-include-6, echo=FALSE,eval = knitr::is_latex_output()}
embed_youtube_alt("j0cfD9290-s")
```
```{r uncertainty-include-7, echo=FALSE,eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=j0cfD9290-s" %>%
  embed_url() %>%
  use_align("center")
# bootstrap
```

```{r uncertainty-include-8, echo=FALSE}
"https://datascience4psych.github.io/slides/d27_bootstrap/d27_bootstrap.html" %>%
  include_url(height = "400px")
```

```{r links, child="admin/md/courselinks.md"}
```
