# (PART) Module 13 {-}

# Welcome to Base R and Simulating Data

```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) # ... embedding tweets
library(vembedr) # ... embedding youtube videos
library(knitr)
library(tidyverse)
```

This module is designed to introduce you to the key ideas behind simulating data. In essence, we are making our own data instead of using someone else's. Eventually, there will be video lectures. Once those videos exist, please watch the videos and work your way through the notes. **The videos will eventually start on the next page.**  You will eventually be able to find the video playlist for this module [here][pl_13]. The slides used to make the videos in this module will be able to be found in the [slides repo][course_slides].

```{r simulations-include-1, echo=FALSE}
"https://twitter.com/AmandaKMontoya/status/1683928825644871685" %>%
  try_include_tweet()
```

## Module Materials

* Slides from Lectures
* Suggested Readings
  * All subchapters of this module, including
    * ...
  * R4DS
    * [TBD](https://r4ds.had.co.nz/), including
      * [TBD]
      * [TBD]
* Activities
  * [TBD]
* Lab
  * [Simulating on Mars](#lab12)

## Estimated Video Length

<!--https://ytplaylist-len.herokuapp.com/ -->

No of videos : TBD

Average length of video : TBD

Total length of playlist : TBD


# Lecture: Getting started with simulating data in R

```{r simulations-include-2, echo=FALSE}
# Designing effective visualizations
# "https://www.youtube.com/watch?v=H9NHHs5VKZc" %>%
#  embed_url() %>%
#  use_align("center")
#

video_url <- sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here][d30_simulations] if you would like to open them full-screen.

```{r simulations-include-3, echo=FALSE}
slide_url(ds4p_urls, "d30_simulations") %>%
  include_url(height = "400px")
```

<!--
## Creating datasets with quantitative and categorical variables

```{r simulations-include-4, echo=FALSE}
# Filler Video
video_url <- sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```


You can follow along with the slides [here][d30_simulations] if you would like to open them full-screen.

```{r simulations-include-5, echo=FALSE}
# slide_url(ds4p_urls, "d13_goodviz") %>%
"https://datascience4psych.github.io/slides/00_ASU/d00_slide.html#1" %>%
  include_url(height = "400px")
```
-->

```{r links, child="admin/md/courselinks.md"}
```
