# (PART) Module 14 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
library(tweetrmd) #... embedding tweets
```
```{r links, child="links.md"}
```

# Welcome to interactive web apps {#rshiny}

This module is designed to introduce you to interactive web apps. We'll focus on rshiny. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-14] . Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].

```{r, echo=FALSE}
include_tweet("https://twitter.com/ianagutierrez/status/1358253894518927362")
```

## Module Materials

* Slides
<!-- 
  * [Tidy data and data wrangling](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
  * [Joining data from multiple sources](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
-->  

* Lab
  * No Lab!


## Topic 1!

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html"
include_url(var_url, height = "400px")
```


## Topic 2!


You can follow along with the slides [here](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html) if they do not appear below.

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```


```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html"
include_url(var_url, height = "400px")
```


