# (PART) Module 14 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
library(tweetrmd) #... embedding tweets
```

# Welcome to interactive web apps {#rshiny}

This module will introduce you to interactive web apps. We'll focus on rshiny. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_14]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

```{r, echo=FALSE}
include_tweet("https://twitter.com/ianagutierrez/status/1358253894518927362")
```

## Module Materials

* Slides
  * [Interactive web apps](https://datascience4psych.github.io/slides/d28_interactive/d28_interactive.html)
* Videos
  * Curated Videos 
* Readings
  * [Rstudio tutorial](https://shiny.rstudio.com/tutorial/written-tutorial/lesson1/)
  * Written Lecture Notes
* Lab
  * No Lab!


# RShiny Overview

```{r, echo=FALSE}
# Filler Video
#video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
#embed_url(video_url) %>%
#  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d28_interactive/d28_interactive.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d28_interactive/d28_interactive.html"
include_url(var_url, height = "400px")
```


# Practical Advice from the Data Professor

I've outsourced the practical videos because, frankly, I'm a novice when it comes to rshiny. I know enough to be dangerous, but not enough to wield that power for teaching. However, my good twitter friend, [the Data Professor](https://twitter.com/thedataprof) has that power. You should check out his [youtube channel](https://www.youtube.com/channel/UCV8e2g4IWQqK71bbzGDEI4Q)


## Web Apps in R: Building your First Web Application in R

```{r, echo=FALSE}
# Web Apps in R: Building your First Web Application in R | Shiny Tutorial Ep 1
video_url="https://www.youtube.com/watch?v=tfN10IUX9Lo"
embed_url(video_url) %>%
  use_align("center")
```

## Web Apps in R: Build Interactive Histogram Web Application in R
```{r, echo=FALSE}
#  Shiny Tutorial Ep2
video_url="https://www.youtube.com/watch?v=lC1Dk6gUbe0"
embed_url(video_url) %>%
  use_align("center")
```

## Web Apps in R: Building Data-Driven Web Application in R
```{r, echo=FALSE}
#  Shiny Tutorial Ep3
video_url="https://www.youtube.com/watch?v=CYXvVuklWRM"
embed_url(video_url) %>%
  use_align("center")
```

## Web Apps in R: Building the Machine Learning Web Application in R
```{r, echo=FALSE}
#  Shiny Tutorial Ep4
video_url="https://www.youtube.com/watch?v=ceg7MMQNln8"
embed_url(video_url) %>%
  use_align("center")
```

## Web Apps in R: Build BMI Calculator web application in R for health monitoring
```{r, echo=FALSE}
#  Shiny Tutorial Ep5
video_url="https://www.youtube.com/watch?v=9EQ6cwBQpvo"
embed_url(video_url) %>%
  use_align("center")
```


```{r links, child="admin/md/courselinks.md"}
```
