# (PART) Module 15 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
library(tweetrmd) #... embedding tweets
```
```{r links, child="links.md"}
```
# Special Topics: Machine, Learn!

This module is designed to introduce machine learning and neural networks. Please watch the videos and work your way through the notes. You can find the video playlist for this module [here][ds4p-pl-15]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].

```{r, echo=FALSE}
include_tweet("https://twitter.com/kareem_carr/status/1366565047019982856")
```

## Module Materials

* Slides
<!-- 
  * [Tidy data and data wrangling](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
  * [Joining data from multiple sources](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
-->  
* Activities
<!-- 
  * [Star Wars!](https://github.com/DataScience4Psych/ae-03-starwars-dataviz)
-->  
* Readings
	* [applied-ml](https://github.com/eugeneyan/applied-ml)

# Machines Learn!

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html"
include_url(var_url, height = "400px")
```

```{r, echo=FALSE}
include_tweet("https://twitter.com/kareem_carr/status/1351338586063458304")
```

# Neural Networks!

## What is a Neural Network?

```{r, echo=FALSE}
# But what is a Neural Network?
video_url="https://www.youtube.com/watch?v=aircAruvnKk"
embed_url(video_url) %>%
  use_align("center")
```

## More on Neural Networks

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html) if they do not appear below.

```{r, echo=FALSE}
var_url="https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html"
include_url(var_url, height = "400px")
```




```{r, echo=FALSE}
include_tweet("https://twitter.com/FrRonconi/status/1358045974451662854")
```

