# (PART) Module 15 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
library(tweetrmd) #... embedding tweets
```


# Special Topics: Machine, Learn!

This bonus module is designed to introduce machine learning and neural networks. Please watch the curated videos and work your way through the notes. 
You can find the video playlist for this module [here][pl_15]. 
Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].

```{r, echo=FALSE}
include_tweet("https://twitter.com/kareem_carr/status/1366565047019982856")

include_tweet("https://twitter.com/kareem_carr/status/1351338586063458304")

include_tweet("https://twitter.com/minimaxir/status/1474913997807755268")
```

## Module Materials


* Readings
  * [A visual introduction to machine learning](http://www.r2d3.us/visual-intro-to-machine-learning-part-1/)



# Neural Networks!

## What is a Neural Network?

```{r, echo=FALSE}
# But what is a Neural Network?
video_url="https://www.youtube.com/watch?v=aircAruvnKk"
embed_url(video_url) %>%
  use_align("center")
```



```{r, echo=FALSE}
include_tweet("https://twitter.com/FrRonconi/status/1358045974451662854")
```

## How does it learn?

The following videos by Youtuber and game developer [Dani](https://www.youtube.com/c/DaniDev/about) illustrate the iterative learning process. Specifically, I think his video on "[Teaching A.I. to Play my Game!](https://www.youtube.com/watch?v=JxdcuHXn9dc)" introduces the basic ideas behind neural networks well and shows the dynamic developement process. His video on "[Stickman A.I. Learns To Walk
](https://www.youtube.com/watch?v=kowCrRPqRMM)" builds upon these ideas and illustrates how the iterative learning process requires thoughtful interventions.


### Teaching A.I. to Play My Game!

```{r, echo=FALSE}
# Teaching A.I. to Play My Game!
video_url="https://www.youtube.com/watch?v=JxdcuHXn9dc"
embed_url(video_url) %>%
  use_align("center")
```

### Stickman A.I. Learns To Walk


```{r, echo=FALSE}
#  Stickman A.I. Learns To Walk
video_url="https://www.youtube.com/watch?v=kowCrRPqRMM"
embed_url(video_url) %>%
  use_align("center")
```



# Natural Language Processing


```{r, echo=FALSE}
# I asked an AI for video ideas, and they were actually good

video_url="https://www.youtube.com/watch?v=TfVYxnhuEdU"
embed_url(video_url) %>%
  use_align("center")

# I asked an AI for video ideas for other YouTubers. It went badly.
video_url="https://www.youtube.com/watch?v=Fw6GjAPpKuA"
embed_url(video_url) %>%
  use_align("center")
```

Resources:
- https://www.vox.com/future-perfect/2019/2/14/18222270/artificial-intelligence-open-ai-natural-language-processing
- https://app.inferkit.com/demo

```{r links, child="admin/md/courselinks.md"}
```