# (PART) Module 16 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("admin/csv/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./admin/csv/ds4p_urls.csv")
library(tweetrmd) # ... embedding tweets
```

# Special Topics: Reproducible reports

This bonus module is designed to introduce reproducible documents. Please watch the curated videos and work your way through the notes. **The videos start on the next page.**  
You can find the video playlist for this module [here][pl_16].
Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [TBD][TBD]
* Suggested Readings
  * All subchapters of this module, including
    * [TOPIC](#basiccare)
  * R4DS
* Activities
  * [Activities][ae03_starwars]
* Lab
  * [LAB](#lab16)

<!--
# Topic 1
```{r, echo=FALSE}
# Filler Video
sample_no_surprises(ds4p_funyoutube$funyoutube) %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here]() if you would like to open them full-screen.
Resources:
<!--
```{r, echo=FALSE}
slide_url(ds4p_urls, "d18_ethics") %>%
  include_url(height = "400px")
```

-->
-->

```{r links, child="admin/md/courselinks.md"}
```
