# (PART\*) Module Last {-}

# Don't Miss The Last Module {-}

```{r include = FALSE}
source("common.R")
#htmltools::tagList(rmarkdown::html_dependency_font_awesome())
library(tidyverse)
library(devtools)
library(DT)

```

```{r links, child="links.md"}
```


This course was designed to be a starting point. You have learned so much in a short span. I am so proud of each and every one of you!!!!! 

```{r}
#library(embedr)
#embed_audio("assets/audio/SoLongFarewellAllParts.mid")
```




```{r, echo=FALSE}
#library(vembedr)
#embed_url("https://www.youtube.com/watch?v=HGtAP_Td8z4") %>%
#  use_align("center")
```
Before you go -- I have some important practical things to walk you thru. Most of these are Wake Forest Specific... such as making sure to connect your github account to a non-WFU email. As well as a well other things...

## Important Wake Forest Stuff

For those of you who are graduating from Wake Forest University, you may not be aware that your email address gets shut off very soon after you graduate. Why do they do this? I do not know and this it is a silly policy. Regardless, before you lose access to your WFU email, you need to add a 2nd email address to your github account. Otherwise you will not be able to get access to your materials after you graduate.

Github has some incredibly useful guides to do this...

* [How to add an email address to your github account](https://docs.github.com/en/github/setting-up-and-managing-your-github-user-account/adding-an-email-address-to-your-github-account)
* [Hot to change your primary email address](https://docs.github.com/en/github/setting-up-and-managing-your-github-user-account/changing-your-primary-email-address)
* [How to set up a backup email address]
 (https://docs.github.com/en/github/setting-up-and-managing-your-github-user-account/setting-a-backup-email-address)

Please do not procrastinate this! It is really important to do this before you lose access!!!!





