
```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
```

# Data and Visualization: Module 2

This module is designed to <>. Please watch the videos and work your way through the notes. You can find the video playlist [here](https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup). Most of the slides used to make the videos in this module can be found in the [slides repo](https://github.com/DataScience4Psych/Slides).

## Module Roadmap 






### Further resources

The above will get your basic setup ready but here are some links if you are interested in reading a bit further.

* [How to Use RStudio][rstudio-support]
* [RStudio's leads for learning R][rstudio-R-help]
* [R FAQ][cran-faq]
* [R Installation and Administration][cran-R-admin]
* [More about add-on packages in the R Installation and Administration Manual][cran-add-ons]
    


```{r links, child="links.md"}
```
