# (PART) Other Coolness {-}

```{r include = FALSE}
source("common.R")

# install.packages("devtools")

if (!require("tweetrmd"))  devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd) #... embedding tweets
library(vembedr)  #... embeding youtube videos
library(knitr)
```

```{r links, child="links.md"}
```
# Good Resources

https://psychnerdjae.github.io/into-the-tidyverse/


# Media without a home yet

## Are you a student? If yes, this is the best data science project for you!

```{r, echo=FALSE}
# Are you a student? If yes, this is the best data science project for you!
embed_url("https://www.youtube.com/watch?v=2goqyY5XBeI") %>%
  use_align("center")
```



## How computer memory works!

```{r, echo=FALSE}
# How computer memory works - Kanawat Senanan
embed_url("https://www.youtube.com/watch?v=p3q5zWCw8J4") %>%
  use_align("center")
```


## Quantum Computers Explained!

```{r, echo=FALSE}
# Quantum Computers Explained
embed_url("https://www.youtube.com/watch?v=JhHMJCUmq28") %>%
  use_align("center")
```


## The Rise of the Machines – Why Automation is Different this Time

```{r, echo=FALSE}
# Quantum Computers Explained
embed_url("https://www.youtube.com/watch?v=WSKi8HfcxEk") %>%
  use_align("center")
```

## Who Would Be King of America if George Washington had been made a monarch?

```{r, echo=FALSE}
# Who Would Be King of America if George Washington had been made a monarch?
embed_url("https://www.youtube.com/watch?v=ZxnBveop5no") %>%
  use_align("center")
```

## Emergence – How Stupid Things Become Smart Together

```{r, echo=FALSE}
# Emergence – How Stupid Things Become Smart Together

embed_url("https://www.youtube.com/watch?v=16W7c0mb-rE") %>%
  use_align("center")
```


## The Birthday Paradox


```{r, echo=FALSE}
# The Birthday Paradox

embed_url("https://www.youtube.com/watch?v=ofTb57aZHZs") %>%
  use_align("center")
```

## Why can't you divide by zero?

```{r, echo=FALSE}
# Why can't you divide by zero? - TED-Ed


embed_url("https://www.youtube.com/watch?v=NKmGVE85GUU") %>%
  use_align("center")
```

## Yea he's chewing up my stats homework but that face though...

<blockquote class="reddit-card" data-card-created="1611607953"><a href="https://www.reddit.com/r/CatsBeingCats/comments/l4fnqx/yea_hes_chewing_up_my_stats_homework_but_that/">Yea he's chewing up my stats homework but that face though...</a> from <a href="http://www.reddit.com/r/CatsBeingCats">r/CatsBeingCats</a></blockquote>
<script async src="//embed.redditmedia.com/widgets/platform.js" charset="UTF-8"></script>

## Coding Kitty

https://hostrider.com/

## Family Trees Cut Short
```{r, echo=FALSE}
include_tweet("https://twitter.com/ShamiraGelbman/status/1354424275583848450")
```

## Democratic databases: science on GitHub

*Nature:* "[Democratic databases: science on GitHub](https://www.nature.com/news/democratic-databases-science-on-github-1.20719)" (Perkel, 2016).


## Ten simple rules for getting started on Twitter as a scientist

https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1007513


## NYT data ethics stuff

https://www.nytimes.com/2021/01/31/technology/facial-recognition-photo-tool.html