# LAB: Academic Freedom {#lab100}

Welcome to the Academic Freedom Monitoring Lab! In this lab, we will take on the role of academic freedom analysts, drawing on real-world political indicators from the **V-Dem** project. Our mission? To evaluate how free scholars are to teach, research, express ideas, and exchange knowledge across time and place.

We'll begin by exploring the 2025 **Academic Freedom Index**, which highlights concerning trends. In particular, the **United States** has experienced significant declines across all five dimensions of academic freedom between 2014 and 2024, most sharply in **freedom to research and teach** and **institutional autonomy**. According to the index, the U.S. is one of only 34 countries to show a statistically significant decline.

By the end of this lab, you will be able to reconstruct the same kinds of indicator-level visualizations that researchers use to monitor these trends—and place the U.S. within a broader global context.


## Learning goals {.unnumbered}

<!--
- Simulate Quantitative Variables: Learn to create simulated quantitative variables representing colonist attributes using `rnorm()`, `runif()`, and `rpois()`.

- Generate Character Variables: Use `rep()` to create categorical variables that represent different groups within the colonists, such as their professions or roles in the colony.

- Replication of Data Simulation: Utilize `replicate()` to repeat the data simulation process, representing multiple scenarios or batches of potential colonists.
-->

## Getting started and warming up {.unnumbered}

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE, # for regression output
  digits = 3
)
```

### Packages {.unnumbered}


We’ll be using **vdemdata** to access the Academic Freedom Index, as well as the **tidyverse** and **ggradar** packages for data manipulation and visualization. Install and load them with:

```{r load-packages, message=FALSE}
if (!require("devtools")) install.packages("devtools")
if (!require("vdemdata")) devtools::install_github("vdeminstitute/vdemdata")
if (!require("tidyverse")) install.packages("tidyverse")
if (!require("remotes")) install.packages("remotes")
if (!require("ggradar")) remotes::install_github("ricardo-bion/ggradar")
library(vdemdata)
library(tidyverse)
library(ggradar)
```

### Exercise 1: Academic Freedom in the United States {.unnumbered}

Let’s begin by examining how academic freedom has changed in the U.S. across five indicators from 2014 to 2024.

**1.1.** Filter the `vdem` dataset to include only the **United States** in the years **2014** and **2024**. Then select the five indicators that make up the Academic Freedom Index.

```{r exercise-one-part-one, error=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
vdem <- vdemdata::vdem %>%
  select(country_name, year, v2cafres, v2cafexch, v2cainsaut, v2casurv, v2clacfree)
```

**1.2.** Normalize these values so that they fall between 0 and 1, which is the format expected by `ggradar()`.

```{r exercise-one-part-two, error=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
vdem_scaled <- vdem %>%
  mutate(across(-c(year, country_name), scales::rescale))


u_data_scaled <- vdem %>% # vdem_scaled  %>%
  filter(country_name %in% c("United States of America"), year %in% c(1954, 2014, 2024)) %>%
  select(year,
    country_name,
    research_teach = v2cafres,
    exchange = v2cafexch,
    autonomy = v2cainsaut,
    integrity = v2casurv,
    expression = v2clacfree
  ) %>%
  mutate(year = as.factor(year)) %>%
  rename(group = year)
```

**1.3.** What do you notice about the raw or rescaled values? Do any indicators show more dramatic declines than others?


### Exercise 2: Building the Radar Plot {.unnumbered}

**2.1.** Now create a radar plot to compare the United States in 2014 and 2024 across the five indicators.

```{r exercise-two-part-one, error=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
u_data_scaled %>%
  filter(country_name == "United States of America") %>%
  select(-country_name) %>%
  ggradar(
    grid.min = 0, grid.mid = .5, grid.max = 1,
    group.line.width = 1.2,
    group.point.size = 3,
    legend.position = "bottom"
  ) +
  scale_color_viridis_d()

u_data_scaled %>%
  filter(country_name == "United Kingdom") %>%
  select(-country_name) %>%
  ggradar(
    grid.min = 0, grid.mid = .5, grid.max = 1,
    group.line.width = 1.2,
    group.point.size = 3,
    legend.position = "bottom"
  ) +
  scale_color_viridis_d()
```

**2.2.** Interpret the shape of the plot. Which dimensions remained relatively high? Which appear to be collapsing?


### Exercise 3: Looking Beyond the U.S. {.unnumbered}

Let’s now explore whether this decline is unique to the U.S. or part of a broader trend.

**3.1.** Identify countries where **freedom to research and teach** declined by **more than 0.2 points** between 2014 and 2024.

```{r exercise-three-part-one, error=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
afi_declines <- vdem %>%
  filter(year %in% c(1954, 2014, 2024)) %>%
  select(country_name, year, v2cafres) %>%
  pivot_wider(names_from = year, values_from = v2cafres) %>%
  mutate(diff = `2024` - `2014`) %>%
  filter(diff < -0.2) %>%
  arrange(diff)
```

**3.2.** Choose **one country** from the list and repeat the radar plot analysis for that country.

```{r exercise-three-part-two, error=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
chosen_country <- "Hungary" # Replace this with your selected country

country_data <- vdem %>%
  filter(country_name == chosen_country, year %in% c(2014, 2024)) %>%
  select(year,
    research_teach = v2cafres,
    exchange = v2cafexch,
    autonomy = v2cainsaut,
    integrity = v2casurv,
    expression = v2clacfree
  ) %>%
  #  mutate(across(-year, scales::rescale)) %>%
  mutate(year = as.factor(year)) %>%
  rename(group = year)

ggradar(country_data,
  grid.min = -3, grid.mid = 0, grid.max = 3,
  group.line.width = 1.2,
  group.point.size = 3,
  legend.position = "bottom"
)
```

**3.3.** In what ways does this country’s trajectory resemble or diverge from the U.S. case?


### Stretch Task: Where Else is Academic Freedom Declining? {.unnumbered}

Pick a region or group of countries (e.g., Latin America, OECD) and investigate whether they follow a similar pattern. Can you identify common features among countries that declined?
