# (PART) Module 99 {-}

```{r include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("data/ds4p_funyoutube.csv", sep="")
ds4p_urls <- read.csv("./data/ds4p_urls.csv")
library(tweetrmd) #... embedding tweets
```
```{r links, child="links.md"}
```


# Welcome to the !template module!

This module is designed to <<be a template for other modules>>. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][ds4p-pl-08]. Most of the slides used to make the videos in this module can be found in the [slides repo][ds4p-slides].


## Module Materials

* Slides from Lectures
  * [LINK](https://datascience4psych.github.io/slides/d17_functions/d17_functions.html)
* Activities
  * [LINK](https://github.com/DataScience4Psych/ae-09-uoeart-functions)
* Suggested Readings
  * All subchapters of this module, including [Notes on Functions][#functions-part1]
  * [r4ds](https://r4ds.had.co.nz/)
    * Sections on [functions](https://r4ds.had.co.nz/functions.html), and
    * [Iterations](https://r4ds.had.co.nz/iteration.html)
* Lab
	* [LAB A][lab08]
	* [LAB B][lab08b]

# Important Topic A! {-}

```{r, echo=FALSE}
include_tweet("https://twitter.com/MissyElliott/status/1386707191617527812")
```


You can follow along with the slides [here](LINK) if they do not appear below.

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d17_functions")
include_url(var_url, height = "400px")
```

## Activity {-}

You can find the materials for this activity [here](LINK). 

```{r, echo=FALSE}
# Filler Video
video_url=sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

```{r, echo=FALSE}
var_url=slide_url(ds4p_urls,"d01_welcome","#1")
include_url(var_url, height = "400px")
```


