--- 
title: "Data Science for Psychologists"
author: "S. Mason Garrison"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: R-Computing-Lab/DataScience4Psych

description: ""
---

# Preamble

This document has been built to accompany my Data Science for Psychologists graduate level R class. The class is a remixed version of Exploratory and Graphical Data Analysis (cite TUKEY), and is heavily influenced by a class I took in graduate school from my wonderfully supportive and patient advisor (Joe Rodgers). 

It has been built using the **bookdown** package and is an active work in progress. 


```{r setup, include=FALSE}

library(tidyverse)
library(citr)
library(RefManageR)
library(bookdown)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown','tidyverse'
), 'packages.bib')


knitr::opts_chunk$set(cache=FALSE)
```
```{r calendar, include=FALSE}
mon <- as.Date("2020-08-24")
cache=F
number_of_classes=29
number_of_weeks=15
advdate <- function(obj, adv, topic = NULL,assignment=NULL) {
 tmon <- obj + 7*(adv-1)
 ttue <- obj + 1 + 7*(adv-1)
 twed <- obj + 2 + 7*(adv-1)
 tthu <- obj + 3 + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 ttue <- format(ttue, format="%m/%d")
 twed <- format(twed, format="%m/%d")
 tthu <- format(tthu, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)

 if (is.null(topic)&is.null(assignment)){
 tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
 } else if(!is.null(assignment)){if(assignment %in% c("Friday","friday","f","fri")){
  tmp <- paste0(tfri)
 }else if(assignment %in% c("wednesday","Wednesday","w","wed","W")){
  tmp <- paste0(twed)
 } else if(assignment %in% c("monday","Monday","m")){
  tmp <- paste0(tmon)
 }
 }
 else{ 
   tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri, " : ", topic[adv])}
 return(tmp)
}

####
#devtools::install_github("crsh/citr")


#bib <- ReadBib("library.bib")
#bib <- c(bib, as.BibEntry(citation()))
#bib[length(bib)]$key = "Rcite"


topic=c(	
	"Introduction", 
	"Week 2",
	"Week 3",
	"Week 4",
	"Week 5",
	"Week 6",
	"Week 7",
	"Week 8",
	"Week 9", 
	"Week 10",
	"Week 11",
	"Week 12",
	"Week 13",
	"Week 14",
	"Week 15",
	"Week 16",
  )
```

