--- 
title: "Data Science for Psychologists"
author: "S. Mason Garrison"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: R-Computing-Lab/DataScience4Psych
description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Prerequisites

# Preamble

This document has been built to accompany my Data Science for Psychologists graduate level R class. The class is a remixed version of Exploratory and Graphical Data Analysis (cite TUKEY), and is heavily influenced by a class I took in graduate school from my wonderfully supportive and patient advisor (Joe Rodgers). 

It has been built using the **bookdown** package and is an active work in progress. 

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```
