--- 
title: "Data Science for Pychologists"
author: "S. Mason Garrison"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: R-Computing-Lab/DataScience4Psych

description: "This is a minimal example of using the bookdown package to write a book. The output format for this example is bookdown::gitbook."
---

# Prerequisites

This is a _sample_ book written in **Markdown**. You can use anything that Pandoc's Markdown supports, e.g., a math equation $a^2 + b^2 = c^2$.

The **bookdown** package can be installed from CRAN or Github:

```{r eval=FALSE}
install.packages("bookdown")
# or the development version
# devtools::install_github("rstudio/bookdown")
```

Remember each Rmd file contains one and only one chapter, and a chapter is defined by the first-level heading `#`.

To compile this example to PDF, you need XeLaTeX. You are recommended to install TinyTeX (which includes XeLaTeX): <https://yihui.name/tinytex/>.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=FALSE, dev='pdf')
mon <- as.Date("2020-08-24")
cache=F
number_of_classes=29
number_of_weeks=15
advdate <- function(obj, adv, topic = NULL,assignment=NULL) {
 tmon <- obj + 7*(adv-1)
 ttue <- obj + 1 + 7*(adv-1)
 twed <- obj + 2 + 7*(adv-1)
 tthu <- obj + 3 + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 ttue <- format(ttue, format="%m/%d")
 twed <- format(twed, format="%m/%d")
 tthu <- format(tthu, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)

 if (is.null(topic)&is.null(assignment)){
 tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
 } else if(!is.null(assignment)){if(assignment %in% c("Friday","friday","f","fri")){
  tmp <- paste0(tfri)
 }else if(assignment %in% c("wednesday","Wednesday","w","wed","W")){
  tmp <- paste0(twed)
 } else if(assignment %in% c("monday","Monday","m")){
  tmp <- paste0(tmon)
 }
 }
 else{ 
   tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri, " : ", topic[adv])}
 return(tmp)
}

####
#devtools::install_github("crsh/citr")
library(tidyverse)
library(citr)
library(RefManageR)



#bib <- ReadBib("library.bib")
#bib <- c(bib, as.BibEntry(citation()))
#bib[length(bib)]$key = "Rcite"








topic=c(
  "Introduction and What is Personality", 
  "Research Design in Personality",
  "Personality Assessment and Testing Controversies",
  "How do people differ? Trait Approach (Assignment 1)",
  "Big Five (Test 1)",
  "Personality over the Lifespan",#  "Spring Break",#  "",
  "Biological and Genetic Approaches to Psychology", 
  "Evolutionary Psychology", 
  "Personality Processes",
  "The Self  (Test 2)", 
  "Freud", 
  "Positive Psychology  (Assignment 2)", 
  "Cross-Cultural Psychology",
  "Personality and Health",
  "Special Topics (Assignment 3)" 
  )
```
```
