---
title: "Parameterized Report Example: `r params$lake`"
author: "Zachary M. Smith"
date: "2/27/2020"
output: html_document
params: 
  lake: Onondaga
---

```{r global-options, echo=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 4)
```

# Introduction

This document was created during the _2020 NABs R Markdown workshop_ and will serve as an example workflow for my future use of __R Markdown__. This document will represent an exploratory analysis of macroinvertebrate metrics and environmental variable associated with `r params$lake` Lake.

# Workflow

1. Open R Studio
2. Create an R Project
3. Add a data folder to the R project directory
4. Create an R Markdown document
5. Use Knit to compile the document.

# Import Data

Import Zachary M. Smith's thesis data provided by Zachary M. Smith at the R Markdown Crash Course Workshop on March 3rd, 2020.

* `unique_id` (character) unique sample ID
* `lake` (character) lake code: caz = Cazenovia lake, onon = Onondaga, and ot = Otisco
* `lat` (numeric) latitude of sampling location
* `long` (numeric) longitude of sampling location
* `substrate_size_d50` (numeric) median partical size from pebble count
* `conductivity` (numeric) specific conductivity (µS/cm)
* `richness` (numeric) taxonomic richness
* `shannon` (numeric) Shannon-Wiener diversity index values
* `pct_ephemeroptera` (numeric) relative abundance of Ephemeroptera taxa
* `pct_amphipoda` (numeric) relative abundance of Amphipoda taxa
* `pct_diptera` (numeric) relative abundance of Diptera taxa
* `dom_1` (numeric) relative abundance of the most dominant taxon observed in each sample
```{r import-data}
thesis.df <- read.csv(file.path(here::here(),
                                "data",
                                "zms_thesis_metrics.csv"),
                      stringsAsFactors = FALSE)
```

# Preprocessing

Load the __tidyverse__ packages into the global environment.
```{r load-packages, message=FALSE}
library(tidyverse)
```

```{r prep-lake}
thesis.df <- thesis.df %>% 
  mutate(lake = case_when(
    lake %in% "caz" ~ "Cazenovia",
    lake %in% "onon" ~ "Onondaga",
    lake %in% "ot" ~ "Otisco",
    TRUE ~ "ERROR"
  ),
  lake = factor(lake, levels = c("Onondaga",
                                 "Otisco",
                                 "Cazenovia")))
```

```{r filter-lake}
thesis.df <- thesis.df %>% 
  filter(lake %in% params$lake)
```

For more details about the __DT__ package visit https://rstudio.github.io/DT/.
```{r dt}
library(DT)

datatable(thesis.df, options = list(scrollX = TRUE))
```

# Study Area

For more details about the __leaflet__ package visit https://rstudio.github.io/leaflet/.
```{r leaflet}
library(leaflet)

pal <- colorFactor(c("#619Cff", "#F8766D", "#00BA38"),
                   domain = c("Cazenovia", "Onondaga", "Otisco"))

leaflet(data = thesis.df,
        options = leafletOptions(minZoom = 7,
                                 maxZoom = 13)) %>% 
  addTiles() %>% 
  addCircleMarkers(~long, ~lat,
                   fillOpacity = 0.75,
                   fillColor = ~pal(lake),
                   stroke = FALSE,
             popup = paste("Sample ID:", thesis.df$unique_id, "<br/>",
                           "Lake:", thesis.df$lake, "<br/>",
                           "Latitude:", thesis.df$lat, "<br/>",
                           "Longitude:", thesis.df$long)
             )

```

# Plot

## Scatter Plot

For more details about the __plotly__ package visit https://plot.ly/ggplot2/.
```{r scatter-sub-dip, message=FALSE}
library(plotly)

scatter.plot <- ggplot(thesis.df, aes(substrate_size_d50, pct_diptera)) +
  geom_point(aes(color = lake)) +
  geom_smooth(method = "lm")

ggplotly(scatter.plot)
```

## Boxplot

```{r boxplot-lake-rich}
ggplot(thesis.df, aes(lake, richness, fill = lake)) +
  geom_boxplot()
```

# Conclusions

1. The relative abundance of Diptera taxa in `r params$lake` Lake ranged from `r min(thesis.df$pct_diptera)` to `r max(thesis.df$pct_diptera)`.
2. The median taxonomic richness observed in `r params$lake` Lake was `r median(thesis.df$richness)`.

