---
title: "Parameterized Report Example: `r params$lake`"
author: "Zachary M. Smith"
date: "2/27/2020"
output: html_document
params: 
  lake: Onondaga
---

```{r global-options, echo=FALSE}
knitr::opts_chunk$set(fig.width = 8, fig.height = 4)
```

# Introduction

This document was created during the _2020 NABs R Markdown workshop_ and will serve as an example workflow for my future use of __R Markdown__. This document will represent an exploratory analysis of macroinvertebrate metrics and environmental variable associated with `r params$lake` Lake.

# Workflow

1. Open R Studio
2. Create an R Project
3. Add a data folder to the R project directory
4. Create an R Markdown document
5. Use Knit to compile the document.

# Import Data

Import Zachary M. Smith's thesis data provided by Zachary M. Smith at the R Markdown Crash Course Workshop on March 3rd, 2020.

* `unique_id` (character) unique sample ID
* `lake` (character) lake code: caz = Cazenovia lake, onon = Onondaga, and ot = Otisco
* `lat` (numeric) latitude of sampling location
* `long` (numeric) longitude of sampling location
* `substrate_size_d50` (numeric) median partical size from pebble count
* `conductivity` (numeric) specific conductivity (µS/cm)
* `richness` (numeric) taxonomic richness
* `shannon` (numeric) Shannon-Wiener diversity index values
* `pct_ephemeroptera` (numeric) relative abundance of Ephemeroptera taxa
* `pct_amphipoda` (numeric) relative abundance of Amphipoda taxa
* `pct_diptera` (numeric) relative abundance of Diptera taxa
* `dom_1` (numeric) relative abundance of the most dominant taxon observed in each sample
```{r import-data}
df_thesis <- read.csv(
  file.path(
    here::here(),
    "data",
    "zms_thesis_metrics.csv"
  ),
  stringsAsFactors = FALSE
)
```

# Preprocessing

Load the __tidyverse__ packages into the global environment.
```{r load-packages, message=FALSE}
library(tidyverse)
```

```{r prep-lake}
df_thesis <- df_thesis %>%
  mutate(
    lake = case_when(
      lake %in% "caz" ~ "Cazenovia",
      lake %in% "onon" ~ "Onondaga",
      lake %in% "ot" ~ "Otisco",
      TRUE ~ "ERROR"
    ),
    lake = factor(lake, levels = c(
      "Onondaga",
      "Otisco",
      "Cazenovia"
    ))
  )
```

```{r filter-lake}
df_thesis <- df_thesis %>%
  filter(lake %in% params$lake)
```

For more details about the __DT__ package visit https://rstudio.github.io/DT/.
```{r dt}
library(DT)

datatable(df_thesis, options = list(scrollX = TRUE))
```

# Study Area

For more details about the __leaflet__ package visit https://rstudio.github.io/leaflet/.
```{r leaflet}
library(leaflet)

pal <- colorFactor(c("#619Cff", "#F8766D", "#00BA38"),
  domain = c("Cazenovia", "Onondaga", "Otisco")
)

leaflet(
  data = df_thesis,
  options = leafletOptions(
    minZoom = 7,
    maxZoom = 13
  )
) %>%
  addTiles() %>%
  addCircleMarkers(~long, ~lat,
    fillOpacity = 0.75,
    fillColor = ~ pal(lake),
    stroke = FALSE,
    popup = paste(
      "Sample ID:", df_thesis$unique_id, "<br/>",
      "Lake:", df_thesis$lake, "<br/>",
      "Latitude:", df_thesis$lat, "<br/>",
      "Longitude:", df_thesis$long
    )
  )
```

# Plot

## Scatter Plot

For more details about the __plotly__ package visit https://plot.ly/ggplot2/.
```{r scatter-sub-dip, message=FALSE}
library(plotly)

p_scatter<- ggplot(df_thesis, aes(substrate_size_d50, pct_diptera)) +
  geom_point(aes(color = lake)) +
  geom_smooth(method = "lm")

ggplotly(p_scatter)
```

## Boxplot

```{r boxplot-lake-rich}
ggplot(df_thesis, aes(lake, richness, fill = lake)) +
  geom_boxplot()
```

# Conclusions

1. The relative abundance of Diptera taxa in `r params$lake` Lake ranged from `r min(df_thesis$pct_diptera)` to `r max(df_thesis$pct_diptera)`.
2. The median taxonomic richness observed in `r params$lake` Lake was `r median(df_thesis$richness)`.

