--- 
title: "Data Science for Psychologists"
author: "S. Mason Garrison"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [book.bib, packages.bib]
biblio-style: apalike
link-citations: yes
github-repo: R-Computing-Lab/DataScience4Psych

description: ""
---

# Preamble

This document has been built to accompany my Data Science for Psychologists graduate level R class. The class is a remixed version of Exploratory and Graphical Data Analysis (cite TUKEY), and is heavily influenced by a class I took in graduate school from my wonderfully supportive and patient advisor (Joe Rodgers). 

It has been built using the **bookdown** package and is an active work in progress. 


```{r setup, include=FALSE}

library(tidyverse)
library(citr)
library(RefManageR)
library(bookdown)

# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown','tidyverse'
), 'packages.bib')


knitr::opts_chunk$set(cache=FALSE)
```
```{r calendar, include=FALSE}
mon <- as.Date("2020-08-24")
cache=F
number_of_classes=29
number_of_weeks=15
advdate <- function(obj, adv, topic = NULL,assignment=NULL) {
 tmon <- obj + 7*(adv-1)
 ttue <- obj + 1 + 7*(adv-1)
 twed <- obj + 2 + 7*(adv-1)
 tthu <- obj + 3 + 7*(adv-1)
 tfri <- obj + 4 + 7*(adv-1)
 tmon <- format(tmon, format="%m/%d")
 ttue <- format(ttue, format="%m/%d")
 twed <- format(twed, format="%m/%d")
 tthu <- format(tthu, format="%m/%d")
 tfri <- format(tfri, format="%m/%d")
 zadv <- sprintf("%02d", adv)

 if (is.null(topic)&is.null(assignment)){
 tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri)
 } else if(!is.null(assignment)){if(assignment %in% c("Friday","friday","f","fri")){
  tmp <- paste0(tfri)
 }else if(assignment %in% c("wednesday","Wednesday","w","wed","W")){
  tmp <- paste0(twed)
 } else if(assignment %in% c("monday","Monday","m")){
  tmp <- paste0(tmon)
 }
 }
 else{ 
   tmp <- paste("Week ",zadv,sep='',", ", tmon," - ",tfri, " : ", topic[adv])}
 return(tmp)
}

####
#devtools::install_github("crsh/citr")


#bib <- ReadBib("library.bib")
#bib <- c(bib, as.BibEntry(citation()))
#bib[length(bib)]$key = "Rcite"


topic=c(	
	"Introduction", 
	"Week 2",
	"Week 3",
	"Week 4",
	"Week 5",
	"Week 6",
	"Week 7",
	"Week 8",
	"Week 9", 
	"Week 10",
	"Week 11",
	"Week 12",
	"Week 13",
	"Week 14",
	"Week 15",
	"Week 16",
  )
```

# Syllabus {#syllabus}

## Course Description

This class provides a survey of theory and research on the structure and function of human personality, with attention to the relationship to cognition, emotion, motivation, and behavior. It requires P-PSY 151. In addition, we will discuss other individual differences, such as intelligence.

```{r child = 'E:/Dropbox/Lab/Teaching/Syllabi/common/content/doingwell_irl.Rmd'}
```


## Broad goals for the course:
- 1. The first is for you to understand what personality psychologists think. We will cover classic and current theories, research, and applications of Personality psychology.
- 2. The second is for you to learn how personality psychologists study their field scientifically. We will discuss the procedures used to test theories and to study issues related to personality.
- 3. The third is to (hopefully) recognize some relevance of personality psychology for your personal and/or professional life.


## Required Materials




### How to use the required text:

The text [@funder_personality_2016] is intended to supplement the lectures. The lectures don't follow the order of chapters in the text and the text covers some material that won't be covered in the lectures, i.e., not all the information in each chapter is perfectly pertinent to the course requirements.


# Course Assignments

## Exams 
There will be 3 exams in this course.  The best 2 out of 3 exam grades will be used to determine your total exam grade.  You can drop any of the exams for any reason BUT YOU MUST PASS THE LAST EXAM (held at TBD on Dec TBD).  If you do not pass the final exam, your score on that exam will count as one of your 2 grades. If you do pass the final exam, then the best 2 grades will be used (irrespective of your score on the final exam).


### Test Dates

- Test 1: Friday, `r advdate(mon, 5, assignment="f")`

- Test 2: Friday, `r advdate(mon, 10, assignment="f")`

- Test 3: During Finals Week


## Short Assignments 
There will be three (3) short assignments in this course. These assignments will allow you to incorporate some of your own interests into the course. Such interests could be related to your career, another class you're taking, a hobby you're exploring, or some other random fancy. They will typically be approximately two (2) pages. I will provide more specific written guidelines at least two weeks before each assignment is due.


The best two (2) out of three (3) short assignments will be used to determine your total short assignments.  You can skip one of the assignments for any reason.


### Short Assignment Dates

- Assignment 1: Friday, `r advdate(mon, 4, assignment="f")`: Critique of an online personality test.

- Assignment 2: Friday, `r advdate(mon, 12, assignment="f")`: Critique of a personality-related media article.

- Assignment 3: Friday, `r advdate(mon, 15, assignment="f")`: Critique of a personality-related YouTube video.

## Engagement Activities

There will be multiple engagement activities in this course. These activities will allow you engage with the material for each module. Details about the specific activities will be provided on canvas. Students must complete two activities per module. For students in the blended section, they can attend their weekly in-person session and count it as an engagement activity.


## Grading Policy

```{r child = 'E:/Dropbox/Lab/Teaching/Syllabi/common/content/grading.Rmd'}
```

- 200 points of your grade will be determined by your best two exams.
    - The final is not optional.
- 50 points of your grade will be determined by completing engagement activities.
<!---
your attendance and participation in engagement activities. Generally, ask questions and answer them.
-->
- 100 points of your grade will be determined by your best two short assignments.


## Course Policies

```{r child = 'E:/Dropbox/Lab/Teaching/Syllabi/common/content/policies.Rmd'}
```

<!---
Schedule
-->

## `r advdate(mon, 1, topic)`
<!---
- Monday: First Day of Class / Introduction
- Wednesday: What is personality? (Chapter 1)
- Friday: Research in Personality: Sources of Data (Chapter 2)
-->
## `r advdate(mon, 2, topic)`
<!---
- Monday: NO CLASS MLK Day
- Wednesday: Research Design: Validity (Chapter 2)
- Friday: Research Design: Generalizability (Chapter 2; Short Assignment 1 Handed Out)
-->
##  `r advdate(mon, 3, topic)` 
<!---
- Monday: Personality Assessment (Chapter 3)
- Tuesday: "Last day to add full-term class"
- Wednesday: Evaluation and Ethics (Chapter 3)
- Friday: Person versus Situation (Chapter 4)
-->

##  `r advdate(mon, 4, topic)`
<!---
- Monday: What are traits? What are the basic traits? (Chapter 4)
- Wednesday: Personality Judgement (Chapter 5)
- Friday: Using Personality Traits to Understand Behavior (Chapter 6; Assignment 1 Due)
-->
##  `r advdate(mon, 5, topic)`
<!---
- Monday: Big Five (Chapter 6)
- Wednesday: Review
- Friday: Test 1
-->
##  `r advdate(mon, 6, topic)`
<!---
- Monday: Brains (Chapter 8)
- Wednesday: More Brains (Chapter 8)
- Friday: Behavior Genetics (Chapter 9)
-->
##  `r advdate(mon, 7, topic)` 
<!---
- Monday: Behavior Genetics (Chapter 9)
- Wednesday: Evolutionary Psychology (Chapter 9)
- Friday: Evolutionary Psychology (Chapter 9)
-->
##  `r advdate(mon, 8, topic)`
<!---
- Monday: Personality over the Lifespan (Chapter 7)
- Wednesday: Personality over the Lifespan (Chapter 7)
- Friday: Personality Processes (Chapter 15)
-->
##  `r advdate(mon, 9, topic)`
<!---
- Monday: No Class
- Wednesday: No Class
- Friday: No Class
-->
##  `r advdate(mon, 10, topic)`
<!---
- Monday: Personality Processes (Chapter 15)
- Wednesday: Review
- Friday: Test 2
-->
##  `r advdate(mon, 11, topic)`
<!---
- Monday: The Self (Chapter 16) Last day to drop with a grade of "W"
- Wednesday: The Self (Chapter 16)
- Friday:  Freud (Chapter 10) 
-->
##  `r advdate(mon, 12, topic)`
<!---
- Monday: Freud (Chapter 10)
- Wednesday: Freud (Chapter 10)
- Friday: Neo-Freud (Chapter 11; Assignment 2 Due)
-->
## `r advdate(mon, 13, topic)`
<!---
- Monday:  Neo-Freud (Chapter 11) 
- Wednesday: Positive Psychology
- Friday: Positive Psychology
-->
##  `r advdate(mon, 14, topic)`
<!---
- Monday: Cross-Cultural Psychology
- Wednesday: Cross-Cultural Psychology
- Friday: Cross-Cultural Psychology
-->

##  `r advdate(mon, 15, topic)`
<!---
- Monday: Personality and Physical Health
- Wednesday: Personality and Mental Health
- Friday: Personality Disorders

##  `r advdate(mon, 16, topic)` 

- Monday: Special Topics
- Wednesday: Review; Assignment 3 Due; Last Day of Class
-->
## Final Exam 

- TBD on DEC TBD

