# LAB: [Lab Title] {#labXX}

```{r include=FALSE}
library(tufte)
library(knitr)
options(
  htmltools.dir.version = FALSE, # for blogdown
  show.signif.stars = FALSE, # for regression output
  digits = 2
)
knitr::opts_chunk$set(eval = FALSE)
```

Brief introduction to the lab topic and its relevance

## Learning goals {.unnumbered}

[Learning goal 1]
[Learning goal 2]
[Learning goal 3]

## Getting started {.unnumbered}
Go to the course GitHub organization and locate your lab repo, which should be named lab-XX-NAME. Grab the URL of the repo, and clone it in RStudio. First, open the R Markdown document lab-XX.Rmd and Knit it. Make sure it compiles without errors. The output will be in the file markdown .md file with the same name.

### Packages {.unnumbered}
In this lab, we will work with the following packages:

```{r}
library(tidyverse)
# Add other necessary packages
```

### Data {.unnumbered}

[Description of the dataset(s) used in this lab]

## Exercises {.unnumbered}

### Part 1: [Section Title] {.unnumbered}

[Exercise 1 description]

```{r}
# Exercise 1 code
```

<details>
  <summary>Click for a hint</summary>
  [Hint content]
</details>

[Exercise 2 description]

🧶 ✅ ⬆️ Knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

### Part 2: [Section Title] {.unnumbered}

[Exercise 3 description]

```{r}
# Exercise 3 code
```

<details>
  <summary>Click for a hint</summary>
  [Hint content]
  
</details>


[Exercise 4 description]

```{r}
# Exercise 4 code
```   


🧶 ✅ ⬆️ Now is a good time to knit your document and commit and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

Wrap up {.unnumbered}
[Summary of the lab and key takeaways]

🧶 ✅ ⬆️ Knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards and review the md document on GitHub to make sure you're happy with the final state of your work.
