# (PART) Module XX {-}

```{r modXX-setup, include = FALSE}
source("common.R")
ds4p_funyoutube <- read.csv("metadata/ds4p_funyoutube.csv", sep = "")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd)
library(vembedr)
```
```{r links, child="links.md"}
```


# Welcome to Data Exploration!

This module is designed to introduce you to exploratory data analysis with R. Please watch the videos and work your way through the notes. **The videos start on the next page.** You can find the video playlist for this module [here][pl_01]. Most of the slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides from Lectures
  * [Exploring Data](https://datascience4psych.github.io/slides/d01_welcome/d01_welcome.html)
  * [Visualizing Distributions](https://datascience4psych.github.io/slides/d02_toolkit/d02_toolkit.html)
* Activities
  * [Penguin Exploration](https://github.com/DataScience4Psych/ae-XX-penguins)
* Suggested Readings
  * All subchapters of this module
  * [R4DS: Exploratory Data Analysis](https://r4ds.had.co.nz/exploratory-data-analysis.html)
  * [R4DS: Data Visualization](https://r4ds.had.co.nz/data-visualisation.html)
* Lab
  * [Penguin Research Station](#labXX)

### Estimated Video Length

<!--https://www.youtube.com/playlist?list=PLKrrdtYgOUYao_7t5ycK4KDXNKaY-ECup & https://ytplaylist-len.herokuapp.com/ -->

No of videos : 5

Average length of video: 10 minutes, 30 seconds

Total length of playlist: 52 minutes, 30 seconds


# LEC: Introduction to Exploratory Data Analysis {-}

```{r modXX-intro-video-latex, echo=FALSE, eval = knitr::is_latex_output()}
embed_youtube_alt("VIDEO_ID_HERE")
```
```{r modXX-intro-video-html, echo=FALSE, eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=VIDEO_ID_HERE" %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here][d01_welcome] if they are too small or do not appear below.

```{r modXX-embed-intro-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome") %>%
  include_url(height = "400px")
```

## Learning goals

By the end of this lecture, you should be able to:

- Describe the role of exploratory data analysis in the data science workflow
- Identify appropriate visualizations for different variable types
- Use summary statistics to characterize distributions

```{r modXX-embed-learning-goals-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome", "#4") %>%
  include_url(height = "400px")
```

## Key Concepts

Exploratory data analysis (EDA) is about asking questions of your data and searching for answers using visualization and transformation. As Tukey (1977) wrote:

> "Far better an approximate answer to the right question, which is often vague, than an exact answer to the wrong question, which can always be made precise."

The EDA cycle involves:

1. Generate questions about your data
2. Search for answers by visualizing, transforming, and modeling
3. Use what you learn to refine your questions or generate new ones

```{r modXX-include-tweet, echo=FALSE}
try_include_tweet("https://twitter.com/hadleywickham/status/1363920786416562179")
```

## Hands-On Practice

Now it's your turn. Try exploring the `mpg` dataset that comes with ggplot2:

```{r modXX-mpg-exploration, eval=FALSE}
library(tidyverse)
glimpse(mpg)

ggplot(mpg, aes(x = displ, y = hwy)) +
  geom_point(aes(color = class)) +
  labs(
    title = "Engine Displacement vs. Highway MPG",
    x = "Engine Displacement (L)",
    y = "Highway Miles per Gallon"
  ) +
  theme_minimal()
```


# LEC: Visualizing Distributions {-}

```{r modXX-dist-video-latex, echo=FALSE, eval = knitr::is_latex_output()}
embed_youtube_alt("VIDEO_ID_HERE")
```
```{r modXX-dist-video-html, echo=FALSE, eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=VIDEO_ID_HERE" %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here](https://datascience4psych.github.io/slides/d02_toolkit/d02_toolkit.html) if you would like to open them full-screen.

```{r modXX-embed-dist-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d02_toolkit") %>%
  include_url(height = "400px")
```


## ACT: Penguin Exploration {-}

```{r modXX-embed-act-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome", "#53") %>%
  include_url(height = "400px")
```

You can find the materials for this activity [here](https://github.com/DataScience4Psych/ae-XX-penguins). The compiled version should look something like the following...

```{r modXX-embed-act-preview, echo=FALSE}
"https://datascience4psych.github.io/ae-XX-penguins/penguins.html" %>%
  include_url(height = "400px")
```


# Wrapping Up {-}

In this module, you've learned about the foundations of exploratory data analysis. Remember that EDA is iterative — you'll revisit these techniques throughout the course as we work with increasingly complex datasets.

```{r modXX-filler-video, echo=FALSE}
video_url <- sample_no_surprises(ds4p_funyoutube$funyoutube)
embed_url(video_url) %>%
  use_align("center")
```

```{r courselinks, child="includes/courselinks.md"}
```
