# ACT: Penguin Exploration {#aeXX}

```{r actXX-setup, include = FALSE}
source("common.R")

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd)
library(vembedr)
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")
```


## Penguin Exploration

```{r actXX-embed-activity-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome", "#53") %>%
  include_url(height = "400px")
```

You can find the materials for the Penguin Exploration activity [here](https://github.com/DataScience4Psych/ae-XX-penguins).
The compiled version should look something like the following... This activity is designed to help you practice data wrangling and visualization with a real ecological dataset.

```{r actXX-embed-compiled-example, echo=FALSE}
"https://datascience4psych.github.io/ae-XX-penguins/penguins.html" %>%
  include_url(height = "400px")
```

<details>
  <summary>Alternative: Iris Dataset Activity</summary>

## Iris Dataset

If you'd like an alternative dataset to practice with, this version uses the classic iris dataset. The core skills are the same — filtering, grouping, summarizing, and visualizing — but with a different biological context.

```{r actXX-embed-alt-activity-slides, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome", "#54") %>%
  include_url(height = "400px")
```

You can find the materials for the Iris version of this activity [here](https://github.com/DataScience4Psych/ae-XX-iris).
The compiled version should look something like the following...

```{r actXX-embed-alt-compiled-example, echo=FALSE}
"https://datascience4psych.github.io/ae-XX-iris/iris.html" %>%
  include_url(height = "400px")
```

</details>


## Learning Outcomes

```{r actXX-embed-learning-outcomes, echo=FALSE, error=TRUE}
slide_url(ds4p_urls, "d01_welcome", "#56") %>%
  include_url(height = "400px")
```

By the end of this activity, you should be able to:

- Load and inspect a dataset using `glimpse()` and `summary()`
- Create basic visualizations using `ggplot2`
- Filter and summarize data using `dplyr` verbs
- Practice the Git/GitHub workflow: clone, edit, commit, push


```{r courselinks, child="includes/courselinks.md"}
```
