# ACT: Exploring the Starwars Dataset {#aeXX}

```{r actscXX-setup, include = FALSE}
source("common.R")

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd)
library(vembedr)
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")

library(tidyverse)
```

In this activity, you'll practice data wrangling and visualization using the `starwars` dataset that comes built into the `dplyr` package. This dataset contains information about 87 characters from the Star Wars universe, including their physical characteristics, homeworld, and species.


## Getting Started {.unnumbered}

### Packages {.unnumbered}

```{r actscXX-install-packages, eval = FALSE}
# Install packages if needed
install.packages("tidyverse")
```

```{r actscXX-load-packages, message=FALSE}
library(tidyverse)
```

### Data {.unnumbered}

The `starwars` dataset is automatically available when you load `dplyr` (part of the tidyverse). Let's take a first look:

```{r actscXX-glimpse-data}
glimpse(starwars)
```

> The `starwars` dataset contains `r nrow(starwars)` characters and `r ncol(starwars)` variables. Some columns are list-columns (films, vehicles, starships) — we'll ignore those for now and focus on the simpler columns.


## Part 1: Exploring the Data {.unnumbered}

**Task 1.** How many distinct species are represented in the dataset? Which species has the most characters?

```{r actscXX-task-one, eval=FALSE}
starwars %>%
  count(___) %>%
  arrange(desc(n))
```

> Hint: The `count()` function groups and counts in one step. Use `arrange(desc(n))` to see the most common species first.


**Task 2.** What is the average height and mass for each species that has at least 3 representatives? Remember to handle missing values.

```{r actscXX-task-two, eval=FALSE}
starwars %>%
  group_by(species) %>%
  filter(n() >= 3) %>%
  summarize(
    n = n(),
    avg_height = ___,
    avg_mass = ___
  )
```


🧶 ✅ ⬆️ Knit, commit, and push your changes to GitHub with the commit message "Completed Part 1: data exploration". Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.


## Part 2: Visualization {.unnumbered}

**Task 3.** Create a scatter plot of `height` vs. `mass`, colored by `species`. Only include species with at least 3 characters (to reduce clutter). What do you notice about the relationship? Is there an outlier?

```{r actscXX-task-three, fig.fullwidth=TRUE, eval=FALSE}
starwars %>%
  group_by(species) %>%
  filter(n() >= 3) %>%
  ungroup() %>%
  ggplot(aes(x = ___, y = ___, color = ___)) +
  geom_point(size = 3, alpha = 0.7) +
  labs(
    title = "Height vs. Mass of Star Wars Characters",
    x = "Height (cm)",
    y = "Mass (kg)"
  ) +
  theme_minimal()
```


**Task 4.** The outlier in the previous plot is Jabba the Hutt (mass = 1358 kg!). Remove Jabba and recreate the plot. Add a trend line using `geom_smooth(method = "lm")`. How does the relationship look now?

```{r actscXX-task-four, fig.fullwidth=TRUE, eval=FALSE}
# Your code here — filter out Jabba, then create the plot
```


🧶 ✅ ⬆️ Knit, commit, and push your changes to GitHub with the commit message "Completed Part 2: visualization". Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.


## Part 3: Data Wrangling Challenge {.unnumbered}

**Task 5.** Create a summary table that shows, for each homeworld with at least 2 characters:

- Number of characters
- Average height
- Most common eye color (mode)

Sort the result by number of characters, descending.

<details>
  <summary>Click for a hint about finding the mode</summary>

  There's no built-in `mode()` function for categorical data in R (confusingly, `mode()` returns the storage type!). Instead, you can use this approach inside `summarize()`:

  ```r
  most_common_eye = names(sort(table(eye_color), decreasing = TRUE))[1]
  ```

  Or more tidyverse-style:

  ```r
  most_common_eye = eye_color[which.max(tabulate(match(eye_color, unique(eye_color))))]
  ```
</details>


## Stretch Goal {.unnumbered}

**Task 6.** Which films feature characters from the most diverse set of homeworlds? To answer this, you'll need to work with the `films` list-column using `unnest()`.

```{r actscXX-task-six, eval=FALSE}
starwars %>%
  select(name, homeworld, films) %>%
  unnest(films) %>%
  group_by(films) %>%
  summarize(
    n_characters = n_distinct(name),
    n_homeworlds = n_distinct(homeworld, na.rm = TRUE)
  ) %>%
  arrange(desc(n_homeworlds))
```


## Wrap Up {.unnumbered}

In this activity, you practiced:

- Exploring data with `glimpse()`, `count()`, and `filter()`
- Summarizing data with `group_by()` and `summarize()`
- Visualizing relationships with `ggplot2`
- Handling outliers and missing data
- Working with list-columns using `unnest()`

🧶 ✅ ⬆️ Knit, commit, and push your changes to GitHub with an appropriate commit message. Make sure to commit and push all changed files so that your Git pane is cleared up afterwards.

```{r courselinks, child="includes/courselinks.md"}
```
