```{r lecXX-setup, include = FALSE}
source("common.R")
ds4p_urls <- read.csv("./metadata/ds4p_urls.csv")

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd)
library(vembedr)
```

# LEC: Hands on Data Transformation {#handson_transform}

## Reshaping data with pivot functions

```{r lecXX-reshape-video-latex, echo=FALSE, eval = knitr::is_latex_output()}
embed_youtube_alt("VIDEO_ID_HERE")
```
```{r lecXX-reshape-video-html, echo=FALSE, eval = knitr::is_html_output()}
"https://www.youtube.com/watch?v=VIDEO_ID_HERE" %>%
  embed_url() %>%
  use_align("center")
```

You can follow along with the slides [here][d09_wrangle] if you would like to open them full-screen.

```{r lecXX-embed-reshape-slides, echo=FALSE, error=TRUE}
ds4p_urls %>%
  slide_url("d09_wrangle") %>%
  include_url(height = "400px")
```


## Getting started

We'll be working with the `table1` dataset from the **tidyr** package, which comes loaded with the tidyverse. This dataset contains tuberculosis case counts for three countries across two years.

```{r lecXX-load-packages, message=FALSE, warning=FALSE}
library(tidyverse)
```

Let's take a look at the data in its tidy form:

```{r lecXX-glimpse-data}
table1
```

This is "tidy" because each variable has its own column, each observation has its own row, and each value has its own cell. But data doesn't always arrive this way!


## Wide vs. long format

Data often comes in "wide" format, where each time point or category gets its own column. Compare `table1` (tidy/long) with `table4a` (wide):

```{r lecXX-show-wide}
table4a # cases in wide format
```

To make this tidy, we need to **pivot longer** — gathering the year columns into key-value pairs:

```{r lecXX-pivot-longer-demo}
table4a %>%
  pivot_longer(
    cols = c(`1999`, `2000`),
    names_to = "year",
    values_to = "cases"
  )
```

Notice how we went from 3 rows and 3 columns to 6 rows and 3 columns. Each row now represents one country in one year — a single observation.


## ACT: Hotels

```{r lecXX-embed-act-slide, echo=FALSE, error=TRUE}
ds4p_urls %>%
  slide_url("d09_wrangle", "#18") %>%
  include_url(height = "400px")
```

You can find the materials for the Hotels activity [here](https://github.com/DataScience4Psych/ae-04-hotels-datawrangling). The compiled version should look something like the following...

```{r lecXX-embed-act-preview, echo=FALSE}
"https://datascience4psych.github.io/ae-04-hotels-datawrangling/hotels-datawrangling.html" %>%
  include_url(height = "400px")
```


## Going wider with `pivot_wider()`

Sometimes you need the reverse operation. If your data is _too_ long, you can spread it wider:

```{r lecXX-pivot-wider-demo}
table2 %>%
  pivot_wider(
    names_from = type,
    values_from = count
  )
```


## Separating and uniting columns

Sometimes a single column contains two pieces of information. The `separate()` function splits one column into multiple:

```{r lecXX-separate-demo}
table3 %>%
  separate(rate, into = c("cases", "population"), sep = "/")
```

The reverse operation, `unite()`, pastes columns together:

```{r lecXX-unite-demo}
table3 %>%
  separate(rate, into = c("cases", "population"), sep = "/") %>%
  unite(rate_combined, cases, population, sep = " out of ")
```


## Putting it all together

Let's practice the full pipeline. Starting with `table4a` (wide cases) and `table4b` (wide population), let's create a tidy dataset with a computed rate:

```{r lecXX-full-pipeline}
tidy_cases <- table4a %>%
  pivot_longer(
    cols = c(`1999`, `2000`),
    names_to = "year",
    values_to = "cases"
  )

tidy_pop <- table4b %>%
  pivot_longer(
    cols = c(`1999`, `2000`),
    names_to = "year",
    values_to = "population"
  )

# Join and compute rate
left_join(tidy_cases, tidy_pop, by = c("country", "year")) %>%
  mutate(rate = cases / population * 10000)
```

This gives us the TB rate per 10,000 people — much more interpretable than raw counts!


## Resources

Tidying data:

* [R for Data Science: Tidy Data](https://r4ds.had.co.nz/tidy-data.html)
* [Pivoting vignette](https://tidyr.tidyverse.org/articles/pivot.html) from the tidyr package
* [RStudio Data Import Cheat Sheet](https://github.com/rstudio/cheatsheets/raw/master/data-import.pdf)

```{r lecXX-include-tweet, echo=FALSE}
try_include_tweet("https://twitter.com/hadleywickham/status/1363920786416562179")
```

```{r courselinks, child="includes/courselinks.md"}
```
