# (PART) Supplementary Materials {-}

```{r specXX-setup, include = FALSE}
source("common.R")
library(tidyverse)
library(DT)
library(knitr)

if (!require("tweetrmd")) devtools::install_github("gadenbuie/tweetrmd")
library(tweetrmd)
library(vembedr)
```

# Recommended R Packages for Psychology {.unnumbered}

This page collects useful R packages organized by the type of task you're trying to accomplish. Whether you're cleaning data, making visualizations, or running statistical models, there's likely a package that can help. Think of this as your field guide to the R ecosystem.


## Data Wrangling {.unnumbered}

The tidyverse provides the core toolkit for getting data into shape:

| Package | Purpose | Key Functions |
|---------|---------|---------------|
| **dplyr** | Data manipulation | `filter()`, `mutate()`, `summarize()`, `group_by()` |
| **tidyr** | Reshaping data | `pivot_longer()`, `pivot_wider()`, `separate()` |
| **stringr** | String manipulation | `str_detect()`, `str_replace()`, `str_extract()` |
| **forcats** | Working with factors | `fct_relevel()`, `fct_reorder()`, `fct_lump()` |
| **lubridate** | Dates and times | `ymd()`, `mdy()`, `interval()` |

> If you're new to R, start with **dplyr** and **ggplot2**. These two packages will carry you through 80% of your data tasks.


## Visualization {.unnumbered}

### Static plots

```{r specXX-example-plot, echo=FALSE, message=FALSE, warning=FALSE, fig.width=8, fig.height=4}
library(palmerpenguins)

penguins %>%
  drop_na() %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, color = species)) +
  geom_point(alpha = 0.7, size = 2) +
  geom_smooth(method = "lm", se = FALSE, formula = "y ~ x") +
  scale_color_viridis_d() +
  labs(
    title = "Palmer Penguins: Flipper Length vs. Body Mass",
    subtitle = "Data from the palmerpenguins package",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  ) +
  theme_minimal()
```

| Package | Purpose |
|---------|---------|
| **ggplot2** | Grammar of graphics — the foundation for R visualization |
| **patchwork** | Combine multiple ggplots into one figure |
| **ggrepel** | Non-overlapping text labels on plots |
| **scales** | Nice axis formatting (commas, percentages, dollars) |

### Interactive plots

| Package | Purpose |
|---------|---------|
| **plotly** | Turn ggplots interactive with `ggplotly()` |
| **DT** | Interactive data tables |
| **leaflet** | Interactive maps |


## Statistical Modeling {.unnumbered}

| Package | Purpose |
|---------|---------|
| **broom** | Tidy model output into data frames |
| **lme4** | Mixed-effects / multilevel models |
| **psych** | Psychometric tools: factor analysis, reliability, descriptives |
| **lavaan** | Structural equation modeling |
| **brms** | Bayesian regression models (via Stan) |


## Reproducibility {.unnumbered}

| Package | Purpose |
|---------|---------|
| **rmarkdown** | Dynamic documents combining code and prose |
| **bookdown** | Write books and long documents (like these course notes!) |
| **here** | Project-relative file paths that work everywhere |
| **renv** | Package version management for reproducible environments |


## Useful Learning Resources {.unnumbered}

- [R for Data Science](https://r4ds.had.co.nz/) — free online textbook by Hadley Wickham & Garrett Grolemund
- [STAT 545](https://stat545.com/) — Jenny Bryan's excellent course on data wrangling
- [Data Science in a Box](https://datasciencebox.org/) — Mine Cetinkaya-Rundel's course materials
- [RStudio Cheatsheets](https://rstudio.com/resources/cheatsheets/) — printable quick-reference guides

```{r specXX-include-tweet, echo=FALSE}
try_include_tweet("https://twitter.com/hadleywickham/status/1363920786416562179")
```


## Fun Stuff {.unnumbered}

Because data science should be fun sometimes.

```{r specXX-fun-video, echo=FALSE}
# Gromit Placing Rail Tracks for 10 Hours
if (knitr::is_html_output()) {
  "https://www.youtube.com/watch?v=fwJHNw9jU_U" %>%
    embed_url() %>%
    use_align("center")
} else if (knitr::is_latex_output()) {
  embed_youtube_alt("fwJHNw9jU_U")
}
```

> "The best thing about R is that it was developed by statisticians. The worst thing about R is that... it was developed by statisticians." — Bo Cowgill


## How to Suggest Additions {.unnumbered}

Found a package or resource that should be on this list? Open a pull request on the [course repo][course_git] or let me know in class. This page is a living document — your contributions make it better for everyone.


```{r links-courselinks, child="includes/courselinks.md"}
```

```{r links-general, child="includes/links.md"}
```
