
```{r include = FALSE}
source("common.R")
library(tidyverse)

library(tweetrmd) #... embedding tweets

```



# (PART\*) Deep Dives and Unmodules {-}

# Data as rhetoric {-}

This is not a module. Although it is designed to introduce you to misleading graphics and data. Please watch the videos and work your way through the notes. **The videos start on the next page.**  You can find the video playlist for this module [here][pl_16]. The slides used to make the videos in this module can be found in the [slides repo][course_slides].


## Module Materials

* Slides
  * [Data and visualization][d03_dataviz]

* Suggested Readings
  * Darrel Huff

* Activities

* Lab

  
### Estimated Video Length

<!-- https://www.youtube.com/playlist?list=PLKrrdtYgOUYZmr_T3PnuxjVIlj0C0kUNI & https://ytplaylist-len.herokuapp.com/ -->

### Technically the Truth

```{r, echo=FALSE}
include_tweet("https://twitter.com/YuyaKiuchi/status/1357712469502885890")
```
# Reading Error Codes


```{r, echo=FALSE}
include_tweet("https://twitter.com/CMastication/status/1458412397371219973")
```
```{r links, child="admin/md/courselinks.md"}
```
